
<!--
Reproduces main Figure 1: locus schematic/ChIP track plots (Gviz) and binarized 200-bp peak-grid maps for the three time points (d0/d7/d15). Inputs (examples): output/fig1.all.k27.bw/fig1.{d0,d7,d15}.tab.tsv. Outputs: outplots/fig1*.pdf. Expects the project outputs described in the repo README.
--> 
fig1.plots

# ===================================================
# a. Schematic-style track plots around the EZH2 gene locus and the corresponding binarized grid map

```{r}
library(Gviz)
library(rtracklayer)
library(GenomicRanges)

# ==========================================
# 1. Parameter settings (adjust these to your own setup)
# ==========================================
gen <- "hg19"  # genome build
chr <- "chr7"  # target chromosome
start_pos <- 148580000  # start position
end_pos   <- 148590000  # end position

# Define the paths of the 8 BigWig files and the names to display
# Recommended to order them by biological logic (e.g. active marks -> repressive marks)
bw_files <- c(
  "DNase"    = "~/Project/ChIP/DNase/DNase-0d.10bp.CPM.bw",
  "H3K27ac"  = "~/Project/ChIP/H3K27ac/H3K27ac-0d.10bp.CPM.bw",
  "H3K4me1"  = "~/Project/ChIP/H3K4me1/H3K4me1-0d.10bp.CPM.bw",
  "H3K4me3"  = "~/Project/ChIP/H3K4me3/H3K4me3-0d.10bp.CPM.bw",
  "H3K9ac"   = "~/Project/ChIP/H3K9ac/H3K9ac-0d.10bp.CPM.bw",
  "H3K36me3" = "~/Project/ChIP/H3K36me3/H3K36me3-0d.10bp.CPM.bw",
  "H3K27me3" = "~/Project/ChIP/H3K27me3/H3K27me3-0d.10bp.CPM.bw",
  "H3K9me3"  = "~/Project/ChIP/H3K9me3/H3K9me3-0d.10bp.CPM.bw"
)

# Define a refined color scheme (academic Morandi palette)
track_colors <- c(
  "#94a3b8", # DNase (neutral gray)
  "#f87171", # H3K27ac (coral red)
  "#4ade80", # H3K4me1 (fresh green)
  "#fbbf24", # H3K4me3 (golden yellow)
  "#fb923c", # H3K9ac (orange)
  "#0d9488", # H3K36me3 (teal)
  "#60a5fa", # H3K27me3 (sky blue)
  "#818cf8"  # H3K9me3 (light purple)
)

# --- Key step: preset the Y-axis upper limit for the 8 tracks ---
# Adjust these values according to the signal strength of your actual data
# If you want all tracks to share one scale, set them to the same value
max_values <- c(20, 0.3, 0.4, 1.0, 0.5, 0.4, 1.0, 0.4) 

# ==========================================
# 2. Build the tracks (Tracks)
# ==========================================

# A. Genome axis track (at the very top)
axisTrack <- GenomeAxisTrack(
  exponent = 0, 
  cex = 0.8, 
  col = "black", 
  fontcolor = "black"
)

# B. List of data tracks
data_tracks <- list()

for (i in seq_along(bw_files)) {
  track_name <- names(bw_files)[i]
  bw_path <- bw_files[i]
  
  # Set the DataTrack parameters as follows:
data_tracks[[track_name]] <- DataTrack(
    range = bw_path,
    genome = gen,
    name = track_name,
    type = "poly",          # filled area plot, resembling the peaks in a schematic
    window = "mean",        # smooth using the mean
    windowSize = 200,        # window size; if it is still not smooth, try 100 or 200
    aggregation = "mean",
    
    # --- Independent Y-axis settings ---
    ylim = c(0, max_values[i]), # key: use the preset i-th upper limit
    showAxis = TRUE,            # show Y-axis ticks
    col.axis = "black",         # tick color
    cex.axis = 0.8,             # tick label font size    ylim = c(0, current_max),      # set the range
    yTicksAt = c(0, max_values[i]),  # key: draw tick marks only at 0 and the maximum

    fill.mountain = c(track_colors[i], track_colors[i]), # fill color
    col.mountain = track_colors[i],                      # border color
    background.title = "white",    # set the label background to white
    col.axis = "black",            # axis color
    col.title = "black",           # label text color
    fontface.title = 2,            # bold label
    cex.title = 0,               # label size
    alpha = 0.8,                   # semi-transparency for a more refined look
    lwd = 1.5                        # line width
  )

}

txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

# Create the gene track (Gene Track)
# This step requires no internet connection and is extremely fast and error-free
biomTrack <- GeneRegionTrack(
  txdb,
  genome = "hg19",
  chromosome = chr,
  start = start_pos,
  end = end_pos,
  name = "Gene",
  
  # --- Visual polish ---
  transcriptAnnotation = "symbol", # display gene names (e.g. MYC)
  collapseTranscripts = "longest",  # show only the longest transcript to keep the track uncluttered
  shape = "arrow",                  # indicate gene orientation with arrows
  fill = "#475569",                 # Morandi dark gray, more refined than pure black
  col = "transparent",              # hide the border
  background.title = "white",
  col.title = "black",
  fontface.title = 2,
  cex.title = 0.7,
  size = 0.5                        # adjust the height ratio of this track
)


# ==========================================
# 3. Plotting and export
# ==========================================

# Combine all tracks
all_tracks <- c(axisTrack, data_tracks, biomTrack)

# Export as PDF (convenient for further polishing in AI)
pdf("outplots/fig1.a_Epigenome_Tracks.pdf", width = 8, height = 6)

plotTracks(
  all_tracks,
  from = start_pos,
  to = end_pos,
  showTitle = FALSE,
  chromosome = chr,
  transcriptAnnotation = "symbol",
  window = -1,           # automatic smoothing
  cex.main = 1.2,
  margin = 40,           # white space
  innerMargin = 5,
  title.width = 1.5      # width of the left label column
)

dev.off()

# ==========================================
# 4. Corresponding binarized grid map
# ==========================================

peak <- paste0("chr19:33853000:33859000")
# Extract the raw binarized matrix
peak_tab <- msc.all[BedInter(peak,msc.all),4:11] %>% t()
library(ggplot2)
library(tidyr)
library(dplyr)

# 1. Prepare the data (assuming your matrix is called peak_tab)
# Explicitly specify your track order (top to bottom)
target_order <- c("H3K27me3","DNase", "H3K27ac", "H3K4me1", "H3K4me3", "H3K9ac", 
                  "H3K36me3", "H3K9me3")

peak_df <- as.data.frame(peak_tab)
peak_df <- read.table(file = "output/fig1.all.k27.bw/fig1.d0.tab.tsv")
colnames(peak_df) <- colnames(peak_tab)
peak_df$Track <- rownames(peak_tab)

# 2. Convert to long format and handle the ordering
plot_data <- peak_df %>%
  pivot_longer(cols = -Track, names_to = "Position", values_to = "Value") %>%
  mutate(Position = as.numeric(Position)) %>%
  # Key point 1: force the top-to-bottom order through factor and its levels
  # rev() is needed because ggplot draws the Y axis from bottom to top; rev keeps the first item at the top
  mutate(Track = factor(Track, levels = rev(target_order))) %>%
  filter(Value == 1)

# 3. Define colors (kept consistent with the previous section)
track_colors <- c(
  "DNase" = "#94a3b8", # DNase (neutral gray)
  "H3K27ac" = "#f87171", # H3K27ac (coral red)
  "H3K4me1" = "#4ade80", # H3K4me1 (fresh green)
  "H3K4me3" = "#fbbf24", # H3K4me3 (golden yellow)
  "H3K9ac" = "#fb923c", # H3K9ac (orange)
  "H3K36me3" = "#0d9488", # H3K36me3 (teal)
  "H3K27me3" = "#60a5fa", # H3K27me3 (sky blue)
  "H3K9me3" = "#818cf8"  # H3K9me3 (light purple)
)

# 4. Plot
p <- ggplot(plot_data, aes(x = Position, y = Track, fill = Track)) +
  # Key points 2 & 4:
  # height = 1 ensures there are no gaps between rows
  # color = "lightgray" adds light gray inner border lines
  # size = 0.1 makes the border lines very thin so they do not obscure the colors
  geom_tile(height = 1, color = "#e5e7eb", size = 0.1) + 
  
  # Set the fill colors
  scale_fill_manual(values = track_colors) +
  
  # Key point 3: no text labels at all
  theme_void() + # theme_void removes all axes, the background and text
  theme(
    legend.position = "none",
    # If you need to keep a small outer margin for alignment, uncomment the line below
    # plot.margin = margin(0, 0, 0, 0) 
  ) +
  
  # Keep the X-axis range exactly consistent with the peak plot
  scale_x_continuous(expand = c(0, 0)) +
  # Keep the Y axis compact with no extra white space
  scale_y_discrete(expand = c(0, 0))

# 5. Output
# It is recommended to control the final shape by specifying the width and height
print(p)
ggsave("outplots/fig1/fig1.a_peak_grid_map.d0.pdf", p, width = 7, height = 7)



{
peak <- paste0("chr19:33853000:33859000")
# Extract the raw binarized matrix
peak_tab <- adi_7d.all[BedInter(peak,msc.all),4:11] %>% t()
library(ggplot2)
library(tidyr)
library(dplyr)

# 1. Prepare the data (assuming your matrix is called peak_tab)
# Explicitly specify your track order (top to bottom)
target_order <- c("H3K27me3","DNase", "H3K27ac", "H3K4me1", "H3K4me3", "H3K9ac", 
                  "H3K36me3",  "H3K9me3")

peak_df <- as.data.frame(peak_tab)
peak_df <- read.table(file = "output/fig1.all.k27.bw/fig1.d7.tab.tsv")
colnames(peak_df) <- colnames(peak_tab)
peak_df$Track <- rownames(peak_tab)

# 2. Convert to long format and handle the ordering
plot_data <- peak_df %>%
  pivot_longer(cols = -Track, names_to = "Position", values_to = "Value") %>%
  mutate(Position = as.numeric(Position)) %>%
  # Key point 1: force the top-to-bottom order through factor and its levels
  # rev() is needed because ggplot draws the Y axis from bottom to top; rev keeps the first item at the top
  mutate(Track = factor(Track, levels = rev(target_order))) %>%
  filter(Value == 1)

# 3. Define colors (kept consistent with the previous section)
track_colors <- c(
  "DNase" = "#94a3b8", # DNase (neutral gray)
  "H3K27ac" = "#f87171", # H3K27ac (coral red)
  "H3K4me1" = "#4ade80", # H3K4me1 (fresh green)
  "H3K4me3" = "#fbbf24", # H3K4me3 (golden yellow)
  "H3K9ac" = "#fb923c", # H3K9ac (orange)
  "H3K36me3" = "#0d9488", # H3K36me3 (teal)
  "H3K27me3" = "#60a5fa", # H3K27me3 (sky blue)
  "H3K9me3" = "#818cf8"  # H3K9me3 (light purple)
)

# 4. Plot
p <- ggplot(plot_data, aes(x = Position, y = Track, fill = Track)) +
  # Key points 2 & 4:
  # height = 1 ensures there are no gaps between rows
  # color = "lightgray" adds light gray inner border lines
  # size = 0.1 makes the border lines very thin so they do not obscure the colors
  geom_tile(height = 1, color = "#e5e7eb", size = 0.1) + 
  
  # Set the fill colors
  scale_fill_manual(values = track_colors) +
  
  # Key point 3: no text labels at all
  theme_void() + # theme_void removes all axes, the background and text
  theme(
    legend.position = "none",
    # If you need to keep a small outer margin for alignment, uncomment the line below
    # plot.margin = margin(0, 0, 0, 0) 
  ) +
  
  # Keep the X-axis range exactly consistent with the peak plot
  scale_x_continuous(expand = c(0, 0)) +
  # Keep the Y axis compact with no extra white space
  scale_y_discrete(expand = c(0, 0))

# 5. Output
# It is recommended to control the final shape by specifying the width and height
print(p)
ggsave("outplots/fig1/fig1.a_peak_grid_map.d7.pdf", p, width = 7, height = 7)
}


{
peak <- paste0("chr19:33853000:33859000")
# Extract the raw binarized matrix
peak_tab <- adi_15d.all[BedInter(peak,msc.all),4:11] %>% t()
peak_df <- read.table(file = "output/fig1.all.k27.bw/fig1.d15.tab.tsv")
colnames(peak_df) <- colnames(peak_tab)
peak_df$Track <- rownames(peak_tab)

# 1. Prepare the data (assuming your matrix is called peak_tab)
# Explicitly specify your track order (top to bottom)
target_order <- c("H3K27me3", "DNase", "H3K27ac", "H3K4me1", "H3K4me3", "H3K9ac", 
                  "H3K36me3", "H3K9me3")


# 2. Convert to long format and handle the ordering
plot_data <- peak_df %>%
  pivot_longer(cols = -Track, names_to = "Position", values_to = "Value") %>%
  mutate(Position = as.numeric(Position)) %>%
  # Key point 1: force the top-to-bottom order through factor and its levels
  # rev() is needed because ggplot draws the Y axis from bottom to top; rev keeps the first item at the top
  mutate(Track = factor(Track, levels = rev(target_order))) %>%
  filter(Value == 1)

# 3. Define colors (kept consistent with the previous section)
track_colors <- c(
  "DNase" = "#94a3b8", # DNase (neutral gray)
  "H3K27ac" = "#f87171", # H3K27ac (coral red)
  "H3K4me1" = "#4ade80", # H3K4me1 (fresh green)
  "H3K4me3" = "#fbbf24", # H3K4me3 (golden yellow)
  "H3K9ac" = "#fb923c", # H3K9ac (orange)
  "H3K36me3" = "#0d9488", # H3K36me3 (teal)
  "H3K27me3" = "#60a5fa", # H3K27me3 (sky blue)
  "H3K9me3" = "#818cf8"  # H3K9me3 (light purple)
)

# 4. Plot
p <- ggplot(plot_data, aes(x = Position, y = Track, fill = Track)) +
  # Key points 2 & 4:
  # height = 1 ensures there are no gaps between rows
  # color = "lightgray" adds light gray inner border lines
  # size = 0.1 makes the border lines very thin so they do not obscure the colors
  geom_tile(height = 1, color = "#e5e7eb", size = 0.1) + 
  
  # Set the fill colors
  scale_fill_manual(values = track_colors) +
  
  # Key point 3: no text labels at all
  theme_void() + # theme_void removes all axes, the background and text
  theme(
    legend.position = "none",
    # If you need to keep a small outer margin for alignment, uncomment the line below
    # plot.margin = margin(0, 0, 0, 0) 
  ) +
  
  # Keep the X-axis range exactly consistent with the peak plot
  scale_x_continuous(expand = c(0, 0)) +
  # Keep the Y axis compact with no extra white space
  scale_y_discrete(expand = c(0, 0))

# 5. Output
# It is recommended to control the final shape by specifying the width and height
print(p)
ggsave("outplots/fig1/fig1.a_peak_grid_map.d14.pdf", p, width = 7, height = 7)
}

plot_data <- rbind(plot_data,c("H3K9me3", 6911194, 1))
plot_data$Position <- as.integer(plot_data$Position)
plot_data$Value <- as.integer(plot_data$Value)


```


# ====================================================
## b. Relationship between the mean read count and the length of all contiguous regions at each time point

```{r}
msc.k27.all <- k27.env$msc.k27.all
adi_7d.k27.all <- k27.env$adi_7d.k27.all
adi_15d.k27.all <- k27.env$adi_15d.k27.all


out1 <- data.frame(msc.k27.all$chr,msc.k27.all$start,msc.k27.all$end,name=paste0(msc.k27.all$chr,":",msc.k27.all$start,":",msc.k27.all$end))
out2 <- data.frame(adi_7d.k27.all$chr,adi_7d.k27.all$start,adi_7d.k27.all$end,name=paste0(adi_7d.k27.all$chr,":",adi_7d.k27.all$start,":",adi_7d.k27.all$end))
out3 <- data.frame(adi_15d.k27.all$chr,adi_15d.k27.all$start,adi_15d.k27.all$end,name=paste0(adi_15d.k27.all$chr,":",adi_15d.k27.all$start,":",adi_15d.k27.all$end))
write.table(out1,file = "output/msc.k27.all.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(out2,file = "output/adi_7d.k27.all.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(out3,file = "output/adi_15d.k27.all.bed",row.names = F,col.names = F,sep = "\t",quote = F)


## In addition, the non-contiguous i.e. isolated regions with a read count of 1 also need to be included:
l1 <- msc.all[msc.all$H3K27me3 == 1,1:3]
msc.inde.k27 <- bt.intersect(a=l1,b=msc.k27.all,v=T)
msc.inde.k27$name <- paste0(msc.inde.k27[,1],":",msc.inde.k27[,2],":",msc.inde.k27[,3])
write.table(msc.inde.k27,file = "output/fig1.all.k27.bw/msc.inde.k27.bed",row.names = F,col.names = F,sep = "\t",quote = F)
l2 <- adi_7d.all[adi_7d.all$H3K27me3 == 1,1:3]
adi_7d.inde.k27 <- bt.intersect(a=l2,b=adi_7d.k27.all,v=T)
adi_7d.inde.k27$name <- paste0(adi_7d.inde.k27[,1],":",adi_7d.inde.k27[,2],":",adi_7d.inde.k27[,3])
write.table(adi_7d.inde.k27,file = "output/fig1.all.k27.bw/adi_7d.inde.k27.bed",row.names = F,col.names = F,sep = "\t",quote = F)
l3 <- adi_15d.all[adi_15d.all$H3K27me3 == 1,1:3]
adi_15d.inde.k27 <- bt.intersect(a=l3,b=adi_15d.k27.all,v=T)
adi_15d.inde.k27$name <- paste0(adi_15d.inde.k27[,1],":",adi_15d.inde.k27[,2],":",adi_15d.inde.k27[,3])
write.table(adi_15d.inde.k27,file = "output/fig1.all.k27.bw/adi_15d.inde.k27.bed",row.names = F,col.names = F,sep = "\t",quote = F)


## The comparison starts below; first a simple comparison for MSC
msc.inde.k27.bw <- read.table(file = "output/fig1.all.k27.bw/msc.inde.k27.out.tab")
msc.k27.all.bw <- read.table(file = "output/fig1.all.k27.bw/msc.k27.all.out.tab")

boxplot(log2(msc.inde.k27.bw$V5+1),log2(msc.k27.all.bw$V5+1))
## Regions in contiguous stretches are certainly significantly higher than isolated ones
msc.all.bw <- rbind(msc.inde.k27.bw,msc.k27.all.bw)
plot(msc.all.bw[,2]/200,log2(msc.all.bw$V5+1))
cor.test(msc.all.bw[,2]/200,log2(msc.all.bw$V5+1))
lm(data.frame(x=msc.all.bw[,2]/200,y=log2(msc.all.bw$V5+1))) %>% summary()
## There is a fairly significant positive correlation, so plots are needed for all three time points


## Boxplot of signal strength versus length, grouped by length
msc.all.bw$len <- msc.all.bw$V2/200
msc.all.bw$V5 <- log2(msc.all.bw$V5+1)

boxplot(msc.all.bw[msc.all.bw$len == 1,"V5"],msc.all.bw[msc.all.bw$len == 2,"V5"],msc.all.bw[msc.all.bw$len == 3,"V5"],msc.all.bw[msc.all.bw$len == 4,"V5"],msc.all.bw[msc.all.bw$len == 5,"V5"],msc.all.bw[msc.all.bw$len == 6,"V5"],msc.all.bw[msc.all.bw$len == 7,"V5"],msc.all.bw[msc.all.bw$len == 8,"V5"],msc.all.bw[msc.all.bw$len == 9,"V5"],msc.all.bw[msc.all.bw$len == 10,"V5"],msc.all.bw[msc.all.bw$len == 10,"V5"])

boxplot(msc.all.bw[msc.all.bw$len == 1,"V5"],msc.all.bw[msc.all.bw$len %in% c(2,3,4,5),"V5"],msc.all.bw[msc.all.bw$len %in% c(6:10),"V5"],msc.all.bw[msc.all.bw$len %in% c(11:20),"V5"],msc.all.bw[msc.all.bw$len > 20,"V5"])


adi_7d.inde.k27.bw <- read.table(file = "output/fig1.all.k27.bw/adi_7d.inde.k27.out.tab")
adi_7d.k27.all.bw <- read.table(file = "output/fig1.all.k27.bw/adi_7d.k27.all.out.tab")

boxplot(log2(adi_7d.inde.k27.bw$V5+1),log2(adi_7d.k27.all.bw$V5+1))
## Regions in contiguous stretches are certainly significantly higher than isolated ones
adi_7d.all.bw <- rbind(adi_7d.inde.k27.bw,adi_7d.k27.all.bw)
plot(adi_7d.all.bw[,2]/200,log2(adi_7d.all.bw$V5+1))
cor.test(adi_7d.all.bw[,2]/200,log2(adi_7d.all.bw$V5+1))
lm(data.frame(x=adi_7d.all.bw[,2]/200,y=log2(adi_7d.all.bw$V5+1))) %>% summary()

adi_7d.all.bw$len <- adi_7d.all.bw$V2/200
adi_7d.all.bw$V5 <- log2(adi_7d.all.bw$V5+1)
boxplot(adi_7d.all.bw[adi_7d.all.bw$len == 1,"V5"],adi_7d.all.bw[adi_7d.all.bw$len %in% c(2,3,4,5),"V5"],adi_7d.all.bw[adi_7d.all.bw$len %in% c(6:10),"V5"],adi_7d.all.bw[adi_7d.all.bw$len %in% c(11:20),"V5"],adi_7d.all.bw[adi_7d.all.bw$len > 20,"V5"])






adi_15d.inde.k27.bw <- read.table(file = "output/fig1.all.k27.bw/adi_15d.inde.k27.out.tab")
adi_15d.k27.all.bw <- read.table(file = "output/fig1.all.k27.bw/adi_15d.k27.all.out.tab")

boxplot(log2(adi_15d.inde.k27.bw$V5+1),log2(adi_15d.k27.all.bw$V5+1))
## Regions in contiguous stretches are certainly significantly higher than isolated ones
adi_15d.all.bw <- rbind(adi_15d.inde.k27.bw,adi_15d.k27.all.bw)
plot(adi_15d.all.bw[,2]/200,log2(adi_15d.all.bw$V5+1))
cor.test(adi_15d.all.bw[,2]/200,log2(adi_15d.all.bw$V5+1))
lm(data.frame(x=adi_15d.all.bw[,2]/200,y=log2(adi_15d.all.bw$V5+1))) %>% summary()

adi_15d.all.bw$len <- adi_15d.all.bw$V2/200
adi_15d.all.bw$V5 <- log2(adi_15d.all.bw$V5+1)
boxplot(adi_15d.all.bw[adi_15d.all.bw$len == 1,"V5"],adi_15d.all.bw[adi_15d.all.bw$len %in% c(2,3,4,5),"V5"],adi_15d.all.bw[adi_15d.all.bw$len %in% c(6:10),"V5"],adi_15d.all.bw[adi_15d.all.bw$len %in% c(11:20),"V5"],adi_15d.all.bw[adi_15d.all.bw$len > 20,"V5"])


library(ggplot2)
library(dplyr)

## =========================
## Parameters you may tweak
## =========================
COL <- "#0072B2"        # firebrick red, a deep red commonly used in papers
k_extreme <- 5          # IQR multiplier: the larger it is, the more only the "extreme few" points are removed. 1.5 is common; 5 here is closer to removing only the "handful of most extreme" points
seed <- 1               # for reproducible scatter downsampling
max_points_to_show <- 1e+8  # maximum number of points shown in the scatter plot (fitting still uses all data after extreme-value removal)

stopifnot(all(c("len","V5") %in% colnames(msc.all.bw)))

## =========================
## 0) Remove extreme outliers (based on V5)
##    Using robust IQR rule: [Q1 - k*IQR, Q3 + k*IQR]
## =========================
df0 <- msc.all.bw %>% dplyr::filter(!is.na(len), !is.na(V5))

q1  <- as.numeric(quantile(df0$V5, 0.25, na.rm = TRUE, type = 7))
q3  <- as.numeric(quantile(df0$V5, 0.75, na.rm = TRUE, type = 7))
iqr <- q3 - q1

lower <- q1 - k_extreme * iqr
upper <- q3 + k_extreme * iqr

df <- df0 %>% dplyr::filter(V5 >= lower, V5 <= upper)

cat("Filtered rows:", nrow(df0) - nrow(df), "removed; kept:", nrow(df), "\n")
cat("V5 extreme filter range: [", lower, ", ", upper, "]\n", sep = "")

## =========================
## Boxplot script only (based on your filtered data df)
## Assumes you already have: df  (after extreme-outlier filtering)
## Requires: ggplot2, dplyr
## =========================

library(ggplot2)
library(dplyr)

## 1) make the same 5 length groups
df_box <- df %>%
  mutate(len_group = case_when(
    len == 1 ~ "1",
    len %in% 2:5 ~ "2-5",
    len %in% 6:10 ~ "6-10",
    len %in% 11:20 ~ "11-20",
    len > 20 ~ ">20",
    TRUE ~ NA_character_
  )) %>%
  dplyr::filter(!is.na(len_group), !is.na(V5)) %>%
  mutate(len_group = factor(len_group, levels = c("1", "2-5", "6-10", "11-20", ">20")))

## 2) flag outliers per group using standard boxplot rule (1.5*IQR)
##    and then downsample outliers to avoid “too many” points shown
outlier_frac <- 0.15   # show at most 15% outliers per group (tune)
outlier_cap  <- 200    # and cap to <=200 outliers per group (tune)
set.seed(1)

df_outliers <- df_box %>%
  group_by(len_group) %>%
  mutate(
    q1 = quantile(V5, 0.25, na.rm = TRUE),
    q3 = quantile(V5, 0.75, na.rm = TRUE),
    iqr = q3 - q1,
    low = q1 - 1.5 * iqr,
    high = q3 + 1.5 * iqr,
    is_outlier = (V5 < low | V5 > high)
  ) %>%
  ungroup() %>%
  dplyr::filter(is_outlier) %>%
  group_by(len_group) %>%
  slice_sample(
    n = 200
  ) %>%
  ungroup()

## 3) main boxplot: hide built-in outliers, then add our sampled outliers
##    + add 25% and 75% "error bars" (Q1/Q3 whiskers)
p1 <- ggplot(df_box, aes(x = len_group, y = V5, fill = len_group)) +
  stat_boxplot(geom = "errorbar",
               width=0.4)+
  geom_boxplot(
    outlier.shape = NA,   # suppress default outlier drawing
    width = 0.65,
    color = "black",
    linewidth = 0.5
  ) +
  ## add sampled outliers as hollow black circles
  geom_point(
    data = df_outliers,
    aes(x = len_group, y = V5),
    inherit.aes = FALSE,
    shape = 21,          # hollow circle
    color = "black",
    fill = "white",
    size = 0.5,
    stroke = 0.3,
    position = position_jitter(width = 0.15, height = 0)
  ) +
  ## light red -> dark red gradient across groups
  scale_fill_manual(
    values = c(  "1"      = "#D6E8F9", "2-5"    = "#A0CCE3", "6-10"   = "#5FA3D0", "11-20"  = "#1F77B4", ">20"    = "#0E4C8C")
  ) +
  labs(
    x = "Length group",
    y = "H3K27me3 log2(CPM+1)",
    title = ""
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "none"
  )


## =========================
## 2) Scatter: len vs V5
##    - show only a sample of points
##    - fit on ALL filtered data
## =========================
set.seed(seed)
n_all <- nrow(df)

df_points <- if (n_all > max_points_to_show) {
  df %>% slice_sample(n = max_points_to_show)
} else {
  df
}

## Stats computed on filtered full data (NOT sampled)
ct  <- cor.test(df$len, df$V5, method = "pearson")
fit <- lm(V5 ~ len, data = df)
r2  <- summary(fit)$r.squared

label_txt <- sprintf(
  "Pearson r = %.3f\np = %.3g\nR² = %.3f\nn = %d",
  unname(ct$estimate), ct$p.value, r2, n_all, nrow(df_points)
)

p2 <- ggplot(df, aes(x = len, y = V5)) +
  ## points: use sampled subset, but keep the same mapping
  scattermore::geom_scattermore(
    data = df_points,
    color = COL,
    #alpha = 0.8,   # more transparent
    pointsize = 3,
    pixels = c(1000,1000)
  ) +
  ## fit line on full filtered data
#   geom_smooth(
#     method = "lm",
#     se = TRUE,
#     color = COL,
#     fill = COL,
#     alpha = 0.18,
#     linewidth = 1
#   ) +
  annotate(
    "text",
    x = Inf, y = Inf,
    label = label_txt,
    hjust = 1.05, vjust = 1.1,
    size = 4,
    color = "black"
  ) +
  labs(
    x = "Length",
    y = "H3K27me3 log2(CPM+1)",
    title = ""
  ) +
  theme_bw(base_size = 16)

## --- Show plots ---
p <- p1 + p2 + patchwork::plot_layout(widths = c(1, 1))
ggsave("outplots/fig1/fig1_b.H3K27me3.signal&length.boxplot&scatter.d0.pdf",p,width=8,height=4,dpi=300)






## =========================
## 0) Remove extreme outliers (based on V5)
##    Using robust IQR rule: [Q1 - k*IQR, Q3 + k*IQR]
## =========================
df0 <- adi_7d.all.bw %>% dplyr::filter(!is.na(len), !is.na(V5))
COL <- "#E69F00" 

q1  <- as.numeric(quantile(df0$V5, 0.25, na.rm = TRUE, type = 7))
q3  <- as.numeric(quantile(df0$V5, 0.75, na.rm = TRUE, type = 7))
iqr <- q3 - q1

lower <- q1 - k_extreme * iqr
upper <- q3 + k_extreme * iqr

df <- df0 %>% dplyr::filter(V5 >= lower, V5 <= upper)

cat("Filtered rows:", nrow(df0) - nrow(df), "removed; kept:", nrow(df), "\n")
cat("V5 extreme filter range: [", lower, ", ", upper, "]\n", sep = "")

## =========================
## Boxplot script only (based on your filtered data df)
## Assumes you already have: df  (after extreme-outlier filtering)
## Requires: ggplot2, dplyr
## =========================

library(ggplot2)
library(dplyr)

## 1) make the same 5 length groups
df_box <- df %>%
  mutate(len_group = case_when(
    len == 1 ~ "1",
    len %in% 2:5 ~ "2-5",
    len %in% 6:10 ~ "6-10",
    len %in% 11:20 ~ "11-20",
    len > 20 ~ ">20",
    TRUE ~ NA_character_
  )) %>%
  dplyr::filter(!is.na(len_group), !is.na(V5)) %>%
  mutate(len_group = factor(len_group, levels = c("1", "2-5", "6-10", "11-20", ">20")))

## 2) flag outliers per group using standard boxplot rule (1.5*IQR)
##    and then downsample outliers to avoid “too many” points shown
outlier_frac <- 0.15   # show at most 15% outliers per group (tune)
outlier_cap  <- 200    # and cap to <=200 outliers per group (tune)
set.seed(1)

df_outliers <- df_box %>%
  group_by(len_group) %>%
  mutate(
    q1 = quantile(V5, 0.25, na.rm = TRUE),
    q3 = quantile(V5, 0.75, na.rm = TRUE),
    iqr = q3 - q1,
    low = q1 - 1.5 * iqr,
    high = q3 + 1.5 * iqr,
    is_outlier = (V5 < low | V5 > high)
  ) %>%
  ungroup() %>%
  dplyr::filter(is_outlier) %>%
  group_by(len_group) %>%
  slice_sample(
    n = 200
  ) %>%
  ungroup()

## 3) main boxplot: hide built-in outliers, then add our sampled outliers
##    + add 25% and 75% "error bars" (Q1/Q3 whiskers)
p1 <- ggplot(df_box, aes(x = len_group, y = V5, fill = len_group)) +
  stat_boxplot(geom = "errorbar",
               width=0.4)+
  geom_boxplot(
    outlier.shape = NA,   # suppress default outlier drawing
    width = 0.65,
    color = "black",
    linewidth = 0.5
  ) +
  ## add sampled outliers as hollow black circles
  geom_point(
    data = df_outliers,
    aes(x = len_group, y = V5),
    inherit.aes = FALSE,
    shape = 1,          # hollow circle
    color = "black",
    fill = "white",
    size = 0.5,
    stroke = 0.3,
    position = position_jitter(width = 0.15, height = 0)
  ) +
  ## light red -> dark red gradient across groups
  scale_fill_manual(
    values = c("1" = "#FBE5B8", "2-5" = "#F3C463", "6-10" = "#E69F00", "11-20"  = "#B87F00", ">20" = "#8A5F00")
  ) +
  labs(
    x = "Length group",
    y = "H3K27me3 log2(CPM+1)",
    title = ""
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "none"
  )


## =========================
## 2) Scatter: len vs V5
##    - show only a sample of points
##    - fit on ALL filtered data
## =========================
set.seed(seed)
n_all <- nrow(df)

df_points <- if (n_all > max_points_to_show) {
  df %>% slice_sample(n = max_points_to_show)
} else {
  df
}

## Stats computed on filtered full data (NOT sampled)
ct  <- cor.test(df$len, df$V5, method = "pearson")
fit <- lm(V5 ~ len, data = df)
r2  <- summary(fit)$r.squared

label_txt <- sprintf(
  "Pearson r = %.3f\np = %.3g\nR² = %.3f\nn = %d",
  unname(ct$estimate), ct$p.value, r2, n_all, nrow(df_points)
)

p2 <- ggplot(df, aes(x = len, y = V5)) +
  ## points: use sampled subset, but keep the same mapping
  scattermore::geom_scattermore(
    data = df_points,
    color = COL,
    #alpha = 0.8,   # more transparent
    pointsize = 3,
    pixels = c(1000,1000)
  ) +
  coord_cartesian(xlim = c(0, 100)) + 
  ## fit line on full filtered data
#   geom_smooth(
#     method = "lm",
#     se = TRUE,
#     color = COL,
#     fill = COL,
#     alpha = 0.18,
#     linewidth = 1
#   ) +
  annotate(
    "text",
    x = Inf, y = Inf,
    label = label_txt,
    hjust = 1.05, vjust = 1.1,
    size = 4,
    color = "black"
  ) +
  labs(
    x = "Length",
    y = "H3K27me3 log2(CPM+1)",
    title = ""
  ) +
  theme_bw(base_size = 16)

## --- Show plots ---
p <- p1 + p2 + patchwork::plot_layout(widths = c(1, 1))
ggsave("outplots/fig1/fig1_b.H3K27me3.signal&length.boxplot&scatter.d7.pdf",p,width=8,height=4,dpi=300)


## =========================
## 0) Remove extreme outliers (based on V5)
##    Using robust IQR rule: [Q1 - k*IQR, Q3 + k*IQR]
## =========================
df0 <- adi_15d.all.bw %>% dplyr::filter(!is.na(len), !is.na(V5))
COL <- "#CC79A7"
q1  <- as.numeric(quantile(df0$V5, 0.25, na.rm = TRUE, type = 7))
q3  <- as.numeric(quantile(df0$V5, 0.75, na.rm = TRUE, type = 7))
iqr <- q3 - q1

lower <- q1 - k_extreme * iqr
upper <- q3 + k_extreme * iqr

df <- df0 %>% dplyr::filter(V5 >= lower, V5 <= upper)

cat("Filtered rows:", nrow(df0) - nrow(df), "removed; kept:", nrow(df), "\n")
cat("V5 extreme filter range: [", lower, ", ", upper, "]\n", sep = "")

## =========================
## Boxplot script only (based on your filtered data df)
## Assumes you already have: df  (after extreme-outlier filtering)
## Requires: ggplot2, dplyr
## =========================

library(ggplot2)
library(dplyr)

## 1) make the same 5 length groups
df_box <- df %>%
  mutate(len_group = case_when(
    len == 1 ~ "1",
    len %in% 2:5 ~ "2-5",
    len %in% 6:10 ~ "6-10",
    len %in% 11:20 ~ "11-20",
    len > 20 ~ ">20",
    TRUE ~ NA_character_
  )) %>%
  dplyr::filter(!is.na(len_group), !is.na(V5)) %>%
  mutate(len_group = factor(len_group, levels = c("1", "2-5", "6-10", "11-20", ">20")))

## 2) flag outliers per group using standard boxplot rule (1.5*IQR)
##    and then downsample outliers to avoid “too many” points shown
outlier_frac <- 0.15   # show at most 15% outliers per group (tune)
outlier_cap  <- 200    # and cap to <=200 outliers per group (tune)
set.seed(1)

df_outliers <- df_box %>%
  group_by(len_group) %>%
  mutate(
    q1 = quantile(V5, 0.25, na.rm = TRUE),
    q3 = quantile(V5, 0.75, na.rm = TRUE),
    iqr = q3 - q1,
    low = q1 - 1.5 * iqr,
    high = q3 + 1.5 * iqr,
    is_outlier = (V5 < low | V5 > high)
  ) %>%
  ungroup() %>%
  dplyr::filter(is_outlier) %>%
  group_by(len_group) %>%
  slice_sample(
    n = 200
  ) %>%
  ungroup()

## 3) main boxplot: hide built-in outliers, then add our sampled outliers
##    + add 25% and 75% "error bars" (Q1/Q3 whiskers)
p1 <- ggplot(df_box, aes(x = len_group, y = V5, fill = len_group)) +
  stat_boxplot(geom = "errorbar",
               width=0.4)+
  geom_boxplot(
    outlier.shape = NA,   # suppress default outlier drawing
    width = 0.65,
    color = "black",
    linewidth = 0.5
  ) +
  ## add sampled outliers as hollow black circles
  geom_point(
    data = df_outliers,
    aes(x = len_group, y = V5),
    inherit.aes = FALSE,
    shape = 1,          # hollow circle
    color = "black",
    size = 0.5,
    stroke = 0.3,
    position = position_jitter(width = 0.15, height = 0)
  ) +
  ## light red -> dark red gradient across groups
  scale_fill_manual(
    values = c("1" = "#F0D4E8", "2-5" = "#DDA5CA", "6-10" = "#CC79A7",
               "11-20" = "#A36185", ">20" = "#7A4864")
  ) +
  labs(
    x = "Length group",
    y = "H3K27me3 log2(CPM+1)",
    title = ""
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "none"
  )


## =========================
## 2) Scatter: len vs V5
##    - show only a sample of points
##    - fit on ALL filtered data
## =========================
set.seed(seed)
n_all <- nrow(df)

df_points <- if (n_all > max_points_to_show) {
  df %>% slice_sample(n = max_points_to_show)
} else {
  df
}

## Stats computed on filtered full data (NOT sampled)
ct  <- cor.test(df$len, df$V5, method = "pearson")
fit <- lm(V5 ~ len, data = df)
r2  <- summary(fit)$r.squared

label_txt <- sprintf(
  "Pearson r = %.3f\np = %.3g\nR² = %.3f\nn = %d",
  unname(ct$estimate), ct$p.value, r2, n_all, nrow(df_points)
)

p2 <- ggplot(df, aes(x = len, y = V5)) +
  ## points: use sampled subset, but keep the same mapping
  scattermore::geom_scattermore(
    data = df_points,
    color = COL,
    #alpha = 0.8,   # more transparent
    pointsize = 3,
    pixels = c(1000,1000)
  )  +
  ## fit line on full filtered data
#   geom_smooth(
#     method = "lm",
#     se = TRUE,
#     color = COL,
#     fill = COL,
#     alpha = 0.18,
#     linewidth = 1
#   ) +
  annotate(
    "text",
    x = Inf, y = Inf,
    label = label_txt,
    hjust = 1.05, vjust = 1.1,
    size = 4,
    color = "black"
  ) +
  labs(
    x = "Length",
    y = "H3K27me3 log2(CPM+1)",
    title = ""
  ) +
  theme_bw(base_size = 16)

## --- Show plots ---
p <- p1 + p2 + patchwork::plot_layout(widths = c(1, 1))
ggsave("outplots/fig1/fig1_b.H3K27me3.signal&length.boxplot&scatter.d14.pdf",p,width=8,height=4,dpi=300)




```




# b. Choosing length = 7 as the ROIs
# The order is:
1. Show the count for each length;
2. Sensitivity analysis: for cutoff = 5, 6, 7, 8, 9, 10, compute separately:
    ROI count
    mean ROI length
    fraction of the genome covered by ROIs
    ROI mean/median H3K27me3 signal
    target gene expression level
    expression difference compared with control regions

# Distribution of length versus count
# This part has three figures; the two supplementary figures show the number of connections and the H3K27me3 signal of the connections compared with l == 1 and with the strength of the background signal
# The main figure shows the correlation analysis between length and signal strength
```{r}



# Sensitivity analysis: at thresholds top30, top20, top10 and top5 with the corresponding l>=4, l>=5, l>=7, l>=9, whether target gene expression differs and is lower than control in all cases, and the expression differences between groups
# Also the number of ROIs, the corresponding genome coverage and the corresponding number of target genes at different thresholds, showing that the differences between thresholds are in fact small and that all of them are repressive, while target gene expression is robustly reduced



# sup1_b.




# fig1_c.cumulative.pdf
k27.env$msc.k27.all[k27.env$msc.k27.all$len >1,"len"] %>% hist
hist(log2(k27.env$msc.k27.all[k27.env$msc.k27.all$len >1,"len"]))
hist(log2(k27.env$adi.7d.k27.all$len+1))

library(ggplot2)

# 1. Simulate/organize the data (replace with your actual variable names)
# Assume your three sample datasets are called data1, data2, data3
# This demonstrates how to merge them into the "long format" required by ggplot
df1 <- data.frame(val = k27.env$msc.k27.all[k27.env$msc.k27.all$len >1,"len"], group = "MSC")
df2 <- data.frame(val = k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >1,"len"], group = "ADI 7d")
df3 <- data.frame(val = k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >1,"len"], group = "ADI 14d")

plot_data <- rbind(df1, df2, df3)
plot_data$group <- factor(plot_data$group,levels=c("MSC","ADI 7d","ADI 14d"))
kruskal.test(val ~ group, data = plot_data)
rstatix::kruskal_effsize(val ~ group, data = plot_data)

## Although the Kruskal-Wallis test shows a statistically significant difference among the three groups (p < 2.2e-16), this is mainly caused by the very large sample size (n = 769,644). The effect size analysis gives η2 = 0.00086 (a very small magnitude), indicating that the length distributions of the three are biologically highly consistent.


# 2. Draw the polished density curve
library(ggplot2)

# 1. Construct demo data (replace with your own log2-transformed data)
# Assume your data have already been combined into a long-format data frame called plot_data
# If not, refer to the rbind step in the previous reply

# 2. Plot
ggplot(plot_data, aes(x = val, fill = group, color = group)) +
  # First layer: draw the histogram (provides the light-colored fill)
  # position = "identity" means overlapping placement rather than stacking
  # alpha is set to a light transparency
  geom_histogram(aes(y = ..count..), 
                 position = "identity", 
                 alpha = 0.2, 
                 binwidth = 0.5, # adjust binwidth according to the spacing of your data
                 color = NA) +   # this layer should have no border lines
  
  # Second layer: draw the frequency polygon (provides the thick curve)
  # This line connects the centers of the frequency points
  geom_freqpoly(aes(y = ..count..), 
                linewidth = 1.5, 
                binwidth = 0.5) + 
  
  # Color settings
  scale_fill_brewer(palette = "Set1") +
  scale_color_brewer(palette = "Set1") +
  
  # Visual polish
  theme_minimal(base_size = 15) +
  theme(
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "top",
    legend.title = element_blank()
  ) +
  labs(
    title = "Frequency Distribution (Discrete)",
    x = "log2(Length)",
    y = "Frequency (Count)"
  )



library(ggplot2)
library(scales) # this package is needed to handle tick labels

# Assume your raw data column is called len (the true length)
# If you only have log2-transformed data, restore it first: plot_data$len <- 2^(plot_data$val)

p <- ggplot(plot_data, aes(x = val, color = group)) +
  # 1. Draw the cumulative distribution curve (ECDF)
  stat_ecdf(linewidth = 1, alpha = 0.8) + 
  
  # 2. Key setting: use an X axis with log2 transformation but display the original values
  # Manually specify the lengths to highlight: 2, 4, 8, 16, 32, 64, 128, 256, 512, 1024
  scale_x_continuous(
    trans = "log2", 
    breaks = c(2, 4, 8, 16, 32, 64, 128, 256, 512, 1024), 
    labels = c("2", "4", "8", "16", "32", "64", "128", "256", "512", "1024")
  ) +
  
  # 3. Add a dashed reference line at y = 0.9
  geom_hline(yintercept = 0.9, linetype = "dashed", color = "gray40", linewidth = 0.8) +
  geom_vline(xintercept = 6, linetype = "dashed", color = "black", linewidth = 0.8) +
  
  # 4. Colors and theme
  scale_color_manual(values = c("#0072B2", "#E69F00" , "#CC79A7")) + # classic Set1 colors
  theme_bw(base_size = 16) +
  theme(
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "none",
    legend.title = element_blank(),
    axis.text= element_text(color="black") # tilt the labels to prevent overlap
  ) +
  
  # 5. Label settings
  labs(
    x = "Length (l)", 
    y = "Cumulative Probability"
  ) +
  
  # 6. Y-axis range optimization
  scale_y_continuous(breaks = seq(0, 1, 0.1), expand = c(0.01, 0.01))

ggsave(file = "outplots/fig1/sup1_a.cumulative.pdf",p,width=4,height=5,dpi=300)

# ==========================================
# Main figure fig1.c: the definition criterion of CRRs together with the counts
# ==========================================


df <- k27.env$msc.k27.all

library(ggplot2)

# Extract the data
df <- k27.env$msc.k27.all

p1 <- ggplot(df, aes(x = len)) +
  # Draw the bar chart
  geom_bar(fill = "#0072B2", color = "white", alpha = 0.8) +
  
  # Draw a dashed line at x = 7
  geom_vline(xintercept = 7, linetype = "dashed", color = "red", size = 0.8) +
  
  # Key settings:
  # 1. limits sets the range from 2 to 20
  # 2. expand = c(0, 0) forces the plot to hug the axis with no blank space
  # 3. breaks sets the tick positions
  scale_x_continuous(
    limits = c(1.5, 20.5), 
    #breaks = 2:20, 
    expand = c(0, 0)
  ) +
  
  # Likewise, for the y axis set expand so that the bottoms of the bars hug the x axis
  scale_y_continuous(
    limits = c(0, 180000),
    expand = expansion(mult = c(0, 0.05)) # 0 at the bottom and 5% at the top so the tallest bar does not touch the edge
  ) +
  
  # Set the labels
  labs(
    x = "length / bin", 
    y = "Count"
  ) +
  
  # Use the classic theme
  theme_classic(base_size = 12) +
  
  # Fine-tuning: make sure the axis line is aligned at position 2
  theme(
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "none",
    legend.title = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text= element_text(color="black"), # tilt the labels to prevent overlap
    plot.margin = margin(10, 20, 10, 10) # slightly increase the margins so that labels are not cut off
  )
ggsave("outplots/fig1/fig1_c.MSC.CRRs.counts.pdf",p1,width = 6,height = 2,dpi = 300)


df <- k27.env$adi_7d.k27.all

p2 <- ggplot(df, aes(x = len)) +
  # Draw the bar chart
  geom_bar(fill = "#E69F00", color = "white", alpha = 0.8) +
  
  # Draw a dashed line at x = 7
  #geom_vline(xintercept = 7, linetype = "dashed", color = "red", size = 0.8) +
  
  # Key settings:
  # 1. limits sets the range from 2 to 20
  # 2. expand = c(0, 0) forces the plot to hug the axis with no blank space
  # 3. breaks sets the tick positions
  scale_x_continuous(
    limits = c(1.5, 20.5), 
    #breaks = 2:20, 
    expand = c(0, 0)
  ) +
  
  # Likewise, for the y axis set expand so that the bottoms of the bars hug the x axis
  scale_y_continuous(
    limits = c(0, 180000),
    expand = expansion(mult = c(0, 0.05)) # 0 at the bottom and 5% at the top so the tallest bar does not touch the edge
  ) +
  
  # Set the labels
  labs(
    x = "length / bin", 
    y = "Count"
  ) +
  
  # Use the classic theme
  theme_classic(base_size = 12) +
  
  # Fine-tuning: make sure the axis line is aligned at position 2
  theme(
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "none",
    legend.title = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text= element_text(color="black"), # tilt the labels to prevent overlap
    plot.margin = margin(10, 20, 10, 10) # slightly increase the margins so that labels are not cut off
  )
ggsave("outplots/fig1/fig1_c.ADI_7d.CRRs.counts.pdf",p2,width = 6,height = 2,dpi = 300)


df <- k27.env$adi_15d.k27.all

p3 <- ggplot(df, aes(x = len)) +
  # Draw the bar chart
  geom_bar(fill = "#CC79A7", color = "white", alpha = 0.8) +
  
  # Draw a dashed line at x = 7
  geom_vline(xintercept = 7, linetype = "dashed", color = "red", size = 0.8) +
  
  # Key settings:
  # 1. limits sets the range from 2 to 20
  # 2. expand = c(0, 0) forces the plot to hug the axis with no blank space
  # 3. breaks sets the tick positions
  scale_x_continuous(
    limits = c(1.5, 20.5), 
    #breaks = 2:20, 
    expand = c(0, 0)
  ) +
  
  # Likewise, for the y axis set expand so that the bottoms of the bars hug the x axis
  scale_y_continuous(
    limits = c(0, 180000),
    expand = expansion(mult = c(0, 0.05)) # 0 at the bottom and 5% at the top so the tallest bar does not touch the edge
  ) +
  
  # Set the labels
  labs(
    x = "length / bin", 
    y = "Count"
  ) +
  
  # Use the classic theme
  theme_classic(base_size = 12) +
  
  # Fine-tuning: make sure the axis line is aligned at position 2
  theme(
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "none",
    legend.title = element_blank(),
    axis.title = element_text(color = "black"),
    axis.text= element_text(color="black"), # tilt the labels to prevent overlap
    plot.margin = margin(10, 20, 10, 10) # slightly increase the margins so that labels are not cut off
  )
ggsave("outplots/fig1/fig1_c.ADI_14d.CRRs.counts.pdf",p3,width = 6,height = 2,dpi = 300)




## Draw the corresponding Venn diagram
library(GenomicRanges)

# For demonstration purposes, we create some simulated GRanges objects:
msc.k27.gr <- GRanges(
  seqnames = k27.env$msc.k27$chr,
  ranges = IRanges(start = k27.env$msc.k27$start, end = k27.env$msc.k27$end)
)

adi_7d.k27.gr <- GRanges(
  seqnames = k27.env$adi_7d.k27$chr,
  ranges = IRanges(start = k27.env$adi_7d.k27$start, end = k27.env$adi_7d.k27$end)
)

adi_14d.k27.gr <- GRanges(
  seqnames = k27.env$adi_15d.k27$chr,
  ranges = IRanges(start = k27.env$adi_15d.k27$start, end = k27.env$adi_15d.k27$end)
)


# ==========================================
# 2. Build the union master template (Master Regions)
# ==========================================
# Merge the coordinates of the three interval sets and reduce them to obtain the set of all non-overlapping intervals in which overlaps may occur
all_regions <- GenomicRanges::reduce(c(msc.k27.gr, adi_7d.k27.gr, adi_14d.k27.gr))

# ==========================================
# 3. Count the overlap of each interval across the three samples
# ==========================================
has_msc     <- GenomicRanges::countOverlaps(all_regions, msc.k27.gr) > 0
has_adi_7d  <- GenomicRanges::countOverlaps(all_regions, adi_7d.k27.gr) > 0
has_adi_14d <- GenomicRanges::countOverlaps(all_regions, adi_14d.k27.gr) > 0

# Combine the results into a Data Frame
df_regions <- data.frame(
  chr   = as.character(GenomicRanges::seqnames(all_regions)),
  start = GenomicRanges::start(all_regions),
  end   = GenomicRanges::end(all_regions),
  msc   = has_msc,
  adi_7d  = has_adi_7d,
  adi_14d = has_adi_14d,
  stringsAsFactors = FALSE
)


# ==========================================
# 4. Classify into 7 categories based on the overlap combination
# ==========================================
# Define a category label function
get_category <- function(m, a7, a14) {
  if (m & !a7 & !a14) return("1")       # only in msc.k27
  if (!m & a7 & !a14) return("2")       # only in adi_7d.k27
  if (!m & !a7 & a14) return("3")       # only in adi_14d.k27
  if (m & a7 & !a14)  return("12")      # shared by msc.k27 and adi_7d.k27
  if (m & !a7 & a14)  return("13")      # shared by msc.k27 and adi_14d.k27
  if (!m & a7 & a14)  return("23")      # shared by adi_7d.k27 and adi_14d.k27
  if (m & a7 & a14)   return("123")     # shared by all three
  return("none")                        # in theory this should never be none
}

# Apply the classification function
df_regions$category <- mapply(get_category, 
                              df_regions$msc, 
                              df_regions$adi_7d, 
                              df_regions$adi_14d)

# ==========================================
# 5. Extract the specific results for the 7 categories
# ==========================================
# We can save the 7 category results separately into a list
results_list <- split(df_regions[, c("chr", "start", "end")], df_regions$category)

# Print the number of intervals in each category
cat("Interval counts per category:\n")
print(table(df_regions$category))

# Examples of extracting specific category results:
# For example, to obtain the result of the three-way intersection (123):
intersection_123 <- results_list[["123"]]
print("Intervals shared by all three (123):")
print(intersection_123)

# For example, to obtain the result of msc.k27-only (1):
only_msc_1 <- results_list[["1"]]


# 2. Define the "disjoint slice" values for each part of the figure (Disjoint counts)
msc_only <- 5723
adi7d_only <- 8928
adi15d_only <- 5302

msc_adi7d_inter <- 3233   # MSC and ADI-7d intersection only
msc_adi15d_inter <- 792    # MSC and ADI-15d intersection only
adi7d_adi15d_inter <- 2790 # ADI-7d and ADI-15d intersection only

triple_inter <- 3212       # intersection of all three

# 3. Convert to the input format required by eulerr (computing the total intersection of each set)
# The eulerr logic is: A&B denotes the full intersection of A and B (including the three-way intersection part)
fit_data <- c(
  "MSC" = msc_only + msc_adi7d_inter + msc_adi15d_inter + triple_inter,
  "ADI-7d" = adi7d_only + msc_adi7d_inter + adi7d_adi15d_inter + triple_inter,
  "ADI-15d" = adi15d_only + msc_adi15d_inter + adi7d_adi15d_inter + triple_inter,
  "MSC&ADI-7d" = msc_adi7d_inter + triple_inter,
  "MSC&ADI-15d" = msc_adi15d_inter + triple_inter,
  "ADI-7d&ADI-15d" = adi7d_adi15d_inter + triple_inter,
  "MSC&ADI-7d&ADI-15d" = triple_inter
)
library(eulerr)
# 4. Fit the model
fit <- euler(fit_data)

# 5. Plot
p <- plot(fit,
     quantities = list(type = "counts", font = 2, cex = 0.8), # show the values; font=2 means bold
     fills = list(fill = c("#1F77B4", "#2CA02C" , "#D62728"), alpha = 0.7), # set the colors and transparency
     labels = list(col = "black", font = 2, cex = 1.2),      # set the label font
     edges = list(col = "white", lex = 2),                   # set the circle border lines
     main = "")                # title

ggsave(file = "outplots/fig1/fig1_d.venn.pdf",p,width=4,height=4,dpi=300)

```


# sup1.d Upset plot showing counts similar to those of a Venn diagram
```{r}
# 1. Install and load the package
# install.packages("UpSetR")
library(UpSetR)

# 2. Build the input vector from your data
# Note: the values in fromExpression represent the "exclusive" count of that specific combination
# For example, 'MSC' = 5723 means the count that occurs only in MSC
msc_only <- 5723
adi7d_only <- 8928
adi15d_only <- 5302

msc_adi7d_inter <- 3233   # MSC and ADI-7d intersection only
msc_adi15d_inter <- 792    # MSC and ADI-15d intersection only
adi7d_adi15d_inter <- 2790 # ADI-7d and ADI-15d intersection only

triple_inter <- 3212       # intersection of all three

# 3. Convert to the input format required by eulerr (computing the total intersection of each set)
# The eulerr logic is: A&B denotes the full intersection of A and B (including the three-way intersection part)
input_data <- c(
  "MSC" = msc_only + msc_adi7d_inter + msc_adi15d_inter + triple_inter,
  "ADI-7d" = adi7d_only + msc_adi7d_inter + adi7d_adi15d_inter + triple_inter,
  "ADI-14d" = adi15d_only + msc_adi15d_inter + adi7d_adi15d_inter + triple_inter,
  "MSC&ADI-7d" = msc_adi7d_inter + triple_inter,
  "MSC&ADI-14d" = msc_adi15d_inter + triple_inter,
  "ADI-7d&ADI-14d" = adi7d_adi15d_inter + triple_inter,
  "MSC&ADI-7d&ADI-14d" = triple_inter
)
target_order <- c("MSC", "ADI-7d", "ADI-14d")

# 3. Draw the UpSet plot

p <- upset(fromExpression(input_data), 
      sets = target_order,       # specify the sets to display and their order
      keep.order = TRUE,         # key parameter: set to TRUE to keep the order specified above
      nintersects = NA, 
      order.by = "freq",         # the top intersection bar chart is still sorted by frequency
      mb.ratio = c(0.7, 0.3),  # adjust the ratio between the top bar chart and the bottom matrix
      main.bar.color = "#696969", # main bar color
      sets.bar.color = c("#1F77B4", "#2CA02C" , "#D62728"), # color of the left set-size bar chart
      matrix.color = "black",     # color of the matrix points
      point.size = 3.5,           # size of the matrix points
      line.size = 1,              # thickness of the connecting lines
      text.scale = c(1.5, 1.2, 1.2, 1.2, 1.5, 1.5) # in order: y-axis label, y-axis ticks, x-axis label, x-axis ticks, set names, numbers above the bars
)
pdf(file = "outplots/fig1/fig1_d.periods.upset.plot.pdf",width=6,height=6)
p
dev.off()
```



# e. Main figure showing that the target gene expression of ROIs versus control at different time points is significantly different from that of the discarded peaks
```{r}
load("RNA-seq.Rdata")
load("gencode.v19.ref.Rdata")

# First produce the comparison plot of target genes among MRR, peak, ROIs and control
# fig1_d.targetGene.tpm.boxplot
msc.peaks.tg <- fread(file = "peaks/H3K27me3-0d.merged.q0.1_peaks.targetGene.tsv")
msc.peaks.tpm <- all.tpm[unique(msc.peaks.tg$target_gene_id),] %>% na.omit()
msc.MRR.tg <- fread(file = "peaks/ROSE/H3K27me3-0d_MRR.targetGene.tsv")
msc.MRR.tpm <- all.tpm[unique(unlist(strsplit(msc.MRR.tg$target_gene_id,",",fixed = T))),] %>% na.omit()
msc.k27.tg <- fread(file = "peaks/msc.k27/msc.k27.len7.bed.targetGene.tsv")
msc.k27.tpm <- all.tpm[unique(unlist(strsplit(msc.k27.tg$target_gene_id,",",fixed = T))),] %>% na.omit()
msc.k27.control.tg <- fread(file = "peaks/msc.k27/msc.k27.len7.control.bed.targetGene.tsv")
msc.k27.control.tpm <- all.tpm[unique(msc.k27.control.tg$target_gene_id),] %>% na.omit()

msc.k27.tpm.df <- data.frame(
    category=c(rep("ROIs (l >= 7)",nrow(msc.k27.tpm)),rep("H3K27me3 MRRs",nrow(msc.MRR.tpm)),rep("H3K27me3 peaks",nrow(msc.peaks.tpm)),rep("ROIs' Controls",nrow(msc.k27.control.tpm))),
    log2tpm=c(log2(msc.k27.tpm[,1]+1),log2(msc.MRR.tpm[,1]+1),log2(msc.peaks.tpm[,1]+1),log2(msc.k27.control.tpm[,1]+1))
)

# Make sure the order is: len < 7, len >= 7, control
msc.k27.tpm.df$category <- factor(msc.k27.tpm.df$category, 
                                  levels = c("ROIs (l >= 7)","H3K27me3 MRRs","H3K27me3 peaks","ROIs' Controls"))


p <- ggplot(msc.k27.tpm.df, aes(x = category, y = log2tpm, fill = category)) +
  # Add error bars; note that width controls the crossbar width and this layer must be placed before geom_boxplot
  stat_boxplot(geom = "errorbar", width = 0.4) +
  # Draw the boxplot
  geom_boxplot(outlier.shape = 21, width = 0.6,color="black") + # outlier.shape = NA hides the outlier points, depending on your needs
  scale_fill_manual(values = c("#1F77B4","#5FA2D1","#8EBFDD","#CCCCCC")) +
  # Visual polish
  theme_classic(base_size = 16) +
  labs(title = "",
       x = "",
       y = "log2(TPM)") +
  theme(
  legend.position = "none",
  axis.text = element_text(color="black"),
  panel.grid.minor = element_blank()
  ) # the legend can be hidden if it is not needed
ggsave(file = "outplots/fig1/fig1_e.tg.log2tpm.msc.pdf",p,width=6,height=4,dpi=300)

adi_7d.peaks.tg <- fread(file = "peaks/H3K27me3-0d.merged.q0.1_peaks.targetGene.tsv")
adi_7d.peaks.tpm <- all.tpm[unique(adi_7d.peaks.tg$target_gene_id),] %>% na.omit()
adi_7d.MRR.tg <- fread(file = "peaks/ROSE/H3K27me3-0d_MRR.targetGene.tsv")
adi_7d.MRR.tpm <- all.tpm[unique(unlist(strsplit(adi_7d.MRR.tg$target_gene_id,",",fixed = T))),] %>% na.omit()
adi_7d.k27.tg <- fread(file = "peaks/adi_7d.k27/adi_7d.k27.len7.bed.targetGene.tsv")
adi_7d.k27.tpm <- all.tpm[unique(unlist(strsplit(adi_7d.k27.tg$target_gene_id,",",fixed = T))),] %>% na.omit()
adi_7d.k27.control.tg <- fread(file = "peaks/adi_7d.k27/adi_7d.k27.len7.control.bed.targetGene.tsv")
adi_7d.k27.control.tpm <- all.tpm[unique(adi_7d.k27.control.tg$target_gene_id),] %>% na.omit()

adi_7d.k27.tpm.df <- data.frame(
    category=c(rep("ROIs (l >= 7)",nrow(adi_7d.k27.tpm)),rep("H3K27me3 MRRs",nrow(adi_7d.MRR.tpm)),rep("H3K27me3 peaks",nrow(adi_7d.peaks.tpm)),rep("ROIs' Controls",nrow(adi_7d.k27.control.tpm))),
    log2tpm=c(log2(adi_7d.k27.tpm[,1]+1),log2(adi_7d.MRR.tpm[,1]+1),log2(adi_7d.peaks.tpm[,1]+1),log2(adi_7d.k27.control.tpm[,1]+1))
)

# Make sure the order is: len < 7, len >= 7, control
adi_7d.k27.tpm.df$category <- factor(adi_7d.k27.tpm.df$category, 
                                  levels = c("ROIs (l >= 7)","H3K27me3 MRRs","H3K27me3 peaks","ROIs' Controls"))


p <- ggplot(adi_7d.k27.tpm.df, aes(x = category, y = log2tpm, fill = category)) +
  # Add error bars; note that width controls the crossbar width and this layer must be placed before geom_boxplot
  stat_boxplot(geom = "errorbar", width = 0.4) +
  # Draw the boxplot
  geom_boxplot(outlier.shape = 21, width = 0.6,color="black") + # outlier.shape = NA hides the outlier points, depending on your needs
  scale_fill_manual(values = c("#E69F00","#ECB240","#F2C57A","#CCCCCC")) +
  # Visual polish
  theme_classic(base_size = 16) +
  labs(title = "",
       x = "",
       y = "log2(TPM)") +
  theme(
  legend.position = "none",
  axis.text = element_text(color="black"),
  panel.grid.minor = element_blank()
  ) # the legend can be hidden if it is not needed
ggsave(file = "outplots/fig1/fig1_e.tg.log2tpm.adi_7d.pdf",p,width=6,height=4,dpi=300)



adi_14d.peaks.tg <- fread(file = "peaks/H3K27me3-0d.merged.q0.1_peaks.targetGene.tsv")
adi_14d.peaks.tpm <- all.tpm[unique(adi_14d.peaks.tg$target_gene_id),] %>% na.omit()
adi_14d.MRR.tg <- fread(file = "peaks/ROSE/H3K27me3-0d_MRR.targetGene.tsv")
adi_14d.MRR.tpm <- all.tpm[unique(unlist(strsplit(adi_14d.MRR.tg$target_gene_id,",",fixed = T))),] %>% na.omit()
adi_14d.k27.tg <- fread(file = "peaks/adi_14d.k27/adi_14d.k27.len7.bed.targetGene.tsv")
adi_14d.k27.tpm <- all.tpm[unique(unlist(strsplit(adi_14d.k27.tg$target_gene_id,",",fixed = T))),] %>% na.omit()
adi_14d.k27.control.tg <- fread(file = "peaks/adi_14d.k27/adi_14d.k27.len7.control.bed.targetGene.tsv")
adi_14d.k27.control.tpm <- all.tpm[unique(adi_14d.k27.control.tg$target_gene_id),] %>% na.omit()

adi_14d.k27.tpm.df <- data.frame(
    category=c(rep("ROIs (l >= 7)",nrow(adi_14d.k27.tpm)),rep("H3K27me3 MRRs",nrow(adi_14d.MRR.tpm)),rep("H3K27me3 peaks",nrow(adi_14d.peaks.tpm)),rep("ROIs' Controls",nrow(adi_14d.k27.control.tpm))),
    log2tpm=c(log2(adi_14d.k27.tpm[,1]+1),log2(adi_14d.MRR.tpm[,1]+1),log2(adi_14d.peaks.tpm[,1]+1),log2(adi_14d.k27.control.tpm[,1]+1))
)

# Make sure the order is: len < 7, len >= 7, control
adi_14d.k27.tpm.df$category <- factor(adi_14d.k27.tpm.df$category, 
                                  levels = c("ROIs (l >= 7)","H3K27me3 MRRs","H3K27me3 peaks","ROIs' Controls"))


p <- ggplot(adi_14d.k27.tpm.df, aes(x = category, y = log2tpm, fill = category)) +
  # Add error bars; note that width controls the crossbar width and this layer must be placed before geom_boxplot
  stat_boxplot(geom = "errorbar", width = 0.4) +
  # Draw the boxplot
  geom_boxplot(outlier.shape = 21, width = 0.6,color="black") + # outlier.shape = NA hides the outlier points, depending on your needs
  scale_fill_manual(values = c("#CC79A7","#D695BC","#E0B0D2","#CCCCCC")) +
  # Visual polish
  theme_classic(base_size = 16) +
  labs(title = "",
       x = "",
       y = "log2(TPM)") +
  theme(
  legend.position = "none",
  axis.text = element_text(color="black"),
  panel.grid.minor = element_blank()
  ) # the legend can be hidden if it is not needed
ggsave(file = "outplots/fig1/fig1_e.tg.log2tpm.adi_14d.pdf",p,width=6,height=4,dpi=300)


```


# f. Overlap with other histone modification signals
## 1. Proportion of covered signal (coverage); 
## 2. Enrichment level of each chip, compared with the background
```{r}

l1 <- bt.intersect(a=msc.all,b=k27.env$msc.k27,wa=T)
l2 <- bt.intersect(a=adi_7d.all,b=k27.env$adi_7d.k27,wa=T)
l3 <- bt.intersect(a=adi_15d.all,b=k27.env$adi_15d.k27,wa=T)

df <- data.frame(
    msc=c(table(l1$V4)[2]/nrow(l1),table(l1$V5)[2]/nrow(l1),table(l1$V7)[2]/nrow(l1),table(l1$V8)[2]/nrow(l1),table(l1$V9)[2]/nrow(l1),table(l1$V10)[2]/nrow(l1),table(l1$V11)[2]/nrow(l1)),
    adi_7d = c(table(l2$V4)[2]/nrow(l2),table(l2$V5)[2]/nrow(l2),table(l2$V7)[2]/nrow(l2),table(l2$V8)[2]/nrow(l2),table(l2$V9)[2]/nrow(l2),table(l2$V10)[2]/nrow(l2),table(l2$V11)[2]/nrow(l2)),
    adi_14d=c(table(l3$V4)[2]/nrow(l3),table(l3$V5)[2]/nrow(l3),table(l3$V7)[2]/nrow(l3),table(l3$V8)[2]/nrow(l3),table(l3$V9)[2]/nrow(l3),table(l3$V10)[2]/nrow(l3),table(l3$V11)[2]/nrow(l3))
)
rownames(df) <- colnames(msc.all)[c(4,5,7:11)]

row_order <- rownames(df)
col_order <- colnames(df)

# 4. Convert the row names into a column and reshape the data into long format
df_long <- df %>%
  mutate(Feature = rownames(df)) %>%
  pivot_longer(cols = -Feature, names_to = "Condition", values_to = "Value")

# 5. Lock the row/column order: convert Feature and Condition to factors and specify levels
# Note: ggplot2 draws the y axis from bottom to top by default; to keep the first row "DNase" at the top, we need to reverse the row order (rev)
df_long$Feature <- factor(df_long$Feature, levels = rev(row_order))
df_long$Condition <- factor(df_long$Condition, levels = col_order)

# 6. Start drawing the heatmap
p <- ggplot(df_long, aes(x = Condition, y = Feature, fill = Value)) +
  geom_tile(color = "white", size = 0.5) +  # set the cell borders to white for a more refined look
  # Set the gradient: 0 is white and high values transition toward warm colors (red); use limits = c(0, NA) to ensure that 0 maps to white
  scale_fill_gradientn(
    colors = c("#ffffff", "#ffebae", "#ff9e3d", "#d7191c"), 
    values = c(0, 0.1, 0.5, 1.0), # weights that control the color transition
    limits = c(0, max(df_long$Value)),
    name = "Value"
  ) +
  # Adjust the polished theme
  theme_minimal(base_size = 16) +
  labs(x = NULL, y = NULL) + # remove the redundant axis labels
  theme(
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    panel.grid = element_blank(), # remove the grid lines
    legend.position = "right"     # place the legend on the right
  ) +
  # Add numeric labels in the cells (optional; delete the geom_text line below if not needed)
  geom_text(aes(label = sprintf("%.3f", Value)), color = "black", size = 4)
ggsave(p,file = "outplots/fig1/fig1_g.other.chip.coverage.pdf",width=6,height=7)


k27.env$msc.k27.control <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig1/msc.k27.control.bed",header = F)
k27.env$adi_7d.k27.control <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig1/adi_7d.k27.control.bed",header = F)
k27.env$adi_15d.k27.control <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig1/adi_14d.k27.control.bed",header = F)

make_gr <- function(df) {
  GRanges(seqnames = df[,1], ranges = IRanges(start = df[,2], end = df[,3]))
}

msc.otherchip.out <- data.frame(matrix(nrow=8,ncol=4))
rownames(msc.otherchip.out) <- mark_names
colnames(msc.otherchip.out) <- c("lower","or","upper","sig")

for (i in 1:length(mark_names)){
    # 1. Compute the overall c(a1,a2) and c(b1,b2)
    k4_pos_bins <- msc.all[msc.all[,i+3] == 1, 1:3]
    gr_k4_active <- make_gr(k4_pos_bins)
    gr_roi <- make_gr(k27.env$msc.k27)
    gr_control <- make_gr(k27.env$msc.k27.control)
    roi_densities <- data.frame(a1=countOverlaps(gr_roi, gr_k4_active),a2=round(width(gr_roi)/200)-countOverlaps(gr_roi, gr_k4_active))
    control_densities <- data.frame(b1=countOverlaps(gr_control, gr_k4_active),b2=round(width(gr_control)/200)-countOverlaps(gr_control, gr_k4_active))

    # 2. Loop 1000 times to compute the OR in each iteration
    set.seed(42) # ensure the results are reproducible
    n_iterations <- 1000
    sample_size <- 1000
    results <- replicate(n_iterations, {
  # Random sampling
        s_roi <- roi_densities[sample(nrow(roi_densities), sample_size),]
        s_con <- control_densities[sample(nrow(control_densities), sample_size*2),]
        out <- fisher.test(matrix(c(a1=sum(s_roi$a1),a2=sum(s_roi$a2),b1=sum(s_con$b1),b2=sum(s_con$b2)),nrow=2,ncol=2))$estimate
        
        return(as.numeric(out))
    })
    # 3. Compute the final OR and confidence interval
    output <- c(quantile(results,0.05),quantile(results,0.5),quantile(results,0.95)) %>% as.numeric
    issig <- ifelse(as.numeric((quantile(results,0.05)-1)*(quantile(results,0.95)-1)) >0 ,"sig","unsig")
    msc.otherchip.out[i,] <- c(output,issig)

}

adi_7d.otherchip.out <- data.frame(matrix(nrow=8,ncol=4))
rownames(adi_7d.otherchip.out) <- mark_names
colnames(adi_7d.otherchip.out) <- c("lower","or","upper","sig")

for (i in 1:length(mark_names)){
    # 1. Compute the overall c(a1,a2) and c(b1,b2)
    k4_pos_bins <- adi_7d.all[adi_7d.all[,i+3] == 1, 1:3]
    gr_k4_active <- make_gr(k4_pos_bins)
    gr_roi <- make_gr(k27.env$adi_7d.k27)
    gr_control <- make_gr(k27.env$adi_7d.k27.control)
    roi_densities <- data.frame(a1=countOverlaps(gr_roi, gr_k4_active),a2=round(width(gr_roi)/200)-countOverlaps(gr_roi, gr_k4_active))
    control_densities <- data.frame(b1=countOverlaps(gr_control, gr_k4_active),b2=round(width(gr_control)/200)-countOverlaps(gr_control, gr_k4_active))

    # 2. Loop 1000 times to compute the OR in each iteration
    set.seed(42) # ensure the results are reproducible
    n_iterations <- 1000
    sample_size <- 1000
    results <- replicate(n_iterations, {
  # Random sampling
        s_roi <- roi_densities[sample(nrow(roi_densities), sample_size),]
        s_con <- control_densities[sample(nrow(control_densities), sample_size*2),]
        out <- fisher.test(matrix(c(a1=sum(s_roi$a1),a2=sum(s_roi$a2),b1=sum(s_con$b1),b2=sum(s_con$b2)),nrow=2,ncol=2))$estimate
        
        return(as.numeric(out))
    })
    # 3. Compute the final OR and confidence interval
    output <- c(quantile(results,0.05),quantile(results,0.5),quantile(results,0.95)) %>% as.numeric
    issig <- ifelse(as.numeric((quantile(results,0.05)-1)*(quantile(results,0.95)-1)) >0 ,"sig","unsig")
    adi_7d.otherchip.out[i,] <- c(output,issig)

}


adi_15d.otherchip.out <- data.frame(matrix(nrow=8,ncol=4))
rownames(adi_15d.otherchip.out) <- mark_names
colnames(adi_15d.otherchip.out) <- c("lower","or","upper","sig")

for (i in 1:length(mark_names)){
    # 1. Compute the overall c(a1,a2) and c(b1,b2)
    k4_pos_bins <- adi_15d.all[adi_15d.all[,i+3] == 1, 1:3]
    gr_k4_active <- make_gr(k4_pos_bins)
    gr_roi <- make_gr(k27.env$adi_15d.k27)
    gr_control <- make_gr(k27.env$adi_15d.k27.control)
    roi_densities <- data.frame(a1=countOverlaps(gr_roi, gr_k4_active),a2=round(width(gr_roi)/200)-countOverlaps(gr_roi, gr_k4_active))
    control_densities <- data.frame(b1=countOverlaps(gr_control, gr_k4_active),b2=round(width(gr_control)/200)-countOverlaps(gr_control, gr_k4_active))

    # 2. Loop 1000 times to compute the OR in each iteration
    set.seed(42) # ensure the results are reproducible
    n_iterations <- 1000
    sample_size <- 1000
    results <- replicate(n_iterations, {
  # Random sampling
        s_roi <- roi_densities[sample(nrow(roi_densities), sample_size),]
        s_con <- control_densities[sample(nrow(control_densities), sample_size*2),]
        out <- fisher.test(matrix(c(a1=sum(s_roi$a1),a2=sum(s_roi$a2),b1=sum(s_con$b1),b2=sum(s_con$b2)),nrow=2,ncol=2))$estimate
        
        return(as.numeric(out))
    })
    # 3. Compute the final OR and confidence interval
    output <- c(quantile(results,0.05),quantile(results,0.5),quantile(results,0.95)) %>% as.numeric
    issig <- ifelse(as.numeric((quantile(results,0.05)-1)*(quantile(results,0.95)-1)) >0 ,"sig","unsig")
    adi_15d.otherchip.out[i,] <- c(output,issig)

}

msc.otherchip.out <- msc.otherchip.out[-3,]
adi_7d.otherchip.out <- adi_7d.otherchip.out[-3,]
adi_15d.otherchip.out <- adi_15d.otherchip.out[-3,]


df <- data.frame(
    category=c(rownames(msc.otherchip.out),rownames(adi_7d.otherchip.out),rownames(adi_15d.otherchip.out)),
    time=c(rep("MSC",7),rep("ADI_7d",7),rep("ADI_14d",7)),
    OR=c(msc.otherchip.out$or,adi_7d.otherchip.out$or,adi_15d.otherchip.out$or),
    lower=c(msc.otherchip.out$lower,adi_7d.otherchip.out$lower,adi_15d.otherchip.out$lower),
    upper=c(msc.otherchip.out$upper,adi_7d.otherchip.out$upper,adi_15d.otherchip.out$upper),
    sig=c(msc.otherchip.out$sig,adi_7d.otherchip.out$sig,adi_15d.otherchip.out$sig)
)
df$OR <- as.numeric(df$OR)
df$lower <- as.numeric(df$lower)
df$upper <- as.numeric(df$upper)
df$category <- factor(df$category,mark_names)
df$time <- factor(df$time,c("MSC","ADI_7d","ADI_14d"))

library(dplyr)
library(ggplot2)
library(scales)

# df: category, time, OR, lower, upper, FDR

# ---- 1) Preprocessing: significance/direction/intensity ----
df2 <- df %>%
  mutate(
    # Convert to Beta values (log2 transformation)
    beta = log2(OR),
    beta_lower = log2(lower),
    beta_upper = log2(upper),
    
    # Color assignment logic
    # sig == "sig" and OR > 1 -> red
    # sig == "sig" and OR < 1 -> blue
    # sig == "unsig" -> gray
    color_group = case_when(
      sig == "unsig" ~ "Non-Significant",
      sig == "sig" & OR > 1 ~ "Risk (Increased)",
      sig == "sig" & OR < 1 ~ "Protective (Decreased)",
      TRUE ~ "Non-Significant"
    )
  )

# Factor ordering (in your input data ADI_15d is stored as ADI_14d, please check)
time_levels <- c("MSC", "ADI_7d", "ADI_14d") 
df2$time <- factor(df2$time, levels = time_levels)

# Category ordering (if mark_names is undefined, default to alphabetical order or specify manually)
# mark_names <- c("DNase", "H3K27ac", "H3K36me3", "H3K4me1", "H3K4me3", "H3K9ac", "H3K9me3")
# df2$category <- factor(df2$category, levels = mark_names)

# 2. Define colors
# Chose a deeper red and a deeper blue for a more academic look
my_colors <- c(
  "Risk (Increased)" = "#a50f15",      # deep red
  "Protective (Decreased)" = "#08519c", # deep blue
  "Non-Significant" = "#BDBDBD"         # gray
)

# 3. Plot
dodge <- position_dodge(width = 0.7)

p <- ggplot(df2, aes(x = category, y = beta, group = time, color = color_group)) +
  # Reference line: Beta = 0 (corresponding to OR = 1)
  geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.6, color = "grey30") +
  
  # Error bars
  geom_errorbar(
    aes(ymin = beta_lower, ymax = beta_upper),
    position = dodge,
    width = 0.2,
    linewidth = 0.8
  ) +
  
  # Points
  geom_point(
    aes(shape = time),
    position = dodge,
    size = 3.5,
    stroke = 1,
    fill = "white" # to make the points hollow, change shape to 21-25
  ) +
  
  # Color mapping
  scale_color_manual(values = my_colors,guide = "none") +
  
  # Shape mapping (corresponding to the three time points in your data)
  scale_shape_manual(values = c("MSC" = 16, "ADI_7d" = 17, "ADI_14d" = 15)) +
  
  # Labels
  labs(
    y = "Beta Estimate",
    x = NULL,
    shape = "Time"
  ) +
  
  # Theme
  theme_classic(base_size = 16) +
  theme(
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "grey95"), # add faint horizontal grid lines to aid alignment
    axis.text.x = element_text(color = "black"), # rotate the x-axis labels to prevent overlap
    axis.text.y = element_text(color = "black"),
    legend.position = "right"
  )

# 4. Save
ggsave(file = "outplots/fig1/fig1_f.other.chip.bivalent.pdf", p, width = 10, height = 7, dpi = 300)


```


# sup1.a Show the fraction of signal occupancy of all histone modifications
```{r}
all.coverage.df <- data.frame(matrix(nrow=8,ncol=3))
rownames(all.coverage.df) <- mark_names
colnames(all.coverage.df) <- c("MSC","ADI_7d","ADI_14d")
for (i in 1:length(mark_names)){
    out1 <- table(msc.all[,mark_names[i]])[2]/length(msc.all[,mark_names[i]]) %>% as.numeric()
    out2 <- table(adi_7d.all[,mark_names[i]])[2]/length(adi_7d.all[,mark_names[i]]) %>% as.numeric()
    out3 <- table(adi_15d.all[,mark_names[i]])[2]/length(adi_15d.all[,mark_names[i]]) %>% as.numeric()
    all.coverage.df[i,] <- c(out1,out2,out3)
}


# Data transformation: from wide format to long format
# Convert the row names into a column and flatten the data with pivot_longer
df_long <- all.coverage.df %>%
  tibble::rownames_to_column(var = "Marker") %>%
  pivot_longer(cols = -Marker, names_to = "Sample", values_to = "Coverage")

# 4. Lock the sample order (so that the colors map to MSC, ADI_7d, ADI_14d in order)
df_long$Sample <- factor(df_long$Sample, levels = c("MSC", "ADI_7d", "ADI_14d"))

# 5. Plot
p <- ggplot(df_long, aes(x = Marker, y = Coverage, fill = Sample)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  scale_fill_manual(values = c("MSC" = "#0072B2", 
                               "ADI_7d" = "#E69F00", 
                               "ADI_14d" = "#CC79A7")) +
  theme_bw(base_size = 16) +
  labs(title = "",
       x = "",
       y = "Coverage Value",
       fill = "Sample Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
  axis.text=element_text(color="black"),
  legend.position="top") # if the labels are too crowded, rotate them by 45 degrees
ggsave(file = "outplots/fig1/sup/sup1_a.all.chip.coverage.pdf", p, width = 12, height = 5, dpi = 300)

```

<!-- 
# sup1.b Show the length and count distribution of the k27.all data
```{r}
hist(k27.env$msc.k27.all$len)
library(ggplot2)

max_y <- 170000
max_x <- 8
p1 <- ggplot(data.frame(len = k27.env$msc.k27.all$len), aes(x = log2(len))) +
  # Draw the histogram: set the fill color, border color and transparency
  geom_histogram(binwidth = 0.6, boundary = 0, fill = "#1F77B4", color = "black") +
  scale_y_continuous(limits = c(0, max_y),expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(limits = c(0, max_x),expand = c(-0.1, 0.1)) +
  # Use a clean theme
  theme_classic(base_size = 16) +
  # Add a title and modify the axis labels
  labs(
    title = ,
    subtitle = ,
    x = expression(log[2](Length)),
    y = ""
  ) +
  # Further fine-tune the details
  theme(
    axis.text = element_text(color="black")
)

p2 <- ggplot(data.frame(len = k27.env$adi_7d.k27.all$len), aes(x = log2(len))) +
  # Draw the histogram: set the fill color, border color and transparency
  geom_histogram(binwidth = 0.6, boundary = 0, fill = "#2CA02C", color = "black") +
  scale_y_continuous(limits = c(0, max_y),expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(limits = c(0, max_x),expand = c(-0.1, 0.1)) +
  # Use a clean theme
  theme_classic(base_size = 16) +
  # Add a title and modify the axis labels
  labs(
    title = ,
    subtitle = ,
    x = expression(log[2](Length)),
    y = ""
  ) +
  # Further fine-tune the details
  theme(
    axis.text = element_text(color="black")
)

p3 <- ggplot(data.frame(len = k27.env$adi_15d.k27.all$len), aes(x = log2(len))) +
  # Draw the histogram: set the fill color, border color and transparency
  geom_histogram(binwidth = 0.6, boundary = 0, fill = "#D62728", color = "black") +
  scale_y_continuous(limits = c(0, max_y),expand = expansion(mult = c(0, 0.1))) +
  scale_x_continuous(limits = c(0, max_x),expand = c(-0.1, 0.1)) +
  # Use a clean theme
  theme_classic(base_size = 16) +
  # Add a title and modify the axis labels
  labs(
    title = ,
    subtitle = ,
    x = expression(log[2](Length)),
    y = ""
  ) +
  # Further fine-tune the details
  theme(
    axis.text = element_text(color="black")
)
library(patchwork)
combined_plot <- p1 + p2  + p3
ggsave(file = "outplots/fig1/sup/sup1_b.all.k27.hist.pdf",combined_plot,width=11,height=4,dpi=300)

``` -->


# sup1.c Show the relationship between ROIs of different lengths and the count and expression of their target genes
```{r}
k27.env$msc.k27.all.pro.tg <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig1/msc.k27.all.tg.bed.tsv",header = T)
k27.env$msc.k27.all.pro.tg <- k27.env$msc.k27.all.pro.tg[k27.env$msc.k27.all.pro.tg$final_type != "intergenic_far",]

k27.env$adi_7d.k27.all.pro.tg <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig1/adi_7d.k27.all.tg.bed.tsv",header = T)
k27.env$adi_7d.k27.all.pro.tg <- k27.env$adi_7d.k27.all.pro.tg[k27.env$adi_7d.k27.all.pro.tg$final_type != "intergenic_far",]

k27.env$adi_15d.k27.all.pro.tg <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig1/adi_14d.k27.all.tg.bed.tsv",header = T)
k27.env$adi_15d.k27.all.pro.tg <- k27.env$adi_15d.k27.all.pro.tg[k27.env$adi_15d.k27.all.pro.tg$final_type != "intergenic_far",]


outtpm.list <- list()
for (i in 2:15){
    tg <- k27.env$msc.k27.all.pro.tg[k27.env$msc.k27.all.pro.tg$len == i,"target_gene_id"]
    outtpm.list[[i]] <- log2(all.tpm[sample(tg,1000),"MSC"]+1)
}

boxplot(outtpm.list[[2]],outtpm.list[[3]],outtpm.list[[4]],outtpm.list[[5]],outtpm.list[[6]],outtpm.list[[7]],outtpm.list[[8]],outtpm.list[[9]],outtpm.list[[10]])


## Do a simple sensitivity analysis: target gene expression for l>=3, l>=5, l>=7, l>=9

outtpm.list <- list()
for (i in c(3,5,7,9,15)){
    tg <- k27.env$msc.k27.all.pro.tg[k27.env$msc.k27.all.pro.tg$len >= i,"target_gene_id"]
    outtpm.list[[i]] <- log2(all.tpm[tg,"MSC"]+1)
}

boxplot(outtpm.list[[3]],outtpm.list[[5]],outtpm.list[[7]],outtpm.list[[9]],outtpm.list[[15]])

msc.allroi.tpm <- data.frame(
    category=c(rep("l >= 3",length(outtpm.list[[3]])),rep("l >= 5",length(outtpm.list[[5]])),rep("l >= 7",length(outtpm.list[[7]])),rep("l >= 9",length(outtpm.list[[9]])),rep("l >= 15",length(outtpm.list[[15]]))),
    log2tpm=c(outtpm.list[[3]],outtpm.list[[5]],outtpm.list[[7]],outtpm.list[[9]],outtpm.list[[15]]))



# 1. Prepare the data (assuming you already have msc.allroi.tpm)
# For demonstration purposes, simulated data are generated here
set.seed(123)
categories_order <- c("l >= 3", "l >= 5", "l >= 7", "l >= 9", "l >= 15")
counts <- c(nrow(k27.env$msc.k27.all[k27.env$msc.k27.all$len >= 3,]),nrow(k27.env$msc.k27.all[k27.env$msc.k27.all$len >= 5,]),nrow(k27.env$msc.k27.all[k27.env$msc.k27.all$len >= 7,]),nrow(k27.env$msc.k27.all[k27.env$msc.k27.all$len >= 9,]),nrow(k27.env$msc.k27.all[k27.env$msc.k27.all$len >= 15,]))

# Cumulative frequency data
cum_data <- data.frame(
  category = factor(categories_order, levels = categories_order),
  cum_val = counts
)
msc.allroi.tpm$category <- factor(msc.allroi.tpm$category,levels=categories_order)

# 2. Draw the upper panel: bar chart (cumulative frequency)
# Requirement: draw it upside down, with the x axis at the top
p1 <- ggplot(cum_data, aes(x = category, y = cum_val)) +
  geom_col(fill = "#69b3a2", width = 0.6, alpha = 0.8) +
  # Use scale_y_reverse so that the bars grow downward
  scale_y_reverse(
    #labels = scales::percent_format(), 
    expand = expansion(mult = c(0.1, 0))) +
  # Move the x axis to the top
  scale_x_discrete(position = "top") +
  labs(y = "Cumulative Frequency", x = NULL) +
  theme_bw(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text = element_text(color="black"),
    panel.grid.minor = element_blank()
  )



# 3. Draw the lower panel: boxplot (gene expression)
p2 <- ggplot(msc.allroi.tpm, aes(x = category, y = log2tpm)) +
  # coef = 0 keeps only the box (25%-75%) and the median line, removing the whiskers
  # outlier.shape = NA removes all outlier points
  stat_boxplot(geom = "errorbar",
        width=0.4)+
  geom_boxplot(aes(fill = category), coef = 0, outlier.shape = NA, alpha = 1, width = 0.6) +
  scale_fill_manual(
    values = c("#D6E8F9","#A0CCE3","#5FA3D0","#1F77B4","#0E4C8C")
  ) +
  labs(y = "Log2(TPM) Expression", x = "Category") +
  theme_bw(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text = element_text(color="black"),
    panel.grid.minor = element_blank()
  )

# 4. Combine the two panels with patchwork
# plot_layout(heights = c(1, 2)) sets the height ratio of the upper and lower panels
combined_plot <- p1 / p2 + 
  patchwork::plot_layout(heights = c(1, 1.5)) 

# Show the plot
print(combined_plot)

ggsave("outplots/fig1/sup/sup1_c.MSC.compare.pdf",combined_plot,width=5,height=6,dpi=300)


## ADI 7d
outtpm.list <- list()
for (i in c(3,5,7,9,17)){
    tg <- k27.env$adi_7d.k27.all.pro.tg[k27.env$adi_7d.k27.all.pro.tg$len >= i,"target_gene_id"]
    outtpm.list[[i]] <- log2(all.tpm[tg,"MSC"]+1)
}

boxplot(outtpm.list[[3]],outtpm.list[[5]],outtpm.list[[7]],outtpm.list[[9]],outtpm.list[[17]])

msc.allroi.tpm <- data.frame(
    category=c(rep("l >= 3",length(outtpm.list[[3]])),rep("l >= 5",length(outtpm.list[[5]])),rep("l >= 7",length(outtpm.list[[7]])),rep("l >= 9",length(outtpm.list[[9]])),rep("l >= 17",length(outtpm.list[[17]]))),
    log2tpm=c(outtpm.list[[3]],outtpm.list[[5]],outtpm.list[[7]],outtpm.list[[9]],outtpm.list[[17]]))



# 1. Prepare the data (assuming you already have msc.allroi.tpm)
# For demonstration purposes, simulated data are generated here
set.seed(123)
categories_order <- c("l >= 3", "l >= 5", "l >= 7", "l >= 9", "l >= 17")
counts <- c(nrow(k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >= 3,]),nrow(k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >= 5,]),nrow(k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >= 7,]),nrow(k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >= 9,]),nrow(k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >= 17,]))

# Cumulative frequency data
cum_data <- data.frame(
  category = factor(categories_order, levels = categories_order),
  cum_val = counts
)
msc.allroi.tpm$category <- factor(msc.allroi.tpm$category,levels=categories_order)

# 2. Draw the upper panel: bar chart (cumulative frequency)
# Requirement: draw it upside down, with the x axis at the top
p1 <- ggplot(cum_data, aes(x = category, y = cum_val)) +
  geom_col(fill = "#69b3a2", width = 0.6, alpha = 0.8) +
  # Use scale_y_reverse so that the bars grow downward
  scale_y_reverse(
    #labels = scales::percent_format(), 
    expand = expansion(mult = c(0.1, 0))) +
  # Move the x axis to the top
  scale_x_discrete(position = "top") +
  labs(y = "Cumulative Frequency", x = NULL) +
  theme_bw(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text = element_text(color="black"),
    panel.grid.minor = element_blank()
  )



# 3. Draw the lower panel: boxplot (gene expression)
p2 <- ggplot(msc.allroi.tpm, aes(x = category, y = log2tpm)) +
  # coef = 0 keeps only the box (25%-75%) and the median line, removing the whiskers
  # outlier.shape = NA removes all outlier points
  stat_boxplot(geom = "errorbar",
        width=0.4)+
  geom_boxplot(aes(fill = category), coef = 0, outlier.shape = NA, alpha = 1, width = 0.6) +
  scale_fill_manual(
    values = c("#E6F4E6", "#B8E0B8", "#7FC97F", "#2CA02C", "#1A6E1A")
  ) +
  labs(y = "Log2(TPM) Expression", x = "Category") +
  theme_bw(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text = element_text(color="black"),
    panel.grid.minor = element_blank()
  )

# 4. Combine the two panels with patchwork
# plot_layout(heights = c(1, 2)) sets the height ratio of the upper and lower panels
combined_plot <- p1 / p2 + 
  patchwork::plot_layout(heights = c(1, 1.5)) 

# Show the plot
print(combined_plot)

ggsave("outplots/fig1/sup/sup1_c.ADI_7d.compare.pdf",combined_plot,width=5,height=6,dpi=300)


## ADI 14d
outtpm.list <- list()
for (i in c(3,5,7,9,15)){
    tg <- k27.env$adi_15d.k27.all.pro.tg[k27.env$adi_15d.k27.all.pro.tg$len >= i,"target_gene_id"]
    outtpm.list[[i]] <- log2(all.tpm[tg,"MSC"]+1)
}

boxplot(outtpm.list[[3]],outtpm.list[[5]],outtpm.list[[7]],outtpm.list[[9]],outtpm.list[[15]])

msc.allroi.tpm <- data.frame(
    category=c(rep("l >= 3",length(outtpm.list[[3]])),rep("l >= 5",length(outtpm.list[[5]])),rep("l >= 7",length(outtpm.list[[7]])),rep("l >= 9",length(outtpm.list[[9]])),rep("l >= 15",length(outtpm.list[[15]]))),
    log2tpm=c(outtpm.list[[3]],outtpm.list[[5]],outtpm.list[[7]],outtpm.list[[9]],outtpm.list[[15]]))



# 1. Prepare the data (assuming you already have msc.allroi.tpm)
# For demonstration purposes, simulated data are generated here
set.seed(123)
categories_order <- c("l >= 3", "l >= 5", "l >= 7", "l >= 9", "l >= 15")
counts <- c(nrow(k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >= 3,]),nrow(k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >= 5,]),nrow(k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >= 7,]),nrow(k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >= 9,]),nrow(k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >= 15,]))

# Cumulative frequency data
cum_data <- data.frame(
  category = factor(categories_order, levels = categories_order),
  cum_val = counts
)
msc.allroi.tpm$category <- factor(msc.allroi.tpm$category,levels=categories_order)

# 2. Draw the upper panel: bar chart (cumulative frequency)
# Requirement: draw it upside down, with the x axis at the top
p1 <- ggplot(cum_data, aes(x = category, y = cum_val)) +
  geom_col(fill = "#69b3a2", width = 0.6, alpha = 0.8) +
  # Use scale_y_reverse so that the bars grow downward
  scale_y_reverse(
    #labels = scales::percent_format(), 
    expand = expansion(mult = c(0.1, 0))) +
  # Move the x axis to the top
  scale_x_discrete(position = "top") +
  labs(y = "Cumulative Frequency", x = NULL) +
  theme_bw(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text = element_text(color="black"),
    panel.grid.minor = element_blank()
  )



# 3. Draw the lower panel: boxplot (gene expression)
p2 <- ggplot(msc.allroi.tpm, aes(x = category, y = log2tpm)) +
  # coef = 0 keeps only the box (25%-75%) and the median line, removing the whiskers
  # outlier.shape = NA removes all outlier points
  stat_boxplot(geom = "errorbar",
        width=0.4)+
  geom_boxplot(aes(fill = category), coef = 0, outlier.shape = NA, alpha = 1, width = 0.6) +
  scale_fill_manual(
    values = c("#FAD2D2", "#F6A6A6", "#EF6A6A", "#D93A3A", "#A80F0F")
  ) +
  labs(y = "Log2(TPM) Expression", x = "Category") +
  theme_bw(base_size = 16) +
  theme(
    legend.position = "none",
    axis.text = element_text(color="black"),
    panel.grid.minor = element_blank()
  )

# 4. Combine the two panels with patchwork
# plot_layout(heights = c(1, 2)) sets the height ratio of the upper and lower panels
combined_plot <- p1 / p2 + 
  patchwork::plot_layout(heights = c(1, 1.5)) 

# Show the plot
print(combined_plot)

ggsave("outplots/fig1/sup/sup1_c.ADI_14d.compare.pdf",combined_plot,width=5,height=6,dpi=300)



```
<!-- 
# sup1.e Alluvial Sankey diagram showing the transitions
```{r}

## When looking at l>=7
l1 <- data.frame(msc.all,index=1:nrow(msc.all))
l1 <- l1[l1$H3K27me3 == 1,c(1,2,3,12)]
l1.k27 <- bt.intersect(a=l1,b=k27.env$msc.k27,wa=T) %>% unique()

l2 <- data.frame(adi_7d.all,index=1:nrow(adi_7d.all))
l2 <- l2[l2$H3K27me3 == 1,c(1,2,3,12)]
l2.k27 <- bt.intersect(a=l2,b=k27.env$adi.7d.k27,wa=T) %>% unique()

l3 <- data.frame(adi_15d.all,index=1:nrow(adi_15d.all))
l3 <- l3[l3$H3K27me3 == 1,c(1,2,3,12)]
l3.k27 <- bt.intersect(a=l3,b=k27.env$adi.15d.k27,wa=T) %>% unique()

l1 <- l1.k27[,4]
l2 <- l2.k27[,4]
l3 <- l3.k27[,4]

all <- unique(c(l1,l2,l3))
l1.out <- setdiff(all,l1)
l2.out <- setdiff(all,l2)
l3.out <- setdiff(all,l3)

intersect(intersect(l1.out,l2.out),l3.out) %>% unique %>% length
intersect(intersect(l1,l2.out),l3.out) %>% unique %>% length
intersect(intersect(l1.out,l2),l3.out) %>% unique %>% length
intersect(intersect(l1,l2),l3.out) %>% unique %>% length
intersect(intersect(l1.out,l2.out),l3) %>% unique %>% length
intersect(intersect(l1,l2.out),l3) %>% unique %>% length
intersect(intersect(l1.out,l2),l3) %>% unique %>% length
intersect(intersect(l1,l2),l3) %>% unique %>% length


# Load the required libraries
library(ggplot2)
library(ggalluvial)

# Build all paths and their flows (example data)
# Each row represents one complete path: T1 -> T2 -> T3
paths <- expand.grid(
  MSC = c(0, 1),
  ADI_7d = c(0, 1),
  ADI_14d = c(0, 1)
)

paths$weight <- c(0,82181,105806,42688,60591,11745,26074,21346)


lodes_data <- paths %>%
  group_by(MSC, ADI_7d, ADI_14d) %>%
  summarise(weight = sum(weight), .groups = "drop") %>%
  mutate(alluvium = row_number()) %>%
  pivot_longer(cols = c(MSC, ADI_7d, ADI_14d), names_to = "x", values_to = "stratum") %>%
  mutate(
    x = factor(x, levels = c("MSC", "ADI_7d", "ADI_14d")),
    stratum = factor(stratum, levels = c(1,0))
  ) %>%
  select(x, stratum, alluvium, weight)

library(ggalluvial)
p <- ggplot(lodes_data,
       aes(x = x, stratum = stratum, alluvium = alluvium, y = weight, fill = stratum)) +
  geom_alluvium(alpha = 0.7, width = 0.5) +
  geom_stratum(width = 0.5, color = "white") +
  labs(x=NULL,y=NULL,title=NULL)+
  theme_minimal(base_size = 16) +
  theme(panel.grid = element_blank(), legend.position = "none",
  axis.text=element_text(color="black"))
ggsave("outplots/fig1/sup/sup1_e.ggalluvial.pdf",p,width=7,height=6,dpi=300)
``` -->


# sup1.f Genomic annotation of k27 regions in different length classes
```{r}
library(ChIPseeker)
library(EnsDb.Hsapiens.v75)
l1 <- k27.env$msc.k27.all[k27.env$msc.k27.all$len <=6,1:3]
l1[,1] <- sapply(strsplit(l1[,1],"chr"),"[",2)
l2 <- k27.env$msc.k27.all[k27.env$msc.k27.all$len > 6,1:3]
l2[,1] <- sapply(strsplit(l2[,1],"chr"),"[",2)
l3 <- k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len <=6,1:3]
l3[,1] <- sapply(strsplit(l3[,1],"chr"),"[",2)
l4 <- k27.env$adi_7d.k27.all[k27.env$adi_7d.k27.all$len >6,1:3]
l4[,1] <- sapply(strsplit(l4[,1],"chr"),"[",2)
l5 <- k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len <=6,1:3]
l5[,1] <- sapply(strsplit(l5[,1],"chr"),"[",2)
l6 <- k27.env$adi_15d.k27.all[k27.env$adi_15d.k27.all$len >6,1:3]
l6[,1] <- sapply(strsplit(l6[,1],"chr"),"[",2)
write.table(l1,file = "output/fig1/msc.k27.k1.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(l2,file = "output/fig1/msc.k27.k2.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(l3,file = "output/fig1/adi_7d.k27.k1.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(l4,file = "output/fig1/adi_7d.k27.k2.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(l5,file = "output/fig1/adi_15d.k27.k1.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(l6,file = "output/fig1/adi_15d.k27.k2.bed",row.names = F,col.names = F,sep = "\t",quote = F)

peaklist <- list(
    MSC1="output/fig1/msc.k27.k1.bed",
    MSC2="output/fig1/msc.k27.k2.bed",
    ADI_7d1="output/fig1/adi_7d.k27.k1.bed",
    ADI_7d2="output/fig1/adi_7d.k27.k2.bed",
    ADI_14d1="output/fig1/adi_15d.k27.k1.bed",
    ADI_14d2="output/fig1/adi_15d.k27.k2.bed")
peak_objects <- lapply(peaklist, function(path) {
    if (!file.exists(path)) {
        stop(paste("File not found:", path, "please check the path or working directory!"))
    }
    # Read with readPeakFile, which returns a GRanges object
    return(readPeakFile(path))
})

# 4. Annotate
peakAnnoList <- lapply(peak_objects, annotatePeak, 
                       TxDb = EnsDb.Hsapiens.v75,
                       tssRegion = c(-2000, 2000), 
                       verbose = FALSE)
p <- plotAnnoBar(peakAnnoList,cex.axis =5)
ggsave(file = "outplots/fig1/sup/sup1_f.all.genome.location.annot.pdf",width=8,height=6,dpi=300)
```
