
<!--
Reproduces main Figure 2: schematic illustrations of ChIP tracks/ridges, dimension-reduction and UMAP of the seven CRR classes, ROI cartoon and per-class per-mark binary-signal z-score heatmap. Inputs (examples): output/clusters.all.binary.colSums.tsv and the clustering outputs (Seurat/UMAP objects). Outputs: outplots/fig2*.pdf.
--> 
fig2.plots

# ====================================================
# a. Schematic: both parts of the input data here need a schematic plot

```{r}
# Load required libraries
library(ggplot2)
library(tidyr)

# 1. Extract data and compute log2 values
# Assume your data are already in the environment
data1 <- log2(k27.env$msc.k27$len)
data2 <- log2(k27.env$adi_7d.k27$len)
data3 <- log2(k27.env$adi_15d.k27$len)

# 2. Build the data frame
# Label each group of data to make later coloring easier
df <- data.frame(
  Value = c(data1, data2, data3),
  Group = rep(c("MSC", "ADI_7d", "ADI_14d"), 
              times = c(length(data1), length(data2), length(data3)))
)

# Ensure Group is a factor so the legend order is preserved
df$Group <- factor(df$Group, levels = c("MSC", "ADI_7d", "ADI_14d"))
# 3. Plot
ggplot(df, aes(x = Value, fill = Group, color = Group)) +
  # alpha sets transparency so overlapping regions are visible
  geom_density(alpha = 0.3, size = 0.8) + 
  # Set axis labels
  labs(
    title = "Density Plot of k27 Lengths",
    x = "log2(Length)",
    y = "Density"
  ) +
  # Use the classic theme
  theme_classic() +
  # Custom colors (optional)
  scale_fill_manual(values = c("#1F77B4", "#2CA02C" , "#D62728")) +
  scale_color_manual(values = c("#1F77B4", "#2CA02C" , "#D62728")) +
  # Optimize the legend position
  theme(legend.position = "top")


```


```{r}
library(ggplot2)
library(ggridges)
library(dplyr)


# 1. Simulated data generator - parameters tuned for visual appeal
set.seed(42)

generate_pretty_signal <- function(name, type, center = 50, weight = 1) {
  n_points <- 5000 * weight # weight increases the number of points, indirectly affecting visual height
  pos <- numeric(0)
  
  if (type == "narrow") {
    # Slightly wider, not too sharp
    pos <- rnorm(n_points, mean = center, sd = 3)
  } else if (type == "medium") {
    pos <- rnorm(n_points, mean = center, sd = 5)
  } else if (type == "bimodal_wide") {
    # Clearly separated bimodal peaks
    pos <- c(rnorm(n_points/2, mean = center - 12, sd = 3), 
             rnorm(n_points/2, mean = center + 12, sd = 3))
  } else if (type == "bimodal_narrow") {
    # Closer bimodal peaks
    pos <- c(rnorm(n_points/2, mean = center - 6, sd = 2.5), 
             rnorm(n_points/2, mean = center + 6, sd = 2.5))
  } else if (type == "hill") {
    # Pronounced arch-shaped hills for H3K36me3/H3K9me3
    pos <- rnorm(n_points, mean = center, sd = 8)
  } else if (type == "plateau") {
    # For H3K27me3, overlapping several high-density normal distributions forms a tall "plateau"
    pos <- c(rnorm(n_points/3, mean = center - 10, sd = 6),
             rnorm(n_points/3, mean = center,      sd = 6),
             rnorm(n_points/3, mean = center + 10, sd = 6))
  }
  return(data.frame(Mark = name, Position = pos))
}

# 2. Build the dataset (manually fine-tune peak positions to avoid complete overlap)
data_list <- list(
  generate_pretty_signal("DNase",    "narrow",         center = 25),
  generate_pretty_signal("H3K27ac",  "narrow",         center = 35),
  generate_pretty_signal("H3K4me3",  "medium",         center = 45),
  generate_pretty_signal("H3K4me1",  "bimodal_wide",   center = 65), # wide bimodal
  generate_pretty_signal("H3K9ac",   "bimodal_narrow", center = 50), # narrow bimodal
  generate_pretty_signal("H3K9me3",  "hill",           center = 40, weight = 1.5), # higher weight, taller peak
  generate_pretty_signal("H3K36me3", "hill",           center = 75, weight = 1.5), # higher weight, taller peak
  generate_pretty_signal("H3K27me3", "plateau",        center = 55, weight = 2.0)  # strongly heightened to form a plateau
)

df <- bind_rows(data_list)

# Set the factor order
df$Mark <- factor(df$Mark, levels = rev(c("DNase", "H3K27ac", "H3K4me3", "H3K4me1", 
                                          "H3K9ac", "H3K36me3", "H3K9me3", "H3K27me3")))

my_colors <- c(
  "DNase" = "#94a3b8", # DNase (neutral gray)
  "H3K27ac" = "#f87171", # H3K27ac (coral red)
  "H3K4me1" = "#4ade80", # H3K4me1 (bright green)
  "H3K4me3" = "#fbbf24", # H3K4me3 (golden yellow)
  "H3K9ac" = "#fb923c", # H3K9ac (orange)
  "H3K36me3" = "#0d9488", # H3K36me3 (teal)
  "H3K27me3" = "#60a5fa", # H3K27me3 (sky blue)
  "H3K9me3" = "#818cf8"  # H3K9me3 (light purple)
  )

# 3. Draw the styled ridge plot
p <- ggplot(df, aes(x = Position, y = Mark, color = Mark,fill=Mark)) +
  # scale controls the degree of overlap: larger values give taller peaks that overlap the layer above
  geom_density_ridges(alpha = 0.5, linewidth = 1.2, scale = 2.0, rel_min_height = 0.0001) +
  # Color scheme: use more vibrant colors
  scale_fill_manual(values = my_colors) +
  scale_color_manual(values = my_colors) +
  
  labs(x = "", y = "") +
  theme_ridges(grid = FALSE) +
  theme(
    legend.position = "none",
    axis.text.y = element_text(color="black"),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line.x = element_line(color = "black",size=0),
    panel.background = element_blank()
  ) +
  xlim(0, 110) # leave a little margin

# Print
print(p)

ggsave("outplots/fig2/fig2_a.simulated.chip.ridges.pdf",width=5,height=4,dpi=300)
```


```{r}
library(ggplot2)
library(patchwork)
library(dplyr)

# 1. Construct simulated data (unchanged)
set.seed(123)
n <- 15
cluster_data <- function(mu, color) {
  data.frame(
    x = rnorm(n, mu[1], 0.5),
    y = rnorm(n, mu[2], 0.5),
    z = rnorm(n, mu[3], 0.5),
    group = color
  )
}

df <- rbind(
  cluster_data(c(2, 8, 5), "Blue"),
  cluster_data(c(5, 8, 2), "Red"),
  cluster_data(c(3, 3, 7), "Yellow"),
  cluster_data(c(8, 3, 4), "Green")
)

my_colors <- c("Blue" = "#5DA5DA", "Red" = "#F15854", "Yellow" = "#FAA43A", "Green" = "#60BD68")

# Run PCA
pca_res <- prcomp(df[, 1:3], scale. = TRUE)
df_pca <- as.data.frame(pca_res$x)
df_pca$group <- df$group

# --- Figure 1: 3D projection schematic ---
project_3d <- function(x, y, z) {
  data.frame(x_proj = x + 0.5 * z, y_proj = y + 0.5 * z)
}
df_3d <- cbind(df, project_3d(df$x, df$y, df$z))

p1 <- ggplot(df_3d) +
  annotate("segment", x=0, xend=10, y=0, yend=0) + 
  annotate("segment", x=0, xend=0, y=0, yend=10) +
  annotate("segment", x=5, xend=15, y=5, yend=5, color="grey80") +
  annotate("segment", x=5, xend=5, y=5, yend=15, color="grey80") +
  annotate("segment", x=0, xend=5, y=0, yend=5, color="grey80") +
  annotate("segment", x=10, xend=15, y=0, yend=5, color="grey80") +
  annotate("segment", x=0, xend=5, y=10, yend=15, color="grey80") +
  annotate("segment", x=10, xend=15, y=10, yend=15, color="grey80") +
  annotate("segment", x=10, xend=10, y=0, yend=10) +
  annotate("segment", x=0, xend=10, y=10, yend=10) +
  annotate("segment", x=5, xend=15, y=15, yend=15, color="grey80") +
  annotate("segment", x=15, xend=15, y=5, yend=15, color="grey80") +
  geom_point(aes(x_proj, y_proj, fill=group), size=3, alpha=0.8,shape=21, color="grey", stroke=0.5) +
  scale_fill_manual(values = my_colors) +
  annotate("text", x=11, y=-0.5, label="x") +
  annotate("text", x=-0.5, y=11, label="y") +
  annotate("text", x=-0.5, y=-0.5, label="z") +
  annotate("segment", x=5, y=5, xend=8, yend=9, arrow=arrow(length=unit(0.2,"cm")), linetype="dashed", color="steelblue", size=0.8) +
  annotate("text", x=9, y=10, label="PC1", color="steelblue", fontface="bold") +
  theme_void(base_size = 16) + theme(legend.position = "none")

# --- Figure 2: 2D PCA (axes changed to solid black) ---
p2 <- ggplot(df_pca, aes(PC1, PC2, fill=group)) +
  geom_point(size=3, alpha=0.8,shape=21, color="grey", stroke=0.5) +
  scale_fill_manual(values = my_colors) +
  # Solid black axes
  geom_segment(aes(x=-3.5, y=-3.5, xend=3.5, yend=-3.5), arrow=arrow(length=unit(0.2,"cm")), color="black", size=0.8) +
  geom_segment(aes(x=-3.5, y=-3.5, xend=-3.5, yend=3.5), arrow=arrow(length=unit(0.2,"cm")), color="black", size=0.8) +
  # Labels
  annotate("text", x=0, y=-4.2, label="PC1", color="black", fontface="bold") +
  annotate("text", x=-4.2, y=0, label="PC2", color="black", angle=90, fontface="bold") +
  coord_cartesian(xlim=c(-4.5, 4.5), ylim=c(-4.5, 4.5)) +
  theme_void(base_size = 16) + theme(legend.position = "none")

# --- Figure 3: 1D PCA (axes changed to solid black) ---
p3 <- ggplot(df_pca, aes(PC1, y=0, fill=group)) +
  geom_point(size=3, alpha=0.8, position = position_jitter(height = 0.05),shape=21, color="grey", stroke=0.5) +
  scale_fill_manual(values = my_colors) +
  # Solid black axis line
  geom_segment(aes(x=-3.5, y=0, xend=3.5, yend=0), arrow=arrow(length=unit(0.2,"cm"), ends="both"), color="black", size=0.8) +
  # Labels
  annotate("text", x=0, y=-0.4, label="PC1", color="black", fontface="bold") +
  ylim(-1, 1) +
  theme_void(base_size = 16) + theme(legend.position = "none")

# --- Combine the panels ---
final_plot <- (p1 | p2 | p3) 

print(final_plot)
ggsave(file = "outplots/fig2/fig2_a_DimRedu.pdf",final_plot,width=12,height=4,dpi=300)
```


fig2_b Radar plot
```{r}
# Load required libraries
library(ggplot2)
library(dplyr)

# 1. Build data: define the start and end position of each stage Peak (on a 1-26 grid)
plot_data <- data.frame(
  stage = c("MSC (Day 0)", "ADI (Day 7)", "ADI (Day 15)", "Unified ROIs"),
  xmin = c(1, 7, 13, 1),   # start grid
  xmax = c(14, 19, 26, 26), # end grid
  y_pos = c(4, 3, 2, 0.5),  # vertical position
  # Use hex color codes for a professional look
  color = c("#4A90E2", "#558B2F", "#C62828", "#2C3E50") 
)

# 2. Start plotting
p <- ggplot() +
  # --- Draw background grid lines (0 to 26) ---
  # Use geom_vline and geom_hline to emulate graph paper
  geom_vline(xintercept = seq(0, 26, 1), color = "gray92", size = 0.3) +
  geom_hline(yintercept = seq(0, 5, 0.5), color = "gray96", size = 0.2) +
  
  # --- Draw the original stage Peaks ---
  # Note: dplyr::filter is used here with an explicit package name to avoid conflicts
  geom_rect(data = dplyr::filter(plot_data, stage != "Unified ROIs"),
            aes(xmin = xmin, xmax = xmax, ymin = y_pos - 0.3, ymax = y_pos + 0.3, fill = color),
            color = "white", size = 0.5) +
  
  # --- Draw the merged Unified ROI ---
  geom_rect(data = dplyr::filter(plot_data, stage == "Unified ROIs"),
            aes(xmin = xmin, xmax = xmax, ymin = y_pos - 0.3, ymax = y_pos + 0.3, fill = color),
            color = "white", size = 0.5) +
  
  # --- Add stage labels on the left ---
  geom_text(data = plot_data, aes(x = -0.5, y = y_pos, label = stage), 
            hjust = 1, fontface = "bold", size = 4, family = "sans") +
  
  # --- Add merge arrows in the middle ---
  annotate("segment", x = 13, xend = 13, y = 1.6, yend = 0.9, 
           arrow = arrow(length = unit(0.2, "cm"), type = "closed"), 
           color = "gray30", size = 0.8) +
  annotate("text", x = 13.5, y = 1.25, label = "Merge & Union", 
           hjust = 0, fontface = "italic", color = "gray30", size = 3.5) +
  
  # --- Color mapping settings ---
  scale_fill_identity() +
  
  # --- Axis and style adjustments ---
  # Relax the left limit (xmin = -6) to fit the text
  scale_x_continuous(limits = c(-6, 27), expand = c(0, 0), breaks = seq(0, 26, 5)) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_void(base_size = 16) +
  labs(
    title = "H3K27me3 ROI Merging Strategy",
    x = "Relative Genomic Position (Bins)",
    y = ""
  ) +
  theme(
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    axis.text = element_text(color = "black"),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(20, 20, 20, 20)
  )

ggsave("outplots/fig2/fig2_a.ROIs.pdf",width=5,height=5,dpi=300)
```

# fig2_a Plot clustering results: UMAP plot and
```{r}
p <- DimPlot(
    seu, group.by = "seurat_clusters",label = T,reduction = "umap",
    label.box = T,raster=T,raster.dpi=c(512,512),label.size=8,
    pt.size = 1.5) + 
    theme_classic(base_size = 16) 
ggsave("outplots/fig2/fig2_a.UMAP.pdf",p,width=8,height=8,dpi=300)
p
dev.off()
```


# fig2_d Radar plot
```{r}
source("Fig2_radar_panel.R")

```


# ====================================================
# fig2_c. Raw values per cluster for each chip and time point (24 dimensions): emission matrix
```{r}
roi_period <- c("d0","d7","d14")
all.output <- data.frame()
for (i in 1:7){
    input <- roi_info.split[roi_info.split$cluster == i,]
    input_list <- roi_list[input$peak]
    for (j in 1:3){
        res_list <- lapply(input_list, function(x) colSums(x[,,j]))
        # Combine into a 10000 * 8 matrix
        result <- do.call(rbind, res_list)
        out <- as.data.frame(colSums(result)) %>% t()
        rownames(out) <- paste0('cluster',i)
        colnames(out) <- paste0(colnames(out),'_',roi_period[j])
        {if (j == 1){cluster_output <- out}
        else{cluster_output <- cbind(cluster_output,out)}}
        
    }
    cluster_output <- as.data.frame(cluster_output)
    cluster_output$sum <- round(sum(input$width)/200)
    all.output <- rbind(all.output,cluster_output)
}
write.table(all.output,file = "output/clusters.all.binary.colSums.tsv",row.names = T,col.names = T,sep = "\t",quote = F)

## Z-score transform the resulting probability density matrix
library(pheatmap)

data <- read.table("output/clusters.all.binary.colSums.tsv", header = TRUE, row.names = 1)

# 2. Drop the final 'sum' column (it interferes with normalization)
counts <- data[, -ncol(data)]

# 3. Remove the effect of cluster size: divide by the row total
counts_prop <- counts / data$sum

# 4. Z-score transform (standardize each chip across all time points)
data_scaled_long <- counts_prop %>%
  rownames_to_column("cluster") %>%
  pivot_longer(cols = -cluster, names_to = "sample", values_to = "val") %>%
  # Extract the marker name (strip the trailing _d0, _d7, _d14)
  mutate(marker = str_remove(sample, "_d(0|7|14)$")) %>%
  # Normalize within each marker group
  group_by(marker) %>%
  mutate(zscore = (val - mean(val)) / sd(val)) %>%
  ungroup()

data_z <- data_scaled_long %>%
  select(cluster, sample, zscore) %>%
  pivot_wider(names_from = sample, values_from = zscore) %>%
  column_to_rownames("cluster")

# Ensure the column order matches the original data
data_z <- data_z[, colnames(counts_prop)]


# 5. Draw the heatmap
# Data cleaning: convert to long format
# Split DNase_d0 into Mark="DNase" and Time="d0"
df <- as.data.frame(data_z)
df_long <- df %>%
  mutate(Cluster = rownames(.)) %>%
  pivot_longer(cols = -Cluster, names_to = "Feature", values_to = "Count") %>%
  # Split the Feature column, for example DNase_d0 -> Mark="DNase", Time="d0"
  separate(Feature, into = c("Mark", "Time"), sep = "_")
df_long$Time <- factor(df_long$Time,c("d0","d7","d14"))
df_long$Cluster <- factor(df_long$Cluster,paste0('cluster',7:1))

# Plot: ggplot2 faceted heatmap
p <- ggplot(df_long, aes(x = Time, y = Cluster, fill = Count)) +
  geom_tile(color = "white") + # draw heatmap tiles
  facet_wrap(~Mark, nrow = 2) + # facet by Mark, a few per row
    scale_fill_distiller(palette = "RdBu", 
                       direction = -1, 
                       limits = c(-2.5, 2.5),
                       oob = scales::squish,
                       name = "Z-score") +
  theme_minimal(base_size = 16) +
  labs(title = "",
       x = "Time Point",
       y = "",
       fill = "Z-score") +
  theme(axis.text = element_text(color="black"),
  panel.border = element_rect(color = "black", fill = NA, linewidth = 0.8),
        strip.background = element_rect(fill = "grey90"), # facet label background
        panel.spacing = unit(0.5, "lines")) # facet spacing
ggsave("outplots/fig2/fig2_c.all.chip.zscore.pdf",p,width=8,height=8,dpi=300)
```




## fig2_e DDRTree pseudotime analysis results
```{r}
library(DDRTree)
library(ggplot2)
library(dplyr)
library(princurve)

# ==================== 1. Prepare data ====================
threshold.emb = 0.6
ndim <- which(res.mfa$eig[,3] >= threshold.emb*100) %>% min()
full_pcs <- as.matrix(res.mfa$ind$coord[, 1:ndim]) # 37000 rows

# Ensure the cell names match
cell_names <- rownames(full_pcs)
cluster_labels <- seu$seurat_clusters[cell_names] # keep the cluster order consistent with the PC matrix

# ==================== 2. DDRTree dimensionality reduction ====================
message("Running DDRTree dimensionality reduction...")
ddrtree_result <- DDRTree(
  X = t(full_pcs),      # input: ndim x 37000
  ndim = 2, 
  ncenter = 1000,        # for 37000 cells, 400-1000 is recommended; 4000 would be very slow
  lambda = 10000,         
  param.gamma = 4,      
  sigma = 0.001
)
message("DDRTree computation finished!")

# ==================== 3. Extract low-dimensional coordinates (Y and Z definitions corrected) ====================
# Z is the projection of cells into the 2D latent space (2 x 37006)
cell_coords <- as.data.frame(t(ddrtree_result$Z))
colnames(cell_coords) <- c("Dim1", "Dim2")
rownames(cell_coords) <- cell_names
cell_coords$cluster <- as.factor(cluster_labels) # this now aligns perfectly with the 37006 rows

# Y is the projection of the principal graph nodes into the 2D latent space (2 x 4000)
node_coords <- as.data.frame(t(ddrtree_result$Y))
colnames(node_coords) <- c("Dim1", "Dim2")


# ==================== 4. Pseudotime analysis (based on cell coordinates Z) ====================
library(igraph)
library(igraph)

# ==================== 1. Build the principal graph and compute node distances ====================
# Get the principal graph node coordinates (4000 x 2)
node_coords_matrix <- t(ddrtree_result$Y)

# Get the adjacency matrix (4000 x 4000)
adj_matrix <- as.matrix(ddrtree_result$stree)

# Create an undirected graph with igraph
g <- graph_from_adjacency_matrix(adj_matrix, mode = "undirected", weighted = TRUE)

# Update the edge weights to the true Euclidean distances between nodes
edges <- as_edgelist(g)
edge_weights <- sapply(1:nrow(edges), function(i) {
  u <- node_coords_matrix[edges[i, 1], ]
  v <- node_coords_matrix[edges[i, 2], ]
  sqrt(sum((u - v)^2)) # Euclidean distance
})
E(g)$weight <- edge_weights


# ==================== 2. Choose the root automatically or manually ====================
# Find the developmental starting point: usually the endpoint (degree-1 node) with the smallest (or largest) Dim1
degrees <- degree(g)
leaves <- which(degrees == 1) # find all terminal nodes

# Automatically pick the terminal node with the smallest Dim1 as the root (adapt this logic to your biological context)
root_node <- leaves[which.max(node_coords_matrix[leaves, 1])]


# ==================== 3. Compute node pseudotime ====================
# Compute shortest-path distances from the root to all principal graph nodes
node_pseudotime <- distances(g, v = root_node, to = V(g), weights = E(g)$weight)
node_pseudotime <- as.numeric(node_pseudotime)


# ==================== 4. Project cells onto the nearest principal graph node (corrected) ====================
message("Computing nearest-neighbor projection of cells onto principal graph nodes...")

# Get the coordinate matrix of the 37006 cells (37006 x 2)
cell_coords_matrix <- as.matrix(cell_coords[, c("Dim1", "Dim2")])

# Quickly compute the nearest node for each of the 37006 cells among the 4000 nodes
# Use Rcpp or efficient matrix operations to avoid slow loops
# Formula: dist_matrix[i, j] = (cell_x - node_x)^2 + (cell_y - node_y)^2
cell_sq <- rowSums(cell_coords_matrix^2)
node_sq <- rowSums(node_coords_matrix^2)
cross_term <- cell_coords_matrix %*% t(node_coords_matrix)

# Square the distance matrix and find, for each row (cell), the column index (node) of the minimum
cell_to_node <- max.col(- (outer(cell_sq, node_sq, "+") - 2 * cross_term))

# Each cell inherits the pseudotime of its nearest node
cell_pseudotime <- node_pseudotime[cell_to_node]

# Normalize to the [0, 1] interval
pseudotime_scaled <- (cell_pseudotime - min(cell_pseudotime)) / 
                     (max(cell_pseudotime) - min(cell_pseudotime))

# Write into the cell_coords data frame
cell_coords$pseudotime <- pseudotime_scaled
message("Pseudotime projection finished!")

# ==================== 5. Plot and save ====================
my_colors <- c(
  "1" = "#1F4E8C", # dark blue
  "2" = "#76A9DC", # light blue    
  "3" = "#F2D36B", # light golden yellow   
  "4" = "#B8860B", # dark mustard yellow  
  "5" = "#C6283D", # strongest red   
  "6" = "#E06C4F", # orange-red
  "7" = "#e29fbb"  # reddish purple
)

library(dplyr)

# Set a random seed for reproducibility
set_seed_val <- 42
set.seed(set_seed_val)

# Group by cluster and randomly sample 20% of the cells per cluster
cell_coords_sub <- cell_coords %>%
  group_by(cluster) %>%
  sample_frac(0.20) %>%
  ungroup() # ungroup to avoid surprises in later plots


# ==================== Extract principal graph edges (edge_coords) ====================
# 1. Extract the principal graph node coordinates and convert them to a data frame
node_coords <- as.data.frame(t(ddrtree_result$Y))
colnames(node_coords) <- c("Dim1", "Dim2")
node_coords$Node_ID <- 1:nrow(node_coords) # number the nodes

# 2. Extract all edges from the adjacency matrix stree
# A value of 1 in the adjacency matrix means the nodes are connected
adj_matrix <- as.matrix(ddrtree_result$stree)
edges_matrix <- which(adj_matrix == 1, arr.ind = TRUE)

# Drop duplicate undirected edges (keep i < j so no line is drawn twice)
edges_matrix <- edges_matrix[edges_matrix[, 1] < edges_matrix[, 2], , drop = FALSE]

# 3. Map both node indices of each edge to their 2D coordinates to build edge_coords
edge_coords <- data.frame(
  x    = node_coords$Dim1[edges_matrix[, 1]],
  y    = node_coords$Dim2[edges_matrix[, 1]],
  xend = node_coords$Dim1[edges_matrix[, 2]],
  yend = node_coords$Dim2[edges_matrix[, 2]]
)



# ---- Optimized Figure 1: DDRTree cluster plot (20% subsample highlighted) ----
p_clu <- ggplot() +
  # 1. Bottom layer: all cells in pale gray as background (overall outline, without distraction)
  geom_point(data = cell_coords_sub, aes(x = Dim1, y = Dim2), 
             size = 0.4, alpha = 0.2, color = "grey90") + 
  # 4. Top layer: principal graph nodes (black outline, white fill, richer than plain shape=1)
  geom_point(data = node_coords, aes(x = Dim1, y = Dim2), 
             color = "black", fill = "grey", size = 1.0, shape = 21, stroke = 0.8) +
  
  # 2. Middle layer: colored cells of the 20% subsample (removes overlap, shows the true distribution)
  geom_point(data = cell_coords_sub, aes(x = Dim1, y = Dim2, color = cluster), 
             size = 0.6, alpha = 0.8) + 
  
  # 3. Top layer: principal graph edges (100% consistent with Figure 2)
  geom_segment(data = edge_coords, aes(x = x, y = y, xend = xend, yend = yend), 
               color = "black", size = 0.6, alpha = 0.9) +

  # 5. Color and label configuration
  scale_color_manual(values = my_colors, name = "Cluster") +
  labs(x = "DDRTree 1", y = "DDRTree 2") +
  
  # 6. Theme polish
  theme_bw(base_size = 16) +
  theme(
    axis.text = element_text(color = "black"),
    panel.grid = element_blank(),
    #aspect.ratio = 1,              # strictly maintain a 1:1 aspect ratio
    legend.position = "right"
  )

# Save Figure 1
ggsave("outplots/fig2/fig2_e1.ddrtree.cluster.pdf", p_clu, width=6, height=6, dpi=300)


# ---- Figure 2: DDRTree pseudotime gradient plot (parameters and layers optimized) ----


# ---- Optimized Figure 2: DDRTree pseudotime gradient plot (20% subsample highlighted) ----
p_pca <- ggplot() +
  # 1. Bottom layer: all cells in pale gray as background
  geom_point(data = cell_coords_sub, aes(x = Dim1, y = Dim2), 
             size = 0.4, alpha = 0.2, color = "grey90") + 

  # 4. Top layer: principal graph edges
  geom_segment(data = edge_coords, aes(x = x, y = y, xend = xend, yend = yend), 
               color = "black", size = 0.8, alpha = 0.9) +
  # 5. Top layer: principal graph nodes
  geom_point(data = node_coords, aes(x = Dim1, y = Dim2), 
             color = "black", fill = "white", size = 1.2, shape = 21, stroke = 0.8) +
  # 2. Middle layer: pseudotime gradient points of the 20% subsample
  geom_point(data = cell_coords_sub, aes(x = Dim1, y = Dim2, color = pseudotime), 
             size = 0.6, alpha = 0.8) +
  
  # 3. Gradient color configuration
  scale_color_viridis_c(
    option = "viridis", 
    direction = -1, 
    name = "Pseudotime",
    breaks = c(0, 0.5, 1.0),
    labels = c("0.0", "0.5", "1.0")
  ) +  
  
  # 6. Labels and theme
  labs(x = "DDRTree 1", y = "DDRTree 2") +
  theme_bw(base_size = 16) +
  theme(
    axis.text = element_text(color = "black"),
    panel.grid = element_blank(),
    #aspect.ratio = 1,
    legend.position = "right"
  )

# Save Figure 2
ggsave("outplots/fig2/fig2_e2.ddrtree.pseudotime.pdf", p_pca, width=6, height=6, dpi=300)


# ---- Figure 3: density distribution of pseudotime across clusters ----
# Build the data frame needed for the density plot
df_density <- data.frame(
  cell = rownames(cell_coords),
  pseudotime = cell_coords$pseudotime,
  cluster = cell_coords$cluster
)

p_density <- ggplot(df_density, aes(x = pseudotime, color = cluster, fill = cluster)) +
  # 1. Draw the density curves:
  #    - size = 1.2 increases line thickness for legibility
  #    - alpha = 0.08 very low fill transparency (a subtle accent that keeps overlapping colors clear)
  #    - to omit shading entirely, set fill = NULL or alpha = 0
  geom_density(size = 1.6, alpha = 0.05) + 
  
  # 2. Color configuration:
  #    - ensure both line color (color) and fill color (fill) use your my_colors
  scale_color_manual(values = my_colors, name = "Cluster") +
  scale_fill_manual(values = my_colors, name = "Cluster") +
    scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  
  # 3. Labels and theme
  labs(x = "Pseudotime", y = "Density") +
  theme_classic(base_size = 16) +
  theme(
    axis.text = element_text(color = "black"),
    axis.line = element_line(color = "black"), # strengthen the axis lines
    legend.position = "right"
  )

ggsave("outplots/fig2/fig2_f.ddrtree.freq.pdf", p_density, width=7, height=6, dpi=300)


p_violin <- ggplot(df_density, aes(x = cluster, y = pseudotime, color = cluster, fill = cluster)) +
  
  # 1. Draw violin plots (the "gourd" look):
  #    - scale = "width" makes every violin equally wide at its maximum (or use "area", "count")
  #    - alpha = 0.5 controls fill transparency
  #    - trim = FALSE prevents the tails from being cut off abruptly
  geom_violin(scale = "width", alpha = 0.5, trim = FALSE, linewidth = 1) + 
  
  # 2. [optional but recommended] add a mini boxplot inside the violin to show the median and quartiles
  geom_boxplot(width = 0.2, color = "black", alpha = 0.7, outlier.shape = NA, 
               fill = "white", linewidth = 0.8) +
  
  # 3. Color configuration (mapped to both line and fill)
  scale_color_manual(values = my_colors, name = "Cluster") +
  scale_fill_manual(values = my_colors, name = "Cluster") +
  
  # 4. Axis adjustments (Y axis is pseudotime, X axis is the clusters)
  scale_y_continuous(expand = c(0.02, 0)) +
  
  # 5. Labels and theme
  labs(x = "Cluster", y = "Pseudotime") +
  theme_classic(base_size = 16) +
  theme(
    axis.text = element_text(color = "black",size = 16),
    axis.text.x = element_text(hjust = 1,size = 16), # rotate 45 degrees to prevent overlap when cluster names are long
    axis.line = element_line(color = "black"), 
    legend.position = "none" # since the X axis is already colored, legend.position = "none" omits the legend
  )

print(p_violin)
ggsave("outplots/fig2/fig2_f.ddrtree.freq.vioplot.pdf", p_violin, width=8.5, height=6, dpi=300)



```



# The following are several supplementary figures 

## sup2_a Bar chart of cell counts per cluster
```{r}
# Load required libraries
library(ggplot2)
library(scales)

# 1. Prepare data
df <- data.frame(
  cluster = paste0('cluster',1:7),
  count = as.data.frame(table(seu$seurat_clusters))[,2]
)

# 2. Define the colors you specified
my_colors <- c(
  "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)

# 3. Plot
p <- ggplot(df, aes(x = cluster, y = count, fill = cluster)) +
  # Draw the bar chart; a thin black border makes the colors easier to distinguish
  geom_bar(stat = "identity", color = "black", linewidth = 0.3) + 
  # Add n=xx labels
  geom_text(aes(label = paste0("n=", count)), 
            vjust = -0.5, 
            size = 4) +
  # Use your custom colors
  scale_fill_manual(values = my_colors) +
  # Set the theme
  theme_classic(base_size = 16) +
  # Hide the legend and set the title
  theme(legend.position = "none") +
  labs(x = "", y = "Number of ROIs") +
  # Reserve space at the top so labels are not clipped
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
ggsave(file = "outplots/fig2/sup2_a.clusters.barplots.pdf",p,width=6,height=6,dpi = 300)
```



## sup2_b Heatmap of Euclidean distances from hierarchical clustering
```{r}
# Assumptions:
# X: 37006 x 29 matrix/data frame
# cluster: length-37006 vector giving the cluster label of each sample

library(dplyr)

emb <- full_pcs
# 1. Combine the cluster information with the dimensionality-reduction matrix
df <- as.data.frame(emb)
df$cluster <- cluster_labels   # works whether cluster is a factor or numeric

# 2. Average by cluster to obtain the center of each cluster
cluster_centers <- df %>%
  group_by(cluster) %>%
  dplyr::summarise(across(where(is.numeric), mean, .names = "{.col}")) %>%
  ungroup()

# Take the cluster column out as row names; the rest is the numeric matrix
center_mat <- as.matrix(cluster_centers[ , -1])   # drop the first cluster column
rownames(center_mat) <- as.character(cluster_centers$cluster)

# 3. Compute the Euclidean distance matrix between clusters
dist_mat <- dist(center_mat, method = "euclidean")

# 4. Perform hierarchical clustering
hc <- hclust(dist_mat, method = "ward.D2")  # the method can be changed to "complete", "ward.D2", etc.

# 5. Draw the dendrogram
plot(hc, main = "Hierarchical Clustering of Cluster Centers",
     xlab = "Cluster", sub = "", ylab = "Euclidean distance")

# install.packages(c("ggplot2", "ggdendro", "patchwork", "dplyr", "tidyr"))
library(ggplot2)
library(ggdendro)
library(patchwork)
library(dplyr)
library(tidyr)

## ---------- 1) Normalize dist_mat to a symmetric matrix + dist ----------
if (inherits(dist_mat, "dist")) {
  D <- dist_mat
  Dmat <- as.matrix(D)
} else {
  Dmat <- as.matrix(dist_mat)
  # If only the lower triangle is filled in, complete the upper triangle
  Dmat[upper.tri(Dmat)] <- t(Dmat)[upper.tri(Dmat)]
  diag(Dmat) <- 0
  D <- as.dist(Dmat)
}

# Ensure row and column names exist (otherwise use 1..n)
if (is.null(rownames(Dmat)) || is.null(colnames(Dmat))) {
  n <- nrow(Dmat)
  nm <- as.character(seq_len(n))
  rownames(Dmat) <- nm
  colnames(Dmat) <- nm
}

## ---------- 2) Hierarchical clustering (directly on your distances) ----------
hc <- hclust(D, method = "ward.D2")

## Match the tree order to the heatmap order (critical)
ord <- hc$order
labs_ord <- rownames(Dmat)
Dmat_ord <- Dmat[labs_ord, labs_ord, drop = FALSE]


df <- as.data.frame(Dmat_ord) |>
  tibble::rownames_to_column("row") |>
  pivot_longer(-row, names_to = "col", values_to = "dist")

# For the half matrix, use the position i/j of rows and columns in the current order
pos <- tibble::tibble(name = labs_ord, idx = seq_along(labs_ord))

df2 <- df |>
  left_join(pos, by = c("row" = "name")) |>
  rename(i = idx) |>
  left_join(pos, by = c("col" = "name")) |>
  rename(j = idx) |>
  mutate(
    # Keep only the lower triangle (including the diagonal); for the upper triangle, change i >= j to i <= j
    keep = i <= j,
    diag = (i == j),
    # Mark the diagonal in gray with a separate fill value
    fill_val = ifelse(diag, NA_real_, dist),
    label = ifelse(keep, sprintf("%.2f", dist), NA_character_)
  ) |>
  filter(keep)

# Warm palette: near (small) = dark, far (large) = light
# Controlled by low/high in scale_fill_gradient: low is the color for small values, high for large values
p_tri <- ggplot(df2, aes(x = col, y = row)) +
  # First draw the distances in the triangle
  geom_tile(aes(fill = fill_val), color = "white", linewidth = 0.3) +
  # Then cover the diagonal with gray squares
  geom_tile(
    data = df2 |> filter(diag),
    fill = "grey70",
    color = "white",
    linewidth = 0.3
  ) +
  # Write the values (the diagonal is written as 0.00; use filter(!diag) below to skip it)
  geom_text(aes(label = label), size = 3) +
  scale_fill_gradient(
    low = "#7f0000",   # small distance: dark warm color
    high = "#fff5eb",  # large distance: light warm color
    na.value = "grey70",
    name = "Distance"
  ) +
  coord_fixed(ratio = 1) +
  labs(x = NULL, y = NULL) +
  theme_minimal(base_size = 16) +
  theme(
    axis.text = element_text(color="black"),
    panel.grid = element_blank(),
    legend.position = "right"
  )

ggsave("outplots/fig2/sup2_b.Euclidean.heatmap.pdf",p_tri,width=6,height=6,dpi=300)
```


## sup2_c True neighbor relationships between clusters and their significance levels
```{r}
## First quantify the relative positions of the clusters
## We are strict here and consider only the neighbor distribution
```{r}

# One-sample proportion test
all.roi.nei.mat <- data.frame(matrix(nrow=7,ncol=7))
all.roi.nei.pval <- data.frame(matrix(nrow=7,ncol=7))
out3 <- c()
for (i in 1:7){
    input <- roi_info.split[roi_info.split$cluster == i,]
    input <- rbind(data.frame(chr=input$chr,start=input$start-200,end=input$start-10,peak=input$peak),data.frame(chr=input$chr,start=input$end+10,end=input$end+200,peak=input$peak))
    out <- bt.intersect(a=roi_info.split,b=input,wa=T) %>% unique
    tab1 <- as.data.frame(table(out$V6))
    tab2 <- as.data.frame(table(roi_info.split$cluster))
    out1 <- c()
    out2 <- c()
    for (j in 1:7){
        k=tab1[j,2]
        n=sum(tab1[,2])
        p0=tab2[j,2]/sum(tab2[,2])
        result <- prop.test(x = k,n = n,p = p0,alternative = "greater",correct = T)
        out1 <- c(out1,result$estimate-p0) # positive if the proportion exceeds the overall value, negative otherwise
        out2 <- c(out2,ifelse(result$p.value<=2.2e-16,2.2e-16,result$p.value))
    }
    out2 <- -log10(out2)
    all.roi.nei.mat[i,] <- out1
    all.roi.nei.pval[i,] <- out2
    out3 <- c(out3,nrow(out)/nrow(input))
}
all.roi.nei.mat$X8 <- out3



library(ggplot2)
library(reshape2)
# Convert the matrix to a long-format data frame
rownames(all.roi.nei.mat) <- paste0('cluster',1:7)
colnames(all.roi.nei.mat) <- c(paste0('cluster',1:7),"sum")
df <- as.data.frame(all.roi.nei.mat[,1:7]) %>%
  rownames_to_column(var = "row")
# Convert to long format
mat_long <- df %>%
  pivot_longer(cols = -row, names_to = "col", values_to = "value")

rownames(all.roi.nei.pval) <- paste0('cluster',1:7)
colnames(all.roi.nei.pval) <- paste0('cluster',1:7)


df <- as.data.frame(all.roi.nei.pval) %>%
  rownames_to_column(var = "row")
# Convert to long format
pval_long <- df %>%
  pivot_longer(cols = -row, names_to = "col", values_to = "pval")

# Merge the two data sets
data <- cbind(mat_long, pval=pval_long$pval)
data$row <- factor(data$row,levels=paste0('cluster',1:7))
data$col <- factor(data$col,levels=paste0('cluster',7:1))
alpha <- 0.05

## 2. Compute -log10(p) for significance; set non-significant cells to NA
data2 <- data %>%
  mutate(
    logp = ifelse(pval > 1.301, pval, NA_real_)
  )

## 3. Plot:
##    - background: significant tiles use a logp gradient from light to dark red; non-significant tiles are uniformly light gray (na.value)
##    - value > 0: circles; value < 0: triangles; size follows the absolute value (larger magnitude, larger point)

p <- ggplot(data2, aes(x = row, y = col)) +
  # Heatmap background: significant tiles shaded by logp, non-significant in light gray
  geom_tile(aes(fill = logp), color = "grey70", linewidth = 1.0) +

  # Positive values: circles (21: filled circle with a stroke), sized by value
  geom_point(
    data = ~ subset(.x, !is.na(logp)),
    aes(size = value),
    shape = 21,
    stroke = 1,
    color = "black"
  ) +

  # Negative values: upward triangles (24: hollow triangle with a stroke, fillable), sized by absolute value
  geom_point(
    data = ~ subset(.x, is.na(logp)),
    aes(size = 0.02),      # == abs(value) because value < 0 here
    shape = 24,
    stroke = 1.0,
    color = "black"
  ) +

  # Background colors: light to dark red for significant values; NA -> light gray otherwise
  scale_fill_gradientn(
    colours  = c("#fee2e2", "#fca5a5", "#ef4444", "#b91c1c"), # light red to dark red
    na.value = "snow",                                      # non-significant: light gray
    name     = "-log10(p)",
    limits   = range(data2$logp, na.rm = TRUE)
  ) +

  # Point size: always interpreted as the absolute value
  scale_size_continuous(
    range = c(2, 10),
    name  = "|value|"
  ) +

  coord_fixed(ratio = 1) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(color="black",angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(color="black"),
    legend.position  = "right"
  ) 
ggsave(file = "outplots/fig2/sup2_c.clu1-7.nei.prop.test.pdf",p,width=7,height=6,dpi=300)


```




# sup2_d Draw the heatmap and the corresponding cluster plot
```{r}
{
feats_coverage <- orig_df[,grep("_coverage",colnames(orig_df),value=T)]
mark_names <- c("DNase", "H3K27ac", "H3K27me3", "H3K36me3", "H3K4me1", "H3K4me3", "H3K9ac", "H3K9me3")
normalize_df_clipped <- function(df) {
  # Combine all numeric columns into one vector (non-numeric columns are assumed to be absent)
  if (!all(sapply(df, is.numeric))) {
    warning("Data frame contains non-numeric columns. They will be ignored or cause errors.")
  }
  
  combined_vec <- unlist(df, use.names = FALSE)
  combined_vec <- combined_vec[!is.na(combined_vec)]  # remove NA
  
  min_val <- quantile(combined_vec, probs = 0.05, na.rm = TRUE)
  max_val <- quantile(combined_vec, probs = 0.95, na.rm = TRUE)
  
  # Guard against a zero denominator
  if (max_val == min_val) {
    stop("0.05 and 0.95 quantiles are equal; cannot normalize.")
  }
  
  # Define the clipping function
  clip_and_scale <- function(x) {
    x_clipped <- pmin(pmax(x, min_val), max_val)  # clip x to [min_val, max_val]
    (x_clipped - min_val) / (max_val - min_val)
  }
  
  df_normalized <- as.data.frame(lapply(df, clip_and_scale))
  names(df_normalized) <- names(df)  # keep the original column names
  
  return(df_normalized)
}
out1 <- data.frame(matrix(nrow=nrow(feats_df),ncol=24),row.names=rownames(feats_df))
colnames(out1) <- as.vector(t(outer(mark_names, c("_d0", "_d7", "_d15"), paste, sep = "")))
for (j in 1:length(mark_names)){
    l1 <- feats_coverage[,grep(mark_names[j],colnames(feats_coverage),value=T)]
    l2 <- normalize_df_clipped(l1)
    out1[,((j-1)*3+1):(j*3)] <- l2
}

row_annot <- data.frame(cluster=cluster_labels,row.names=names(cluster_labels))
#col_annot <- data.frame(time=rep(c(1,2,3),8),row.names=colnames(out1))
ord1 <- order(row_annot$cluster)
#ord2 <- order(col_annot$time)
out1_sorted <- out1[ord1,]
#out2_sorted <- out2[ord1,ord2]
row_ann_sorted <- row_annot[ord1, , drop = FALSE]
#col_ann_sorted <- col_annot[ord2, , drop = FALSE]
library(pheatmap)
gap_rows <- sapply(unique(row_ann_sorted[,1]),function(x) which(row_ann_sorted[,1] == x)[1])[-1]-1
gap_cols <- seq(3,24,3)


library(dendextend)
order_within_group_hclust <- function(mat, row_annot, dist_method = "euclidean", hclust_method = "ward.D2"){
  ord_all <- character(0)
  for(g in unique(row_annot$cluster)){
    idx <- which(row_annot$cluster == g)
    sub <- mat[idx, , drop = FALSE]
    if(nrow(sub) <= 2){
      ord_all <- c(ord_all, rownames(sub))
      next
    }
    d <- dist(sub, method = dist_method)
    hc <- hclust(d, method = hclust_method)
    ord <- rownames(sub)[hc$order]
    ord_all <- c(ord_all, ord)
  }
  mat[ord_all, , drop = FALSE]
}
out1_hclust <- order_within_group_hclust(out1_sorted,row_ann_sorted)
row_ann_hclust <- data.frame(group=row_ann_sorted[match(rownames(out1_hclust),rownames(row_ann_sorted)),1],row.names=rownames(out1_hclust))
row_ann_hclust$group <- as.character(row_ann_hclust$group)
out1_hclust <- out1_hclust[,c(1:3,4:6,13:15,16:18,19:21,10:12,22:24,7:9)]


library(RColorBrewer)
#pdf("outplots/hclust.pheatmap.pdf",width=4000,height=6000)

# --- 1. Define the main heatmap colors (a more refined red-yellow-blue) ---
# Use 11 gradient steps for smoother transitions
my_heatmap_color <- colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))(101)
my_heatmap_color <- colorRampPalette(c("#5171be", "#f7f7f7", "#EB5354"))(101)

# --- 2. Define annotation colors (reflecting the repression/activation gradient) ---
# 1-4 decreasing repression (dark blue -> light sky blue)
# 5-7 decreasing activation (dark red -> light orange)
#row_ann_hclust$cluster <- as.character(row_ann_hclust$group)

# 2. Define the color list (note: "cluster" here must exactly match the column name in row_ann_hclust

ann_colors <- list(
    group = c(
        "1" = "#1F4E8C", # dark blue
        "2" = "#76A9DC", # light blue    
        "3" = "#F2D36B", # light golden yellow   
        "4" = "#B8860B", # dark mustard yellow  
        "5" = "#C6283D", # strongest red   
        "6" = "#E06C4F", # orange-red
        "7" = "#e29fbb"  # reddish purple
    )
)

# --- 2. Recompute the split lines ---
# To place lines between d0, d7 and d15 as well:
# With 24 columns in total and one group per 3 columns, you may also want lines between 1|2 and 2|3
# then gaps_col would not be just 3, 6, 9... but 1, 2, 3, 4, 5, 6... 
# but such dense lines look poor. Prefer thick lines between groups of 3 columns and inspect the thin lines in the exported PDF.
# Here we restore the normal between-group splits:
all_gaps_col <- seq(3, ncol(out1_hclust) - 1, by = 3)

# --- 3. Plot (corrected version) ---
png("outplots/fig2/sup2_d.hclust.pheatmap.png",width=3000,height=3000,res=300)
pheatmap(
    as.matrix(out1_hclust), 
    scale = "none",
    cluster_cols = FALSE,
    cluster_rows = FALSE, 
    
    # [key change 1] remove the border to fix the washed-out look and warnings
    border_color = NA, 
    
    # Annotation
    annotation_row = row_ann_hclust, 
    annotation_colors = ann_colors,
    
    # Split lines
    gaps_row = gap_rows,
    gaps_col = all_gaps_col,
    
    # Colors
    color = my_heatmap_color,
    
    # [key change 2] if the colors still look pale, set the breaks manually
    # assuming your data range is 0-1, the settings below shift the mid tones toward 0.5
    # breaks = seq(0, 1, length.out = 101), 
    
    # Rendering optimization
    use_raster = TRUE,
    raster_device = "png", 
    raster_quality = 5, 
    
    show_rownames = FALSE,
    show_colnames = FALSE,
    legend_breaks = c(0, 0.5, 1), 
    legend_labels = c("Low", "Mid", "High"),
    main = "Refined Chromatin Landscape"
) %>% print()
dev.off()
}
```


# ====================================================
## sup2_e. chromHMM annotation results
## Compare our annotation results with chromHMM
```{r}
chromHMM.annot <- read.table(file = "~/Project/chromHMM/18chrhmm_mixed.annot.tsv",header = T,sep = "\t")
msc_chromHMM.results <- fread(file = "~/Project/chromHMM/learnmodel/MSC_18_chrhmm_segments.bed",header = F)
adi_7d_chromHMM.results <- fread(file = "~/Project/chromHMM/learnmodel/ADI-7d_18_chrhmm_segments.bed",header = F)
adi_14d_chromHMM.results <- fread(file = "~/Project/chromHMM/learnmodel/ADI-14d_18_chrhmm_segments.bed",header = F)

# First draw an overall stacked bar chart of the chromHMM annotations
source("~/Project/chromHMM/src/chromHMM_composition.R")

test.composition <- chromHMM_composition(
  win = l1.win,
  chromHMM.annot = chromHMM.annot,
  time = "MSC",
  cluster = "cluster1"
)

all.composition <- data.frame()
for (i in 1:7){
    input <- read.table(file = paste0('output/clusters/CRRs/cluster',i,'.bed'))[,1:3]
    input.bg <- read.table(file = paste0('output/clusters/controls/cluster',i,'.bg.bed'))[,1:3]
    
    output <- data.frame()
    l1 <- bt.intersect(a=msc_chromHMM.results,b=input,wo=T) %>% unique()
    l1.win <- bt.makewindows(b=as.data.frame(l1),w = 200, i = "src")
    l1.win <- bt.intersect(a=l1.win,b=input,wa=T) %>% unique()
    l1.win <- annotate_bed_chromhmm(chromHMM.annot,l1.win)
    l1.composition <- chromHMM_composition(win = l1.win,chromHMM.annot = chromHMM.annot, time = "MSC",  cluster = paste0('cluster',i))


    l2 <- bt.intersect(a=adi_7d_chromHMM.results,b=input,wo=T) %>% unique()
    l2.win <- bt.makewindows(b=as.data.frame(l2),w = 200, i = "src")
    l2.win <- bt.intersect(a=l2.win,b=input,wa=T) %>% unique()
    l2.win <- annotate_bed_chromhmm(chromHMM.annot,l2.win)
    l2.composition <- chromHMM_composition(win = l2.win,chromHMM.annot = chromHMM.annot, time = "ADI-7d",cluster = paste0('cluster',i))


    l3 <- bt.intersect(a=adi_14d_chromHMM.results,b=input,wo=T) %>% unique()
    l3.win <- bt.makewindows(b=as.data.frame(l3),w = 200, i = "src")
    l3.win <- bt.intersect(a=l3.win,b=input,wa=T) %>% unique()
    l3.win <- annotate_bed_chromhmm(chromHMM.annot,l3.win)
    l3.composition <- chromHMM_composition(win = l3.win,chromHMM.annot = chromHMM.annot, time = "ADI-14d",cluster = paste0('cluster',i))    

    all.composition <- bind_rows(all.composition,l1.composition,l2.composition,l3.composition)
}


annotation.order <- c(
  "Active promoter",
  "Active enhancer",
  "Primed enhancer",
  "Transcribed",
  "Mixed/Bivalent",
  "Polycomb repressed",
  "Heterochromatin/repeats",
  "Quiescent"
)

time.order <- c(
  "MSC",
  "ADI-7d",
  "ADI-14d"
)

cluster.order <- paste0(
  "cluster",
  1:7
)

plot.data <- all.composition %>%
  mutate(
    Annotation = factor(
      Annotation,
      levels = annotation.order
    ),

    Time = factor(
      Time,
      levels = time.order
    ),

    Cluster = factor(
      Cluster,
      levels = cluster.order
    )
  )

chromHMM.colors <- c(
  "Active promoter" = "#980000",
  "Active enhancer" = "#E64B35",
  "Primed enhancer" = "#F4B183",
  "Transcribed" = "#70AD47",
  "Mixed/Bivalent" = "#8064A2",
  "Polycomb repressed" = "#2E75B6",
  "Heterochromatin/repeats" = "#595959",
  "Quiescent" = "#D9D9D9"
)

source("~/Project/chromHMM/src/chromHMM_composition.R")
p1 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster1",
  show.y.axis = TRUE
)

p2 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster2"
)

p3 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster3"
)

p4 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster4"
)

p5 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster5",
  show.y.axis = TRUE
)

p6 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster6"
)

p7 <- make_chromHMM_cluster_panel(
  data = plot.data,
  cluster.name = "cluster7"
)
p.chromHMM.composition <- (
  (
    p1 | p2 | p3 | p4
  ) /
  (
    p5 | p6 | p7 | guide_area()
  )
) +

  plot_layout(
    guides = "collect",
    widths = c(
      1,
      1,
      1,
      1
    ),
    heights = c(
      1,
      1
    )
  )

p.chromHMM.composition <-
  p.chromHMM.composition &

  theme(
    legend.position = "right",

    legend.justification = "center",

    legend.box.just = "left",

    legend.title = element_text(
      face = "bold",
      size = 9
    ),

    legend.text = element_text(
      size = 8.5
    ),

    legend.key.width = unit(
      0.45,
      "cm"
    ),

    legend.key.height = unit(
      0.38,
      "cm"
    )
  )

ggsave(
  filename = "outplots/fig2/sup2_d.ChromHMM_composition.pdf",
  plot = p.chromHMM.composition,
  width = 10.5,
  height = 7,
  units = "in"
)



## 

annotate_bed_chromhmm <- function(annot_map,bed_df) {

  annot_map[,2] <- paste0('E',annot_map[,2])
  annot_map <- setNames(annot_map$Annotation, annot_map$stage)
  # 1. Define the mapping from E1-E18 to specific annotations
  # 2. Ensure the input variable is a data.frame or otherwise modifiable
  bed_df <- as.data.frame(bed_df, stringsAsFactors = FALSE)
  
  # 3. Replace the fourth column (assuming the fourth column by name or position)
  # Use match for efficient vectorized replacement; unmatched entries are left unchanged
  matched_idx <- match(bed_df[, 4], names(annot_map))
  
  # Replace only successful matches so that non-E1-E18 entries do not become NA
  to_replace <- !is.na(matched_idx)
  bed_df[to_replace, 4] <- annot_map[bed_df[to_replace, 4]]
  
  # 4. Return the processed data frame variable
  return(bed_df)
}

source("~/Project/chromHMM/src/chromHMM_enrichment.R")

all.output <- data.frame()
for (i in 1:7){
    input <- read.table(file = paste0('output/clusters/CRRs/cluster',i,'.bed'))[,1:3]
    input.bg <- read.table(file = paste0('output/clusters/controls/cluster',i,'.bg.bed'))[,1:3]
    
    output <- data.frame()
    l1 <- bt.intersect(a=msc_chromHMM.results,b=input,wo=T) %>% unique()
    l1.win <- bt.makewindows(b=as.data.frame(l1),w = 200, i = "src")
    l1.win <- bt.intersect(a=l1.win,b=input,wa=T) %>% unique()
    l1.win <- annotate_bed_chromhmm(chromHMM.annot,l1.win)
    l1.bg <- bt.intersect(a=msc_chromHMM.results,b=input.bg,wo=T) %>% unique()
    l1.bg.win <- bt.makewindows(b=as.data.frame(l1.bg),w = 200, i = "src")
    l1.bg.win <- bt.intersect(a=l1.bg.win,b=input.bg,wa=T) %>% unique()
    l1.bg.win <- annotate_bed_chromhmm(chromHMM.annot,l1.bg.win)

    out1 <- chromHMM_enrichment(l1.win,l1.bg.win,chromHMM.annot)
    out1 <- data.frame(Time = "MSC",Cluster = paste0('cluster',i),Annotation = out1$Annotation,log2_fold_enrichment = out1$log2_fold_enrichment,p_value= out1$p_value)


    l2 <- bt.intersect(a=adi_7d_chromHMM.results,b=input,wo=T) %>% unique()
    l2.win <- bt.makewindows(b=as.data.frame(l2),w = 200, i = "src")
    l2.win <- bt.intersect(a=l2.win,b=input,wa=T) %>% unique()
    l2.win <- annotate_bed_chromhmm(chromHMM.annot,l2.win)
    l2.bg <- bt.intersect(a=adi_7d_chromHMM.results,b=input.bg,wo=T) %>% unique()
    l2.bg.win <- bt.makewindows(b=as.data.frame(l2.bg),w = 200, i = "src")
    l2.bg.win <- bt.intersect(a=l2.bg.win,b=input.bg,wa=T) %>% unique()
    l2.bg.win <- annotate_bed_chromhmm(chromHMM.annot,l2.bg.win)

    out2 <- chromHMM_enrichment(l2.win,l2.bg.win,chromHMM.annot)
    out2 <- data.frame(Time = "ADI-7d",Cluster = paste0('cluster',i),Annotation = out2$Annotation,log2_fold_enrichment = out2$log2_fold_enrichment,p_value= out2$p_value)


    l3 <- bt.intersect(a=adi_14d_chromHMM.results,b=input,wo=T) %>% unique()
    l3.win <- bt.makewindows(b=as.data.frame(l3),w = 200, i = "src")
    l3.win <- bt.intersect(a=l3.win,b=input,wa=T) %>% unique()
    l3.win <- annotate_bed_chromhmm(chromHMM.annot,l3.win)
    l3.bg <- bt.intersect(a=adi_14d_chromHMM.results,b=input.bg,wo=T) %>% unique()
    l3.bg.win <- bt.makewindows(b=as.data.frame(l3.bg),w = 200, i = "src")
    l3.bg.win <- bt.intersect(a=l3.bg.win,b=input.bg,wa=T) %>% unique()
    l3.bg.win <- annotate_bed_chromhmm(chromHMM.annot,l3.bg.win)

    out3 <- chromHMM_enrichment(l3.win,l3.bg.win,chromHMM.annot)
    out3 <- data.frame(Time = "ADI-14d",Cluster = paste0('cluster',i),Annotation = out3$Annotation,log2_fold_enrichment = out3$log2_fold_enrichment,p_value= out3$p_value)


    output <- rbind(out1,out2,out3)
    all.output <- rbind(all.output,output)
}

all.enrichment <- all.output %>%
  mutate(
    FDR_global = p.adjust(
      p_value,
      method = "BH"
    )
  )

source("~/Project/chromHMM/src/chromHMM_enrichment.plots.R")
p.chromHMM <- plot_chromHMM_enrichment(
  enrichment.data = all.enrichment,
  effect.col = "log2_fold_enrichment",
  fdr.col = "FDR_global",
  color.limit = 4,
  min.effect.for.star = 1
)

ggsave(
  filename = "outplots/fig2/sup2_f.ChromHMM_enrichment_heatmap.pdf",
  plot = p.chromHMM,
  width = 13,
  height = 5.5,
  units = "in"
)

ggsave(
  filename = "outplots/fig2/sup2_f.ChromHMM_enrichment_heatmap.png",
  plot = p.chromHMM,
  width = 13,
  height = 5.5,
  units = "in",
  dpi = 600
)


```


# Supplementary Table S2
# The clustering sensitivity analysis output referenced by Supplementary
# Table S2 is produced by pipeline/src/run_05_sensitivity_clustering.R;
# no figure is drawn from it in this script.



