fig3.plots
# =====================================================
## fig3_a Enrichment at the regions of target genes relative to control
```{r}

loc.annot.class <- c("promoter","exon","intron","distal_regulatory","intergenic_far")
all.loc.annot.df <- data.frame()
for (i in 1:7){
    input <- roi_info.split[roi_info.split$cluster == i,]
    input.tg <- read.table(file = paste0("output/peaks/cluster_tg/cluster",i,".rois.targetGene.tsv"),header = T)
    input.bg.tg <- read.table(file = paste0("output/peaks/cluster_tg/cluster",i,".bg.targetGene.tsv"),header = T)
    output <- data.frame()
    for (j in 1:length(loc.annot.class)){
        a1=input.tg[input.tg$final_type == loc.annot.class[j],] %>% nrow()
        a2=nrow(input.tg) - a1
        b1=input.bg.tg[input.bg.tg$final_type == loc.annot.class[j],] %>% nrow()
        b2=nrow(input.bg.tg) - b1
        result <- fisher.test(matrix(c(a1,a2,b1,b2),nrow=2,ncol=2))
        out <- data.frame(cluster = paste0('cluster',i),annot = loc.annot.class[j],OR = result$estimate,upper = result$conf.int[2],lower = result$conf.int[1],pval=result$p.value)
        output <-rbind(output,out)
    }
    all.loc.annot.df <- rbind(all.loc.annot.df, output)

}


# Load required libraries
library(ggplot2)
library(dplyr)

# 1. Prepare the data (the data frame is assumed to be named all.loc.annot.df)
# For illustration only: make sure the data are converted to numeric if this has not been done yet
df <- all.loc.annot.df
df$pval <- ifelse(df$pval < 2.2e-16,2.2e-16,df$pval)

# Convert OR to beta (log e)
df$beta <- log(df$OR)

# Handle the significance magnitude:
# Set a threshold (for example p < 0.05).
# Non-significant points are given a fixed small value; significant points are mapped from -log10(p) onto a larger range.
df <- df %>%
  mutate(
    is_sig = ifelse(pval < 0.05, "Significant", "Not Significant"),
    log_p = -log10(pval),
    # Create a size variable for plotting:
    # Non-significant points all get a very small value (e.g. 1), while significant points start at 4 and scale upwards
    plot_size = ifelse(pval >= 0.05, 1, 4 + (log_p / max(log_p) * 6))
  )

df <- df[df$annot != "intergenic_far",]
# 2. Factor ordering (keep the logical order)
df$cluster <- factor(df$cluster, levels = rev(unique(df$cluster)))
df$annot <- factor(df$annot, levels = c("promoter", "exon", "intron", "distal_regulatory"))

# 3. Plot
p <- ggplot(df, aes(x = annot, y = cluster)) +
  # Draw background grid lines (optional, adds visual guidance)
  geom_point(aes(size = plot_size, color = beta)) +
  
  # Colour settings: use gradient2 with a very narrow transition around 0
  # so that even 0.001 immediately takes on a colour bias
  scale_color_gradient2(
    low = "#377EB8",      # blue (Diluted)
    mid = "white",        # white at 0
    high = "#E41A1C",     # red (Enriched)
    midpoint = 0,
    name = expression(beta == log(OR)) # legend title displaying the formula
  ) +
  
  # Size settings: drop the continuous mapping and use the precomputed plot_size directly
  scale_size_identity(
    guide = "legend", 
    name = "Significance",
    breaks = c(1, 3, 5),
    labels = c("p >= 0.05", "p < 0.01", "p < 1e-10") # these labels need to be adjusted to the actual data
  ) +
  
  theme_bw(base_size = 16) +
  labs(
    x = "Target-relative position",
    y = ""
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1 ,color = "black"),
    axis.text.y = element_text(color = "black"),
    panel.grid.minor = element_blank()
  )

ggsave("outplots/fig3/fig3_a.loc.tg.annot.dotplot.pdf",p,width=6,height=8,dpi=300)


```

# =====================================================
## sup3_a Genomic region annotation with ChIPseeker
```{r}
library(ChIPseeker)

library(EnsDb.Hsapiens.v75)
#library(org.Hs.eg.db)
peaklist <- list(
    Cluster1="output/peaks/roi_cluster1.bed",
    Cluster2="output/peaks/roi_cluster2.bed",
    Cluster3="output/peaks/roi_cluster3.bed",
    Cluster4="output/peaks/roi_cluster4.bed",
    Cluster5="output/peaks/roi_cluster5.bed",
    Cluster6="output/peaks/roi_cluster6.bed",
    Cluster7="output/peaks/roi_cluster7.bed")
peakAnnoList <- lapply(peaklist, annotatePeak, TxDb=EnsDb.Hsapiens.v75,
                       tssRegion=c(-2000, 2000), verbose=FALSE)
p <- plotAnnoBar(peakAnnoList,cex.axis =5)
p2 <- p +
  theme_bw(base_size = 16) +
  labs(title = "",
       x = "",
       y = "Percentage") +
  theme(
    axis.text.x = element_text(size = 16,color="black"),
    axis.text.y = element_text(size = 16,color="black"),
    axis.title.x = element_text(size = 16,color="black"),
    axis.title.y = element_text(size = 16,color="black"),
    legend.text = element_text(size = 16,color="black"),
    legend.title = element_text(size = 18,color="black")
  )

print(p2)
ggsave(file = "outplots/fig3/sup3_a.genome.location.annot.pdf",p2,width=10,height=8,dpi=300)


```


# =====================================================
## sup3_b Ridge plot of the distance between each cluster and its target genes
```{r}

clu.distance.list <- list()
for (i in 1:7){
    clu.tg <- fread(file = paste0("output/peaks/cluster_tg/cluster",i,".rois.targetGene.tsv.tsv"))
    clu.tg <- clu.tg[clu.tg$final_type != "intergenic_far",]
    clu.tg$center <- (clu.tg$start + clu.tg$end)/2
    clu.tg$tss <- ifelse(ref[match(clu.tg$target_gene_id,ref$gene_id),"strand"] == "+",ref[match(clu.tg$target_gene_id,ref$gene_id),"start"],ref[match(clu.tg$target_gene_id,ref$gene_id),"end"])
    clu.tg$true_distance <- abs(clu.tg$center-clu.tg$tss)
    hist(log10(clu.tg$true_distance+1))
    clu.distance.list[[i]] <- sqrt(clu.tg$true_distance)
}

library(data.table)
library(ggplot2)

## 1. Convert the list into a long table
plot.df <- rbindlist(lapply(seq_along(clu.distance.list), function(i){
    data.table(
        cluster = paste0("cluster", i),
        log_distance = clu.distance.list[[i]]
    )
}))

## 2. Set the order: cluster1 to cluster7 from top to bottom
plot.df$cluster <- factor(plot.df$cluster, levels = paste0("cluster", 1:7))

## 3. Compute the sample size of each cluster (optional)
n.df <- plot.df[, .(n = .N), by = cluster]
n.df[, cluster_lab := paste0(cluster, "  (n=", n, ")")]

plot.df <- merge(plot.df, n.df[, .(cluster, cluster_lab)], by = "cluster", all.x = TRUE)
plot.df$cluster_lab <- factor(plot.df$cluster_lab, levels = n.df$cluster_lab[match(paste0("cluster", 1:7), n.df$cluster)])

## 4. Compute a common y axis upper limit first so that all panels match
dens_list <- lapply(split(plot.df$log_distance, plot.df$cluster), density)
ymax <- max(sapply(dens_list, function(x) max(x$y))) * 1.05


my_colors <- c(
  "cluster1" = "#084594",
  "cluster2" = "#2171b5",
  "cluster3" = "#6baed6",
  "cluster4" = "#bdd7e7",
  "cluster5" = "#b10026",
  "cluster6" = "#e31a1c",
  "cluster7" = "#feb24c"
)

## 5. Plot
p <- ggplot(plot.df, aes(x = log_distance, color = cluster,fill = cluster)) +
    geom_density(alpha = 0.3, linewidth = 0.8) +
    facet_wrap(~ cluster_lab, ncol = 1, scales = "fixed", strip.position = "left") +
    scale_color_manual(values = my_colors) +
    scale_fill_manual(values = my_colors) +
    scale_x_continuous(
        breaks = sqrt(c(0, 100, 1000, 10000, 50000, 100000, 200000)),
        labels = c("0", "100 bp", "1 kb", "10 kb", "50 kb", "100 kb", "200 kb"),
        expand = c(0, 0)
    ) +
    coord_cartesian(xlim = c(0, sqrt(200000)), ylim = c(0, ymax)) +
    labs(
        x = "Distance from ROI center to target gene TSS",
        y = "Density"
    ) +
    theme_bw(base_size = 16) +
    theme(
        panel.border = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        strip.background = element_blank(),
        strip.placement = "outside",
        strip.text.y.left = element_text(angle = 0, hjust = 1, size = 16, face = "bold"),
        axis.text = element_text(size = 16, color = "black"),
        panel.spacing = unit(0.15, "lines"),
        legend.position = "none"
    )

print(p)

ggsave("outplots/fig3/sup3_b.distance.pdf", p, width = 10, height = 8)


```




# =====================================================
## fig3_b Distribution of the number of target genes per cluster
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)

all.rois.tg <- data.frame()
for (i in 1:7){
    input.tg <- read.table(file = paste0("output/peaks/cluster_tg/cluster",i,".rois.targetGene.tsv.tsv"),header = T)
    input.tg$cluster <- paste0('cluster',i)
    all.rois.tg <- rbind(all.rois.tg,input.tg)

}
roi.proximal.tg <- all.rois.tg
colnames(roi.proximal.tg) <- c("chr","start","end","peak","id","type","distance","cluster")


library(tximport)
library(DESeq2)

period <- c("MSC","ADI-7d","ADI-15d")
sample_list <- c("MSC_rep1","MSC_rep2","MSC_rep3","ADI-7d_rep1","ADI-7d_rep2","ADI-7d_rep3","ADI-15d_rep1","ADI-15d_rep2","ADI-15d_rep3")

# Define genes that are expressed
tg.keeped <- setdiff(ref$gene_id,rownames(all.tpm[all.tpm$MSC < 1 & all.tpm$ADI_7d < 1 & all.tpm$ADI_15d < 1,]))

files <- paste0('~/Project/19-NG/RNA-seq/out/',sample_list,'.genes.results')
txi <- tximport(files, type="rsem")

# Deal with non-positive lengths (0/NA/negative)
txi$length[!is.finite(txi$length) | txi$length <= 0] <- 1
sampleTable <- data.frame(
  row.names = names(files),
  condition = c("MSC","MSC","MSC","ADI-7d","ADI-7d","ADI-7d","ADI-15d","ADI-15d","ADI-15d")
)
sampleTable$condition <- factor(sampleTable$condition)

log2FC_cutoff <- 0.5
padj_cutoff   <- 0.05

dds <- DESeqDataSetFromTximport(txi, sampleTable, ~ condition)
dds <- DESeq(dds)
res <- results(dds)
res <- results(dds,contrast = c("condition", "ADI-15d","MSC")) %>% as.data.frame %>% na.omit()

res1 <- results(dds,contrast = c("condition", "ADI-7d","MSC")) %>% as.data.frame %>% na.omit()
res2 <- results(dds,contrast = c("condition", "ADI-15d","ADI-7d")) %>% as.data.frame %>% na.omit()

res1 <- res1[rownames(res1) %in% tg.keeped,]
res1$change <- "NotSig"
res1$change[res1$log2FoldChange >=  log2FC_cutoff & res1$padj < padj_cutoff] <- "Up"
res1$change[res1$log2FoldChange <= -log2FC_cutoff & res1$padj < padj_cutoff] <- "Down"

res2 <- res2[tg.keeped,]
res2$change <- "NotSig"
res2$change[res2$log2FoldChange >=  log2FC_cutoff & res2$padj < padj_cutoff] <- "Up"
res2$change[res2$log2FoldChange <= -log2FC_cutoff & res2$padj < padj_cutoff] <- "Down"


all.clusters.tg.res.annot <- data.frame()
for (i in 1:7){
    input <- roi.proximal.tg[roi.proximal.tg$cluster == paste0('cluster',i),]
    input.res1 <- res1[rownames(res1) %in% input$id,]
    input.res2 <- res2[rownames(res2) %in% input$id,]
    out1 <- c(table(input.res1$change)[1],table(input.res1$change)[2],table(input.res1$change)[3],length(unique(input$id))-length(intersect(input$id,tg.keeped)))/length(unique(input$id))
    out2 <- c(table(input.res2$change)[1],table(input.res2$change)[2],table(input.res2$change)[3],length(unique(input$id))-length(intersect(input$id,tg.keeped)))/length(unique(input$id))
    output <- data.frame(
        cluster = paste0('cluster',i),
        category= c("d7-d0","d14-d7"),
        repressed = c(out1[1],out2[1]),
        unchanged = c(out1[2],out2[2]),
        induced = c(out1[3],out2[3]),
        silent = c(out1[4],out2[4])
    )
    all.clusters.tg.res.annot <- rbind(all.clusters.tg.res.annot,output)
}


# 2. Data cleaning and factor ordering
plot_data <- all.clusters.tg.res.annot %>%
  pivot_longer(cols = c(repressed, unchanged, induced, silent), 
               names_to = "fate", values_to = "proportion")

# Set the Fate order (left to right: Induced -> Unchanged -> Repressed -> Silent)
plot_data$fate <- factor(plot_data$fate, 
                         levels = rev(c("induced", "repressed", "unchanged", "silent")))

# Set the Cluster order (so that the plot runs 1-7 from top to bottom)
plot_data$cluster <- factor(plot_data$cluster, 
                            levels = rev(paste0("cluster", 1:7)))

# Set the Category order
plot_data$category <- factor(plot_data$category, levels = c("d7-d0", "d14-d7"))
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
# 3. Plot
p <- ggplot(plot_data, aes(x = cluster, y = proportion, fill = fate)) +
  # Draw the bars and add a thin white border for a more refined look
  geom_bar(stat = "identity", position = "stack", width = 0.7, color = "white", linewidth = 0.2) +
  # Flip the coordinate axes
  coord_flip() +
  # Facets: stacked vertically without a shared Y axis scale (to keep them independent)
  facet_grid(category ~ ., scales = "free_y", space = "free_y") +
  # Colour scheme
  scale_fill_manual(values = c(
    "induced"   = "#E64B35FF", # coral red (NPG)
    "unchanged" = "#F0F0F0",   # very light grey
    "repressed" = "#4DBBD5FF", # greyish blue (NPG)
    "silent"    = "#3C3C3CFF"  # dark charcoal grey
  )) +
  # Format the Y axis as percentages
  scale_y_continuous(labels = percent_format(), expand = c(0, 0)) +
  # Theme styling
  theme_minimal(base_size = 16) +
  theme(
    # Remove the background grid
    panel.grid = element_blank(),
    # Style the facet strips
    strip.background = element_rect(fill = "gray95", color = NA),
    strip.text = element_text(face = "bold"),
    # Axis text
    axis.text = element_text(color = "black"),
    axis.title = element_blank(),
    # Legend position
    legend.position = "right",
    legend.title = element_blank(),
    # Increase the spacing between panels
    panel.spacing = unit(1, "lines"),
    plot.margin = margin(10, 20, 10, 10)
  )

# 4. Output
print(p)
ggsave("outplots/fig3/fig3_b.Gene_Fate_Proportion.pdf", p, width = 7, height = 8,dpi=300)

```



# =====================================================
## fig3_c Boxplot of target gene expression per cluster
```{r}
tg.keeped <- all.tpm[all.tpm$MSC >=1 | all.tpm$ADI_7d >=1 | all.tpm$ADI_15d >=1,] %>% row.names()
tg.keeped <- intersect(tg.keeped,ref$gene_id)

all.tpm.df <- data.frame()
for (i in 1:7){
    input <- roi.proximal.tg[roi.proximal.tg$cluster == paste0('cluster',i),"id"]
    input <- intersect(input,tg.keeped)
    out <- log2(all.tpm[input,]+1)
    output <- data.frame(
        cluster=rep(paste0('cluster',i),3*nrow(out)),
        category=c(rep('MSC',nrow(out)),rep('ADI-7d',nrow(out)),rep('ADI-15d',nrow(out))),
        tpm=c(out$MSC,out$ADI_7d,out$ADI_15d)
    )
    print(c(
        t.test(out$MSC,out$ADI_7d,paired = T)$p.value,
        t.test(out$MSC,out$ADI_15d,paired = T)$p.value,
        t.test(out$ADI_7d,out$ADI_15d,paired = T)$p.value
    ))
    all.tpm.df <- rbind(all.tpm.df,output)
}

data <- all.tpm.df
data$category <- factor(data$category,c("MSC","ADI-7d","ADI-15d"))
data$cluster <- factor(data$cluster,paste0('cluster',1:7))

data$row_group <- ifelse(data$cluster %in% c("cluster1", "cluster2", "cluster3", "cluster4"), "Group A", "Group B")

# If cluster is numeric 1-7, use instead:
# data$row_group <- ifelse(data$cluster <= 4, "Group A", "Group B")

# 2. Plot
p <- ggplot(data, aes(x = cluster, y = tpm, fill = category)) +
  stat_boxplot(geom = "errorbar",
               width=0.4,
               position = position_dodge(0.8)) +
  geom_boxplot(position = position_dodge(width = 0.8),
               outlier.shape = NA) +  
  scale_fill_manual(values = c("#4b92c3", "#56b356" , "#de5253")) + 
  coord_cartesian(ylim = c(min(data$tpm)+0.5, max(data$tpm))) +
  labs(y = "log(Target Gene TPM +1)", x = " ", fill = "Class") +
  theme_bw(base_size = 16) +
  
  # --- Key modification: faceting ---
  facet_wrap(~row_group, ncol = 1, scales = "free_x") + 
  
  theme(
    axis.text.x = element_text(colour = "black"),  
    axis.text.y = element_text(colour = "black"),  
    axis.title.y = element_text(face = "bold"),   
    legend.position = "right",
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank(), 
    panel.border = element_rect(color = "black"),
    
    # Hide the facet strip background and text (if the "Group A" / "Group B" bars are not wanted)
    strip.background = element_blank(),
    strip.text = element_blank()
  )

# Save the figure (note: faceting may require a larger height, e.g. 7 or 8)
ggsave(p, file = "outplots/fig3/fig3_d.all.clu.tg.tpm.pdf", width = 8, height = 8, dpi = 300)

```



```{r}
library(clusterProfiler)
library(org.Hs.eg.db)

cluP.go <- function(input.tg,ref.tg,keyType = "SYMBOL",ont,qvalue=0.05){
    ego <- enrichGO(
        gene = input.tg,
        universe = ref.tg,
        OrgDb = org.Hs.eg.db,
        keyType = keyType,
        ont = ont,
        pAdjustMethod = "BH",
        pvalueCutoff = 0.05,
        qvalueCutoff = qvalue,
        minGSSize = 10,
        maxGSSize = 300,
        readable = TRUE
    )
    if (nrow(ego) == 0){
        return(data.frame(ID = NA,Description = NA,qvalue = NA))
    }
    else{
        ego2 <- as.data.frame(ego)
        ego2 <- data.frame(ID = ego2$ID,Description = ego2$Description,qvalue = ego2$qvalue)
    }

}


up.tg <- res1[res1$change == "Up",] %>% rownames()
down.tg <- res1[res1$change == "Down",] %>% rownames

up.tg <- ref[ref$gene_id %in% up.tg,"gene_name"]
down.tg <- ref[ref$gene_id %in% down.tg,"gene_name"]

all.go.up <- data.frame()
all.go.down <- data.frame()
for (i in 1:7){
    clu.tg <- intersect(roi.proximal.tg[roi.proximal.tg$cluster == paste0('cluster',i),"id"],tg.keeped)
    clu.tg <- ref[ref$gene_id %in% clu.tg,"gene_name"]
    clu.up.tg <- intersect(clu.tg,up.tg)
    clu.down.tg <- intersect(clu.tg,down.tg)
    universe <- ref[ref$gene_id %in% tg.keeped,"gene_name"]
    out1=cluP.go(input.tg = clu.up.tg,ref.tg = universe,ont = "BP")
    out2=cluP.go(input.tg = clu.up.tg,ref.tg = universe,ont = "MF")
    out3=cluP.go(input.tg = clu.up.tg,ref.tg = universe,ont = "CC")
    output <- rbind(out1,out2,out3)
    output$cluster <- paste0('cluster',i)
    all.go.up <- rbind(all.go.up,output)

    out1=cluP.go(input.tg = clu.down.tg,ref.tg = universe,ont = "BP")
    out2=cluP.go(input.tg = clu.down.tg,ref.tg = universe,ont = "MF")
    out3=cluP.go(input.tg = clu.down.tg,ref.tg = universe,ont = "CC")
    output <- rbind(out1,out2,out3)
    output$cluster <- paste0('cluster',i)
    all.go.down <- rbind(all.go.down,output)
}
 
all.go.up$logqvalue <- -log10(all.go.up$qvalue)
all.go.down$logqvalue <- -log10(all.go.down$qvalue)
all.go.up <- na.omit(all.go.up)
all.go.up.mat <- data.frame(matrix(nrow=length(unique(all.go.up$Description)),ncol=7),row.names=unique(all.go.up$Description))
colnames(all.go.up.mat) <- paste0('cluster',1:7)
for (i in 1:length(unique(all.go.up$Description))){
    input <- all.go.up[all.go.up$Description == unique(all.go.up$Description)[i],]
    all.go.up.mat[i,input$cluster] <- input$logqvalue
}
fwrite(all.go.up.mat,file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/WebGestaltR/d7-d0.all.go.up.mat.tsv",sep = "\t",row.names = T)

all.go.down <- na.omit(all.go.down)
all.go.down.mat <- data.frame(matrix(nrow=length(unique(all.go.down$Description)),ncol=7),row.names=unique(all.go.down$Description))
colnames(all.go.down.mat) <- paste0('cluster',1:7)
for (i in 1:length(unique(all.go.down$Description))){
    input <- all.go.down[all.go.down$Description == unique(all.go.down$Description)[i],]
    all.go.down.mat[i,input$cluster] <- input$logqvalue
}
fwrite(all.go.down.mat,file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/WebGestaltR/d7-d0.all.go.down.mat.tsv",sep = "\t",row.names = T)


up.tg <- res2[res2$change == "Up",] %>% rownames()
down.tg <- res2[res2$change == "Down",] %>% rownames

up.tg <- ref[ref$gene_id %in% up.tg,"gene_name"]
down.tg <- ref[ref$gene_id %in% down.tg,"gene_name"]

all.go.up <- data.frame()
all.go.down <- data.frame()
for (i in 1:7){
    clu.tg <- intersect(roi.proximal.tg[roi.proximal.tg$cluster == paste0('cluster',i),"id"],tg.keeped)
    clu.tg <- ref[ref$gene_id %in% clu.tg,"gene_name"]
    clu.up.tg <- intersect(clu.tg,up.tg)
    clu.down.tg <- intersect(clu.tg,down.tg)
    universe <- ref[ref$gene_id %in% tg.keeped,"gene_name"]
    out1=cluP.go(input.tg = clu.up.tg,ref.tg = universe,ont = "BP")
    out2=cluP.go(input.tg = clu.up.tg,ref.tg = universe,ont = "MF")
    out3=cluP.go(input.tg = clu.up.tg,ref.tg = universe,ont = "CC")
    output <- rbind(out1,out2,out3)
    output$cluster <- paste0('cluster',i)
    all.go.up <- rbind(all.go.up,output)

    out1=cluP.go(input.tg = clu.down.tg,ref.tg = universe,ont = "BP")
    out2=cluP.go(input.tg = clu.down.tg,ref.tg = universe,ont = "MF")
    out3=cluP.go(input.tg = clu.down.tg,ref.tg = universe,ont = "CC")
    output <- rbind(out1,out2,out3)
    output$cluster <- paste0('cluster',i)
    all.go.down <- rbind(all.go.down,output)
}
 
all.go.up$logqvalue <- -log10(all.go.up$qvalue)
all.go.down$logqvalue <- -log10(all.go.down$qvalue)
all.go.up <- na.omit(all.go.up)
all.go.up.mat <- data.frame(matrix(nrow=length(unique(all.go.up$Description)),ncol=7),row.names=unique(all.go.up$Description))
colnames(all.go.up.mat) <- paste0('cluster',1:7)
for (i in 1:length(unique(all.go.up$Description))){
    input <- all.go.up[all.go.up$Description == unique(all.go.up$Description)[i],]
    all.go.up.mat[i,input$cluster] <- input$logqvalue
}
fwrite(all.go.up.mat,file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/WebGestaltR/d14-d7.all.go.up.mat.tsv",sep = "\t",row.names = T)

all.go.down <- na.omit(all.go.down)
all.go.down.mat <- data.frame(matrix(nrow=length(unique(all.go.down$Description)),ncol=7),row.names=unique(all.go.down$Description))
colnames(all.go.down.mat) <- paste0('cluster',1:7)
for (i in 1:length(unique(all.go.down$Description))){
    input <- all.go.down[all.go.down$Description == unique(all.go.down$Description)[i],]
    all.go.down.mat[i,input$cluster] <- input$logqvalue
}
fwrite(all.go.down.mat,file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/WebGestaltR/d14-d7.all.go.down.mat.tsv",sep = "\t",row.names = T)




```



## sup3_c. Expression differences of target genes relative to control
```{r}
all.clu.tpm <- data.frame()
for (i in 1:7){
    l1 <- fread(file = paste0('output/peaks/cluster_tg/cluster',i,'.rois.targetGene.tsv.tsv'))
    l2 <- fread(file = paste0('output/peaks/cluster_tg/cluster',i,'.bg.targetGene.tsv'))
    var1 <- dispersions(dds)[match(unique(l1$target_gene_id),rownames(dds))]
    var2 <- dispersions(dds)[match(unique(l2$target_gene_id),rownames(dds))]

    tpm1 <- log2(all.tpm[unique(l1$target_gene_id),]+1)
    tpm2 <- log2(all.tpm[unique(l2$target_gene_id),]+1)
    output <- data.frame(
        cluster = paste0('cluster',i),
        category = c(rep("input",nrow(tpm1)),rep("control",nrow(tpm2)),rep("input",nrow(tpm1)),rep("control",nrow(tpm2)),rep("input",nrow(tpm1)),rep("control",nrow(tpm2))),
        period = c(rep("d0",nrow(tpm1)+nrow(tpm2)),rep("d7",nrow(tpm1)+nrow(tpm2)),rep("d14",nrow(tpm1)+nrow(tpm2))),
        tpm=c(tpm1$MSC,tpm2$MSC,tpm1$ADI_7d,tpm2$ADI_7d,tpm1$ADI_15d,tpm2$ADI_15d)
    )
    all.clu.tpm <- rbind(all.clu.tpm,output)
}


library(ggplot2)
library(dplyr)

# Set the factor order
all.clu.tpm <- all.clu.tpm %>%
  mutate(
    cluster = factor(cluster, levels = paste0("cluster", 1:7)),
    period = factor(period, levels = c("d0", "d7", "d14")),
    category = factor(category, levels = c("input", "control")),
    fill_group = case_when(
      category == "control" ~ "control",
      category == "input" & period == "d0" ~ "input_d0",
      category == "input" & period == "d7" ~ "input_d7",
      category == "input" & period == "d14" ~ "input_d14"
    )
  )

all.clu.tpm$fill_group <- factor(
  all.clu.tpm$fill_group,
  levels = c("input_d0", "input_d7", "input_d14", "control")
)
sig_df <- all.clu.tpm %>%
  group_by(cluster, period) %>%
  summarise(
    y = max(tpm, na.rm = TRUE) * 0.9,
    .groups = "drop"
  ) %>%
  mutate(label = "***")


p <- ggplot(all.clu.tpm, aes(x = period, y = tpm, fill = fill_group)) +
  geom_boxplot(
    aes(group = interaction(period, category)),
    position = position_dodge(width = 0.8),
    width = 0.7,
    outlier.size = 0.25,
    outlier.alpha = 0.25
  ) +
  facet_wrap(~ cluster, ncol = 4, scales = "free_y") +
  scale_fill_manual(
    values = c(
      "input_d0" = "#1F77B4",
      "input_d7" = "#2CA02C",
      "input_d14" = "#D62728",
      "control" = "grey70"
    ),
    breaks = c("input_d0", "input_d7", "input_d14", "control"),
    labels = c("Input d0", "Input d7", "Input d14", "Control")
  ) +
  labs(
    x = "Period",
    y = "TPM",
    fill = NULL
  ) +
  theme_bw(base_size = 16) +
  theme(
    strip.background = element_rect(fill = "white", colour = "black"),
    strip.text = element_text(face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    legend.position = "right"
  )

p2 <- p +
  geom_segment(
    data = sig_df,
    aes(
      x = as.numeric(period) - 0.2,
      xend = as.numeric(period) + 0.2,
      y = y,
      yend = y
    ),
    inherit.aes = FALSE
  ) +
  geom_text(
    data = sig_df,
    aes(
      x = period,
      y = y,
      label = label
    ),
    vjust = -0.3,
    size = 4,
    inherit.aes = FALSE
  )

ggsave(file = "outplots/fig3/sup3_c.all.tpm.boxplots.pdf",p2,width=10,height=8)


```





```{r}
############################
## 0. Packages
############################
suppressPackageStartupMessages({
  library(clusterProfiler)
  library(org.Hs.eg.db)
  library(GOSemSim)
  library(dplyr)
  library(tidyr)
  library(tibble)
  library(purrr)
  library(pheatmap)
})

############################
## 1. Parameters
############################
outdir <- "GO_global_module_results"
dir.create(outdir, showWarnings = FALSE, recursive = TRUE)

keyType_use <- "SYMBOL"
q_cutoff <- 0.05
p_cutoff <- 0.05
minGSSize_use <- 10
maxGSSize_use <- 300

# GO semantic similarity cutoff
# larger -> stricter merge; smaller -> more aggressive merge
sim_cut_BP <- 0.5
sim_cut_MF <- 0.5
sim_cut_CC <- 0.5

cluster_ids <- paste0("cluster", 1:7)

############################
## 2. Helper functions
############################

safe_max <- function(x) {
  if (length(x) == 0 || all(is.na(x))) NA_real_ else max(x, na.rm = TRUE)
}

safe_mean <- function(x) {
  if (length(x) == 0 || all(is.na(x))) NA_real_ else mean(x, na.rm = TRUE)
}

cluP.go <- function(input.tg, ref.tg, keyType = "SYMBOL", ont,
                    qvalue = 0.05,
                    pvalue = 0.05,
                    minGSSize = 10,
                    maxGSSize = 300) {

  input.tg <- unique(input.tg)
  ref.tg <- unique(ref.tg)

  if (length(input.tg) < minGSSize) {
    return(data.frame(
      ID = NA_character_,
      Description = NA_character_,
      Ontology = ont,
      qvalue = NA_real_,
      stringsAsFactors = FALSE
    ))
  }

  ego <- tryCatch({
    clusterProfiler::enrichGO(
      gene = input.tg,
      universe = ref.tg,
      OrgDb = org.Hs.eg.db::org.Hs.eg.db,
      keyType = keyType,
      ont = ont,
      pAdjustMethod = "BH",
      pvalueCutoff = pvalue,
      qvalueCutoff = qvalue,
      minGSSize = minGSSize,
      maxGSSize = maxGSSize,
      readable = TRUE
    )
  }, error = function(e) {
    NULL
  })

  if (is.null(ego)) {
    return(data.frame(
      ID = NA_character_,
      Description = NA_character_,
      Ontology = ont,
      qvalue = NA_real_,
      stringsAsFactors = FALSE
    ))
  }

  ego.df <- as.data.frame(ego)

  if (nrow(ego.df) == 0) {
    return(data.frame(
      ID = NA_character_,
      Description = NA_character_,
      Ontology = ont,
      qvalue = NA_real_,
      stringsAsFactors = FALSE
    ))
  }

  data.frame(
    ID = ego.df$ID,
    Description = ego.df$Description,
    Ontology = ont,
    qvalue = ego.df$qvalue,
    stringsAsFactors = FALSE
  )
}

run_go_for_clusters <- function(direction = c("up", "down"),
                                res1,
                                ref,
                                roi.proximal.tg,
                                tg.keeped,
                                cluster_ids,
                                keyType = "SYMBOL",
                                q_cutoff = 0.05,
                                p_cutoff = 0.05,
                                minGSSize = 10,
                                maxGSSize = 300) {
  direction <- match.arg(direction)

  if (direction == "up") {
    target_gene_id <- rownames(res1[res1$change == "Up", , drop = FALSE])
  } else {
    target_gene_id <- rownames(res1[res1$change == "Down", , drop = FALSE])
  }

  target_gene_name <- unique(ref[ref$gene_id %in% target_gene_id, "gene_name"])
  universe <- unique(ref[ref$gene_id %in% tg.keeped, "gene_name"])

  all.go <- data.frame()

  for (clu.id in cluster_ids) {
    message("Running ", direction, " GO for ", clu.id, "...")

    clu.tg.id <- intersect(
      roi.proximal.tg[roi.proximal.tg$cluster == clu.id, "id"],
      tg.keeped
    )

    clu.tg <- unique(ref[ref$gene_id %in% clu.tg.id, "gene_name"])
    clu.target.tg <- intersect(clu.tg, target_gene_name)

    out_bp <- cluP.go(
      input.tg = clu.target.tg,
      ref.tg = universe,
      keyType = keyType,
      ont = "BP",
      qvalue = q_cutoff,
      pvalue = p_cutoff,
      minGSSize = minGSSize,
      maxGSSize = maxGSSize
    )

    out_mf <- cluP.go(
      input.tg = clu.target.tg,
      ref.tg = universe,
      keyType = keyType,
      ont = "MF",
      qvalue = q_cutoff,
      pvalue = p_cutoff,
      minGSSize = minGSSize,
      maxGSSize = maxGSSize
    )

    out_cc <- cluP.go(
      input.tg = clu.target.tg,
      ref.tg = universe,
      keyType = keyType,
      ont = "CC",
      qvalue = q_cutoff,
      pvalue = p_cutoff,
      minGSSize = minGSSize,
      maxGSSize = maxGSSize
    )

    output <- dplyr::bind_rows(out_bp, out_mf, out_cc) %>%
      dplyr::mutate(cluster = clu.id)

    all.go <- dplyr::bind_rows(all.go, output)
  }

  all.go
}

prepare_go_df <- function(all.go.df) {
  all.go.df %>%
    dplyr::filter(!is.na(ID), !is.na(Description), !is.na(qvalue)) %>%
    dplyr::distinct(cluster, ID, Description, Ontology, .keep_all = TRUE) %>%
    dplyr::mutate(score = -log10(qvalue))
}

build_go_modules <- function(df, ont = "BP", sim_cut = 0.5) {
  subdf <- df %>%
    dplyr::filter(Ontology == ont) %>%
    dplyr::distinct(ID, Description, Ontology)

  if (nrow(subdf) == 0) {
    return(data.frame())
  }

  if (nrow(subdf) == 1) {
    return(
      subdf %>%
        dplyr::mutate(module = paste0(ont, "_M1")) %>%
        dplyr::select(ID, Description, Ontology, module)
    )
  }

  semData <- GOSemSim::godata(
    annoDb = "org.Hs.eg.db",
    ont = ont
  )

  sim_mat <- GOSemSim::mgoSim(
    GO1 = subdf$ID,
    GO2 = subdf$ID,
    semData = semData,
    measure = "Wang",
    combine = NULL
  )

  sim_mat <- as.matrix(sim_mat)
  rownames(sim_mat) <- subdf$ID
  colnames(sim_mat) <- subdf$ID
  sim_mat[is.na(sim_mat)] <- 0
  diag(sim_mat) <- 1

  dist_mat <- stats::as.dist(1 - sim_mat)
  hc <- stats::hclust(dist_mat, method = "average")
  module_id <- stats::cutree(hc, h = 1 - sim_cut)

  data.frame(
    ID = names(module_id),
    module = paste0(ont, "_M", module_id),
    stringsAsFactors = FALSE
  ) %>%
    dplyr::left_join(subdf, by = "ID") %>%
    dplyr::select(ID, Description, Ontology, module)
}

pick_module_representative <- function(go_df, term2module_df) {
  go_df %>%
    dplyr::left_join(term2module_df, by = c("ID", "Description", "Ontology")) %>%
    dplyr::filter(!is.na(module)) %>%
    dplyr::group_by(module, ID, Description, Ontology) %>%
    dplyr::summarise(
      n_cluster = dplyr::n_distinct(cluster),
      mean_score = safe_mean(score),
      max_score = safe_max(score),
      .groups = "drop"
    ) %>%
    dplyr::arrange(
      module,
      dplyr::desc(n_cluster),
      dplyr::desc(mean_score),
      dplyr::desc(max_score),
      Description
    ) %>%
    dplyr::group_by(module) %>%
    dplyr::slice(1) %>%
    dplyr::ungroup() %>%
    dplyr::rename(module_label = Description)
}

build_term_level_matrix <- function(go_df) {
  go_df %>%
    dplyr::select(ID, Description, Ontology, cluster, score) %>%
    tidyr::pivot_wider(
      id_cols = c(ID, Description, Ontology),
      names_from = cluster,
      values_from = score
    ) %>%
    dplyr::arrange(Ontology, Description)
}

build_module_matrix <- function(go_df, term2module_df, module_label_df) {
  go_df %>%
    dplyr::left_join(term2module_df, by = c("ID", "Description", "Ontology")) %>%
    dplyr::filter(!is.na(module)) %>%
    dplyr::group_by(module, cluster) %>%
    dplyr::summarise(score = safe_max(score), .groups = "drop") %>%
    dplyr::left_join(
      module_label_df %>% dplyr::select(module, module_label, Ontology),
      by = "module"
    ) %>%
    dplyr::distinct() %>%
    tidyr::pivot_wider(
      id_cols = c(module, module_label, Ontology),
      names_from = cluster,
      values_from = score
    ) %>%
    dplyr::arrange(Ontology, module_label)
}

plot_module_heatmap <- function(module_mat_df, outfile, cluster_ids, main = "GO module heatmap") {
  if (nrow(module_mat_df) == 0) return(NULL)

  mat <- module_mat_df %>%
    dplyr::select(dplyr::any_of(cluster_ids)) %>%
    as.data.frame()

  rownames(mat) <- paste(module_mat_df$Ontology, module_mat_df$module_label, sep = " | ")
  mat <- as.matrix(mat)
  mat[is.na(mat)] <- 0

  if (nrow(mat) < 2 || ncol(mat) < 2) return(NULL)

  grDevices::pdf(outfile, width = 8, height = max(6, nrow(mat) * 0.18))
  pheatmap::pheatmap(
    mat,
    scale = "row",
    clustering_method = "ward.D2",
    border_color = NA,
    fontsize_row = 7,
    fontsize_col = 10,
    main = main
  )
  grDevices::dev.off()
}

build_module_members <- function(term2module_df) {
  term2module_df %>%
    dplyr::group_by(module, Ontology) %>%
    dplyr::summarise(
      member_terms = paste(Description, collapse = " ; "),
      member_ids = paste(ID, collapse = " ; "),
      n_terms = dplyr::n(),
      .groups = "drop"
    ) %>%
    dplyr::arrange(Ontology, module)
}

get_module_presence <- function(module_mat_df, cluster_ids) {
  module_mat_df %>%
    dplyr::mutate(
      n_cluster_present = rowSums(
        !is.na(as.matrix(dplyr::select(., dplyr::any_of(cluster_ids))))
      )
    )
}

run_global_module_pipeline <- function(all.go.df,
                                       sim_cut_BP = 0.5,
                                       sim_cut_MF = 0.5,
                                       sim_cut_CC = 0.5) {
  go_df <- prepare_go_df(all.go.df)

  term_mat <- build_term_level_matrix(go_df)

  term2module_bp <- build_go_modules(go_df, ont = "BP", sim_cut = sim_cut_BP)
  term2module_mf <- build_go_modules(go_df, ont = "MF", sim_cut = sim_cut_MF)
  term2module_cc <- build_go_modules(go_df, ont = "CC", sim_cut = sim_cut_CC)

  term2module <- dplyr::bind_rows(term2module_bp, term2module_mf, term2module_cc)

  module_label <- pick_module_representative(go_df, term2module)

  module_mat <- build_module_matrix(go_df, term2module, module_label)

  module_members <- build_module_members(term2module)

  list(
    go_df = go_df,
    term_mat = term_mat,
    term2module = term2module,
    module_label = module_label,
    module_mat = module_mat,
    module_members = module_members
  )
}

############################
## 3. Run GO enrichment
############################
all.go.up <- run_go_for_clusters(
  direction = "up",
  res1 = res1,
  ref = ref,
  roi.proximal.tg = roi.proximal.tg,
  tg.keeped = tg.keeped,
  cluster_ids = cluster_ids,
  keyType = keyType_use,
  q_cutoff = q_cutoff,
  p_cutoff = p_cutoff,
  minGSSize = minGSSize_use,
  maxGSSize = maxGSSize_use
)

all.go.down <- run_go_for_clusters(
  direction = "down",
  res1 = res1,
  ref = ref,
  roi.proximal.tg = roi.proximal.tg,
  tg.keeped = tg.keeped,
  cluster_ids = cluster_ids,
  keyType = keyType_use,
  q_cutoff = q_cutoff,
  p_cutoff = p_cutoff,
  minGSSize = minGSSize_use,
  maxGSSize = maxGSSize_use
)

############################
## 4. Build global GO modules
############################
res.up <- run_global_module_pipeline(
  all.go.df = all.go.up,
  sim_cut_BP = sim_cut_BP,
  sim_cut_MF = sim_cut_MF,
  sim_cut_CC = sim_cut_CC
)

res.down <- run_global_module_pipeline(
  all.go.df = all.go.down,
  sim_cut_BP = sim_cut_BP,
  sim_cut_MF = sim_cut_MF,
  sim_cut_CC = sim_cut_CC
)

############################
## 5. Export tables
############################
write.table(all.go.up,
            file = file.path(outdir, "all.go.up.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(all.go.down,
            file = file.path(outdir, "all.go.down.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.up$go_df,
            file = file.path(outdir, "all.go.up.cleaned.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.down$go_df,
            file = file.path(outdir, "all.go.down.cleaned.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.up$term_mat,
            file = file.path(outdir, "go_term_mat_up.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.down$term_mat,
            file = file.path(outdir, "go_term_mat_down.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.up$term2module,
            file = file.path(outdir, "go_term2module_up.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.down$term2module,
            file = file.path(outdir, "go_term2module_down.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.up$module_label,
            file = file.path(outdir, "go_module_label_up.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.down$module_label,
            file = file.path(outdir, "go_module_label_down.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.up$module_mat,
            file = file.path(outdir, "go_module_mat_up.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(res.down$module_mat,
            file = file.path(outdir, "go_module_mat_down.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

############################
## 6. Plot heatmaps
############################
plot_module_heatmap(
  res.up$module_mat,
  outfile = file.path(outdir, "go_module_heatmap_up.pdf"),
  main = "GO module heatmap (Up)"
)

plot_module_heatmap(
  res.down$module_mat,
  outfile = file.path(outdir, "go_module_heatmap_down.pdf"),
  main = "GO module heatmap (Down)"
)

############################
## 7. Optional: shared/specific modules
############################
get_module_presence <- function(module_mat_df, cluster_ids) {
  module_mat_df %>%
    dplyr::mutate(
      n_cluster_present = rowSums(!is.na(as.matrix(dplyr::select(., dplyr::any_of(cluster_ids)))))
    )
}

up_module_presence <- get_module_presence(res.up$module_mat, cluster_ids)
down_module_presence <- get_module_presence(res.down$module_mat, cluster_ids)

write.table(up_module_presence,
            file = file.path(outdir, "go_module_mat_up.with_presence.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

write.table(down_module_presence,
            file = file.path(outdir, "go_module_mat_down.with_presence.tsv"),
            sep = "\t", quote = FALSE, row.names = FALSE)

message("Done. Results saved in: ", outdir)
```





## sup3_g. Multivariable linear regression plots, in particular the regression for clu567, because the target genes of clu567 are significantly upregulated.
```{r}
period <- c("d0","d7","d15")
period_delta <- c("d0_d7","d7_d15","d0_d15")
feats_coverage <- orig_df[,grep("_coverage",colnames(orig_df),value=T)]
chip <- c("DNase","H3K27ac","H3K27me3","H3K36me3","H3K4me1","H3K4me3","H3K9ac","H3K9me3")


Create_period_df <- function(input.feats.df,input.proximal.tg.df,period_temp){ # this function computes the feature matrix for a single time point
    chip <- c("DNase","H3K27ac","H3K27me3","H3K36me3","H3K4me1","H3K4me3","H3K9ac","H3K9me3")
    input.tg <- unique(input.proximal.tg.df$id)
    output <- data.frame()
    for (j in 1:length(input.tg)){
        tg.tpm <- log2(all.tpm[input.tg[j],which(period == period_temp)]+1)
        peaks <- input.proximal.tg.df[input.proximal.tg.df$id == input.tg[j],"peak"]
        df <- input.feats.df[peaks,]
        out_feats <- data.frame(matrix(nrow=nrow(df),ncol=length(chip)))
        colnames(out_feats) <- chip
        for (k in 1:length(chip)){
            out_feats[,k] <- df[,paste0(chip[k],'_',period_temp,'_coverage')]
        }
        out <- cbind(data.frame(gene_id=input.tg[j],tpm=tg.tpm),out_feats)
        output <- rbind(output,out)
    }
    return(output)

}


Create_delta_df <- function(input.feats.df,input.proximal.tg.df,delta_class){ # this function computes the feature matrix of the difference between two time points
    chip <- c("DNase","H3K27ac","H3K27me3","H3K36me3","H3K4me1","H3K4me3","H3K9ac","H3K9me3")
    input.tg <- unique(input.proximal.tg.df$id)
    period1 <- unlist(strsplit(delta_class,"_"))[1]
    period2 <- unlist(strsplit(delta_class,"_"))[2]
    output <- data.frame()
    for (j in 1:length(input.tg)){
        tg.tpm.delta <- log2(all.tpm[input.tg[j],which(period == period2)]+1)-log2(all.tpm[input.tg[j],which(period == period1)]+1)
        peaks <- input.proximal.tg.df[input.proximal.tg.df$id == input.tg[j],"peak"]
        df <- input.feats.df[peaks,]
        out_feats <- data.frame(matrix(nrow=nrow(df),ncol=length(chip)))
        colnames(out_feats) <- chip
        for (k in 1:length(chip)){
            out_feats[,k] <- df[,paste0(chip[k],'_',period2,'_coverage')]-df[,paste0(chip[k],'_',period1,'_coverage')]
        }
        out <- cbind(data.frame(gene_id=input.tg[j],tpm=tg.tpm.delta),out_feats)
        output <- rbind(output,out)
    }
    return(output)

}


all.delta.period <- data.frame()
{
    input <- roi_info.split[roi_info.split$cluster %in% c(5,6,7),]
    input.feats.df <- feats_coverage[roi_info.split$cluster %in% c(5,6,7),]
    input.proximal.tg.df <- roi.proximal.tg[roi.proximal.tg$cluster %in% paste0('cluster',c(5,6,7)) & roi.proximal.tg$id %in% tg.keeped,]
    output1 <- data.frame()
    output2 <- data.frame()
    for (j in 1:3){
        out.delta.df <- Create_delta_df(input.feats.df,input.proximal.tg.df,period_delta[j])
        ## Examine the distribution of 1s for each chip: a chip whose content is very low (below 0.05) has no explanatory power and should be dropped.
        ratio2 <- sapply(out.delta.df[,-c(1,2)], function(x) mean(x == 0, na.rm = TRUE))
        out.delta.df[,colnames(out.delta.df) %in% names(ratio2[ratio2 >= 0.95])] <- 0
        model2 <- lm(tpm ~ DNase + H3K27ac + H3K27me3 + H3K36me3 + 
                H3K4me1 + H3K4me3 + H3K9ac + H3K9me3, 
                data = out.delta.df)
        coef_matrix2 <- as.data.frame(summary(model2)$coefficients)[-1,]
        #coef_matrix2[,4] <- p.adjust(coef_matrix2[,4])
        output2 <- rbind(output2,data.frame(cluster=paste0('cluster567'),period=period_delta[j],chip=chip,estimate=coef_matrix2[match(chip,rownames(coef_matrix2)),1],se=coef_matrix2[match(chip,rownames(coef_matrix2)),2],pval=coef_matrix2[match(chip,rownames(coef_matrix2)),4]))
     
    }
    all.delta.period <- rbind(all.delta.period,output2)
}


all.delta.period <- data.frame()
{
    input <- roi_info.split[roi_info.split$cluster %in% c(7),]
    input.feats.df <- feats_coverage[roi_info.split$cluster %in% c(7),]
    input.proximal.tg.df <- roi.proximal.tg[roi.proximal.tg$cluster %in% paste0('cluster',c(7)) & roi.proximal.tg$id %in% tg.keeped,]
    output1 <- data.frame()
    output2 <- data.frame()
    for (j in 1:3){
        out.delta.df <- Create_delta_df(input.feats.df,input.proximal.tg.df,period_delta[j])
        ## Examine the distribution of 1s for each chip: a chip whose content is very low (below 0.05) has no explanatory power and should be dropped.
        ratio2 <- sapply(out.delta.df[,-c(1,2)], function(x) mean(x == 0, na.rm = TRUE))
        out.delta.df[,colnames(out.delta.df) %in% names(ratio2[ratio2 >= 0.95])] <- 0
        model2 <- lm(tpm ~ DNase + H3K27ac + H3K27me3 + H3K36me3 + 
                H3K4me1 + H3K4me3 + H3K9ac + H3K9me3, 
                data = out.delta.df)
        coef_matrix2 <- as.data.frame(summary(model2)$coefficients)[-1,]
        #coef_matrix2[,4] <- p.adjust(coef_matrix2[,4])
        output2 <- rbind(output2,data.frame(cluster=paste0('cluster567'),period=period_delta[j],chip=chip,estimate=coef_matrix2[match(chip,rownames(coef_matrix2)),1],se=coef_matrix2[match(chip,rownames(coef_matrix2)),2],pval=coef_matrix2[match(chip,rownames(coef_matrix2)),4]))
     
    }
    all.delta.period <- rbind(all.delta.period,output2)
}




# 1. Load required libraries
library(ggplot2)
library(dplyr)

# 2. Prepare the data
data <- all.delta.period[1:16,]

chip_order <- unique(data$chip)
data <- data %>%
  mutate(
    logP = -log10(pval),
    # Define the significance groups
    group = case_when(
      pval < 0.05 & estimate > 0 ~ "Sig Positive",
      pval < 0.05 & estimate < 0 ~ "Sig Negative",
      TRUE ~ "Non-significant"
    ),
    # Error bar logic: positive values to the right (estimate+se), negative values to the left (estimate-se)
    error_limit = ifelse(estimate >= 0, estimate + se, estimate - se),
    # Keep the original order: reverse the factor levels so that the first entry appears at the top of the plot
    chip = factor(chip, levels = rev(chip_order))
  )

# 4. Define the plotting function
draw_regression_plot <- function(df_sub, title_name) {
  ggplot(df_sub, aes(x = estimate, y = chip, fill = group, alpha = logP)) +
    # Error bar layer: placed beneath the bars
    # xmin and xmax always span estimate and the computed error_limit
    geom_errorbarh(aes(xmin = estimate, xmax = error_limit), 
                   height = 0.2, color = "black", alpha = 1) +
    # Bar chart layer
    geom_col(color = "black", linewidth = 0.5) +
    # Colour mapping
    scale_fill_manual(values = c("Sig Positive" = "#E41A1C",   # red
                                 "Sig Negative" = "#377EB8",   # blue
                                 "Non-significant" = "grey80")) +
    # Alpha mapping (the smaller the P value, the more significant and the darker the colour)
    scale_alpha_continuous(range = c(0.7, 1), guide = "none") +
    # Zero reference line
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    labs(title = "",
         x = "Estimate (Coefficient)",
         y = "",
         fill = "Significance") +
    theme_bw(base_size = 16) +
    theme(
      panel.grid.minor = element_blank(),
      axis.text = element_text(color = "black")
    )
}

# 5. Generate and display the plots
plot_d0_d7 <- draw_regression_plot(filter(data, period == "d0_d7"), "d0_d7")
plot_d7_d15 <- draw_regression_plot(filter(data, period == "d7_d15"), "d7_d15")

print(plot_d0_d7)
print(plot_d7_d15)

ggsave("outplots/fig3/fig3_e.d0-d7.clu567.estimate.barplots.pdf",plot_d0_d7,width=6,height=5,dpi=300)
ggsave("outplots/fig3/fig3_e.d15-d7.clu567.estimate.barplots.pdf",plot_d7_d15,width=6,height=5,dpi=300)

```



```{r}

all.delta.period <- data.frame()
{
    input <- roi_info.split[roi_info.split$cluster %in% c(1,2,3,4),]
    input.feats.df <- feats_coverage[roi_info.split$cluster %in% c(1,2,3,4),]
    input.proximal.tg.df <- roi.proximal.tg[roi.proximal.tg$cluster %in% paste0('cluster',c(1,2,3,4)) & roi.proximal.tg$id %in% tg.keeped,]
    output1 <- data.frame()
    output2 <- data.frame()
    for (j in 1:3){
        out.delta.df <- Create_delta_df(input.feats.df,input.proximal.tg.df,period_delta[j])
        ## Examine the distribution of 1s for each chip: a chip whose content is very low (below 0.05) has no explanatory power and should be dropped.
        ratio2 <- sapply(out.delta.df[,-c(1,2)], function(x) mean(x == 0, na.rm = TRUE))
        out.delta.df[,colnames(out.delta.df) %in% names(ratio2[ratio2 >= 0.95])] <- 0
        model2 <- lm(tpm ~ DNase + H3K27ac + H3K27me3 + H3K36me3 + 
                H3K4me1 + H3K4me3 + H3K9ac + H3K9me3, 
                data = out.delta.df)
        coef_matrix2 <- as.data.frame(summary(model2)$coefficients)[-1,]
        coef_matrix2[,4] <- p.adjust(coef_matrix2[,4])
        output2 <- rbind(output2,data.frame(cluster=paste0('cluster567'),period=period_delta[j],chip=chip,estimate=coef_matrix2[match(chip,rownames(coef_matrix2)),1],se=coef_matrix2[match(chip,rownames(coef_matrix2)),2],pval=coef_matrix2[match(chip,rownames(coef_matrix2)),4]))
     
    }
    all.delta.period <- rbind(all.delta.period,output2)
}




# 1. Load required libraries
library(ggplot2)
library(dplyr)

# 2. Prepare the data
data <- all.delta.period[1:16,]

chip_order <- unique(data$chip)
data <- data %>%
  mutate(
    logP = -log10(FDR),
    # Define the significance groups
    group = case_when(
      FDR < 0.05 & estimate > 0 ~ "Sig Positive",
      FDR < 0.05 & estimate < 0 ~ "Sig Negative",
      TRUE ~ "Non-significant"
    ),
    # Error bar logic: positive values to the right (estimate+se), negative values to the left (estimate-se)
    error_limit = ifelse(estimate >= 0, estimate + se, estimate - se),
    # Keep the original order: reverse the factor levels so that the first entry appears at the top of the plot
    chip = factor(chip, levels = rev(chip_order))
  )

# 4. Define the plotting function
draw_regression_plot <- function(df_sub, title_name) {
  ggplot(df_sub, aes(x = estimate, y = chip, fill = group, alpha = logP)) +
    # Error bar layer: placed beneath the bars
    # xmin and xmax always span estimate and the computed error_limit
    geom_errorbarh(aes(xmin = estimate, xmax = error_limit), 
                   height = 0.2, color = "black", alpha = 1) +
    # Bar chart layer
    geom_col(color = "black", linewidth = 0.5) +
    # Colour mapping
    scale_fill_manual(values = c("Sig Positive" = "#E41A1C",   # red
                                 "Sig Negative" = "#377EB8",   # blue
                                 "Non-significant" = "grey80")) +
    # Alpha mapping (the smaller the P value, the more significant and the darker the colour)
    scale_alpha_continuous(range = c(0.7, 1), guide = "none") +
    # Zero reference line
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
    labs(title = "",
         x = "Estimate (Coefficient)",
         y = "",
         fill = "Significance") +
    theme_bw(base_size = 16) +
    theme(
      panel.grid.minor = element_blank(),
      axis.text = element_text(color = "black")
    )
}

# 5. Generate and display the plots
plot_d0_d7 <- draw_regression_plot(dplyr::filter(data, period == "d0_d7"), "d0_d7")
plot_d7_d15 <- draw_regression_plot(dplyr::filter(data, period == "d7_d15"), "d7_d15")

print(plot_d0_d7)
print(plot_d7_d15)

ggsave("outplots/fig3/sup3_d.d0-d7.clu1234.estimate.barplots.pdf",plot_d0_d7,width=6,height=6,dpi=300)
ggsave("outplots/fig3/sup3_d.d15-d7.clu1234.estimate.barplots.pdf",plot_d7_d15,width=6,height=6,dpi=300)
```





