
<!--
Reproduces main Figure 4: mean-signal line/region plots built from deepTools computeMatrix tables plus related annotation panels. Inputs (examples): output/computeMatrix/{H3K27me3,H3K27ac}/{0d,7d,15d}.matrix.tab. Outputs: outplots/fig4*.pdf.
--> 




fig4.b: mean signal of H3K27me3 and H3K27ac in clu567
```{r}
library(patchwork)
# clu567.mean.df is assumed to exist with the structure shown
# For illustration the data structure is kept as is; the actual run uses your own df directly

clu567.tab.peaks <- c(
    roi_info.split[roi_info.split$cluster == 5,"peak"],
    roi_info.split[roi_info.split$cluster == 6,"peak"],
    roi_info.split[roi_info.split$cluster == 7,"peak"]
)

l1 <- rbind(
    data.frame(signal = rowMeans(read.table("~/Project/chromHMM/DiffCluster/output/computeMatrix/H3K27me3/0d.matrix.tab",skip=3,header = F,sep = "\t")[,101:200]),category = "H3K27me3",time = "d0",peak = clu567.tab.peaks),
    data.frame(signal = rowMeans(read.table("~/Project/chromHMM/DiffCluster/output/computeMatrix/H3K27me3/7d.matrix.tab",skip=3,header = F,sep = "\t")[,101:200]),category = "H3K27me3",time = "d7",peak = clu567.tab.peaks),
    data.frame(signal = rowMeans(read.table("~/Project/chromHMM/DiffCluster/output/computeMatrix/H3K27me3/15d.matrix.tab",skip=3,header = F,sep = "\t")[,101:200]),category = "H3K27me3",time = "d15",peak = clu567.tab.peaks)
)

l2 <- rbind(
    data.frame(signal = rowMeans(read.table("~/Project/chromHMM/DiffCluster/output/computeMatrix/H3K27ac/0d.matrix.tab",skip=3,header = F,sep = "\t")[,101:200]),category = "H3K27ac",time = "d0",peak = clu567.tab.peaks),
    data.frame(signal = rowMeans(read.table("~/Project/chromHMM/DiffCluster/output/computeMatrix/H3K27ac/7d.matrix.tab",skip=3,header = F,sep = "\t")[,101:200]),category = "H3K27ac",time = "d7",peak = clu567.tab.peaks),
    data.frame(signal = rowMeans(read.table("~/Project/chromHMM/DiffCluster/output/computeMatrix/H3K27ac/15d.matrix.tab",skip=3,header = F,sep = "\t")[,101:200]),category = "H3K27ac",time = "d15",peak = clu567.tab.peaks)
)

clu567.mean.df <- rbind(l1,l2)
clu567.mean.df$time <- factor(clu567.mean.df$time,levels = c("d0","d7","d15"))


# 1. Split the data
df_me3 <- clu567.mean.df %>% dplyr::filter(category == "H3K27me3")
df_ac  <- clu567.mean.df %>% dplyr::filter(category == "H3K27ac")

# 2. Define the colour gradients
# H3K27me3: blue scale, d0 (dark) -> d7 (medium) -> d15 (light)
# The colours of the three time points are set manually here, or via scale_color_manual
# Dark blue, medium blue and light blue are used here
cols_me3 <- c("d0" = "#08306B", "d7" = "#4292C6", "d15" = "#C6DBEF") 

# H3K27ac: red scale, d0 (light) -> d7 (medium) -> d15 (dark)
# Light red, medium red and dark red are used here
cols_ac  <- c("d0" = "#FCAE91", "d7" = "#FB6A4A", "d15" = "#CB181D")

# 3. First plot: H3K27me3 (blue scale)
p1 <- ggplot(df_me3, aes(x = time, y = signal, fill = time, color = time)) +
stat_boxplot(geom = "errorbar",
               width = 0.2,
               size = 1.0,
               position = position_dodge(0.8)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1, alpha = 1) +
  scale_fill_manual(values = cols_me3) +
  scale_color_manual(values = c("black","black","black")) +
  labs(title = "", x = "", y = "log2(H3K27me3 CPM +1)") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", # show the legend at the bottom
        legend.direction = "horizontal",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text.x = element_text(size = 10),
        panel.grid.major.x = element_blank())

# 4. Second plot: H3K27ac (red scale)
p2 <- ggplot(df_ac, aes(x = time, y = signal, fill = time, color = time)) +
stat_boxplot(geom = "errorbar",
               width = 0.2,
               size = 1.0,
               position = position_dodge(0.8)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1, alpha = 1) +
  scale_fill_manual(values = cols_ac) +
  scale_color_manual(values = c("black","black","black")) +
  labs(title = "", x = "", y = "log2(H3K27ac CPM +1)") +
  theme_minimal(base_size = 16) +
  theme(legend.position = "bottom", # show the legend at the bottom
        legend.direction = "horizontal",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text.x = element_text(size = 10),
        panel.grid.major.x = element_blank())

# 5. Combine the plots vertically
# The / symbol denotes vertical stacking
# plot_layout heights can adjust the height ratio between the upper and lower panels
final_plot <- p1 + p2 + 
  patchwork::plot_layout(width = c(1, 1)) & # global theme adjustment (if needed)
  theme(plot.margin = margin(10, 10, 10, 10))

# Display the plot
print(final_plot)

# Save if needed
ggsave("outplots/fig4/fig4_b.clu567.signal.mean.pdf", final_plot, width = 5, height = 5,dpi = 300)


```


# ==============================================================
## fig4.c Sankey diagram showing the bivalent state of all cluster5/6/7 bins and their transitions
```{r}
input.df <- roi_list[roi_info.split[roi_info.split$cluster %in% c(5,6,7),"peak"]]
# Load required libraries
library(tidyverse)
library(ggalluvial) # for drawing the Sankey diagram

# -------------------------------------------------------

# 1. Define the state assignment function (vectorised for speed)
get_state <- function(ac, me3) {
  case_when(
    ac == 1 & me3 == 1 ~ "Bivalent",
    ac == 1 & me3 == 0 ~ "H3K27ac_only",
    ac == 0 & me3 == 1 ~ "H3K27me3_only",
    ac == 0 & me3 == 0 ~ "None",
    TRUE ~ NA_character_
  )
}

# 2. Process the data: extract the state transitions of all ROIs
print("Processing ROI states...")

all_bins_states <- map_dfr(1:length(input.df), function(i) {
  roi <- input.df[[i]]
  
  # Extract column 2 (H3K27ac) and column 3 (H3K27me3)
  # Time points 1, 2, 3
  t1_ac <- roi[, 2, 1]
  t1_me3 <- roi[, 3, 1]
  
  t2_ac <- roi[, 2, 2]
  t2_me3 <- roi[, 3, 2]
  
  t3_ac <- roi[, 2, 3]
  t3_me3 <- roi[, 3, 3]
  
  # Build the state table of this ROI
  data.frame(
    roi_id = names(input.df)[i],
    bin_id = 1:nrow(roi),
    d0 = get_state(t1_ac, t1_me3),
    d7 = get_state(t2_ac, t2_me3),
    d15 = get_state(t3_ac, t3_me3)
  )
})

# 3. Count the transition frequencies (for the Sankey diagram)
sankey_data <- all_bins_states %>%
  group_by(d0, d7, d15) %>%
  summarise(Freq = n(), .groups = 'drop')

# Inspect the first few rows of the summary
print(head(sankey_data))

# 4. Overall proportion of each state per time point (abundance level)
abundance_stats <- all_bins_states %>%
  pivot_longer(cols = starts_with("d"), names_to = "Period", values_to = "State") %>%
  group_by(Period, State) %>%
  summarise(Count = n(), .groups = 'drop') %>%
  group_by(Period) %>%
  mutate(Percentage = Count / sum(Count) * 100)

print(abundance_stats)

library(tidyverse)
library(ggalluvial)

# 1. Prepare the data and set the factor order strictly (Order)
# Order from top to bottom: H3K27ac_only > Bivalent > H3K27me3_only > None
# In ggplot the stacking order is normally determined by the factor levels
target_levels <- rev(c("None", "H3K27me3_only", "Bivalent", "H3K27ac_only"))

threshold <- sum(sankey_data$Freq) * 0.005 # filter out paths below 0.5%
sankey_data_final <- sankey_data %>%
  filter(Freq >= threshold)
sankey_data_final$d0 <- factor(sankey_data_final$d0, levels = target_levels)
sankey_data_final$d7 <- factor(sankey_data_final$d7, levels = target_levels)
sankey_data_final$d15 <- factor(sankey_data_final$d15, levels = target_levels)

# 2. Define the colour scheme precisely
my_colors <- c(
  "H3K27ac_only"  = "#E41A1C", # vivid red/orange, represents Active
  "Bivalent"      = "#E6AB02", # professional gold/dark yellow
  "H3K27me3_only" = "#377EB8", # steady blue, represents Repressed
  "None"          = "#7F7F7F"  # deepened grey to keep the flows legible
)

# 3. Plot
p <- ggplot(sankey_data_final,
       aes(y = Freq, axis1 = d0, axis2 = d7, axis3 = d15)) +
  # Draw the flow lines
  # A larger knot.pos makes the curves smoother and reduces clutter at crossings
  geom_alluvium(aes(fill = d0), width = 1/8, alpha = 0.45, knot.pos = 0.4) +
  # Draw the stratum bars
  geom_stratum(aes(fill = after_stat(stratum)), width = 1/6, color = "white", size = 0.3) +
  # Labels: bold white stands out better on a dark background
  geom_text(stat = "stratum", aes(label = after_stat(stratum)), 
            size = 3, fontface = "bold", color = "black") +
  # Apply the custom colours
  scale_fill_manual(values = my_colors) +
  # Set the X axis
  scale_x_discrete(limits = c("d0", "d7", "d15"), expand = c(.1, .1)) +
  # Publication-quality theme adjustments
  theme_minimal(base_size = 16) +
  theme(
    text = element_text(family = "sans"),
    axis.text.y = element_text(color = "black", size = 10),
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.x = element_text(size = 12, face = "bold", color = "black"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "#EEEEEE"),
    legend.position = "right",
    legend.title = element_text(face = "bold")
  ) +
  labs(
    y = "Number of Bins",
    fill = "Chromatin State",
    title = "Dynamics of H3K27ac & H3K27me3 Transitions"
  )

ggsave("outplots/fig4/fig4_c.Sankey_Final_Version.pdf", p, width = 10, height = 8)
```



# =================================================
## fig4_d Descriptive result showing that bivalent structure is more frequent in cluster5, 6 and 7 than in the background
```{r}
clu567.bed <- roi_info.split[roi_info.split$cluster %in% c(5,6,7),]
clu567.control <- rbind(read.table("output/clusters/controls/cluster5.bg.bed"),read.table("output/clusters/controls/cluster6.bg.bed"),read.table("output/clusters/controls/cluster7.bg.bed")) %>% unique()
clu567.control <- bt.sort(i=clu567.control,faidx="~/reference/hg19.chrom.sizes")

clu567.control$V2 <- round(clu567.control$V2/200)*200
clu567.control$V3 <- round(clu567.control$V3/200)*200
clu567.control$V4 <- paste0(clu567.control$V1,":",clu567.control$V2,":",clu567.control$V3)

clu567.control.counts <- c(0,0,0)
for (i in 1:nrow(clu567.control)){
    input <- clu567.control[i,]
    index <- BedInter(input[1,4],msc.all)
    d0 <- msc.all[index,]
    d7 <- adi_7d.all[index,]
    d15 <- adi_15d.all[index,]
    a1 <- ifelse(any(d0$H3K27me3 == 1 & d0$H3K27ac == 1),1,0)
    a2 <- ifelse(any(d7$H3K27me3 == 1 & d7$H3K27ac == 1),1,0)
    a3 <- ifelse(any(d15$H3K27me3 == 1 & d15$H3K27ac == 1),1,0)
    clu567.control.counts <- clu567.control.counts + c(a1,a2,a3)
    #print(i)
}


clu567.control.biv.d0 <- bt.intersect(a=clu567.control,b=msc.all[msc.all$H3K27ac == 1 & msc.all$H3K27me3 == 1,],wa=T) %>% unique()
clu567.control.biv.d7 <- bt.intersect(a=clu567.control,b=adi_7d.all[adi_7d.all$H3K27ac == 1 & adi_7d.all$H3K27me3 == 1,],wa=T) %>% unique()
clu567.control.biv.d15 <- bt.intersect(a=clu567.control,b=adi_15d.all[adi_15d.all$H3K27ac == 1 & adi_15d.all$H3K27me3 == 1,],wa=T) %>% unique()

clu567.bed.biv.d0 <- bt.intersect(a=clu567.bed,b=msc.all[msc.all$H3K27ac == 1 & msc.all$H3K27me3 == 1,],wa=T) %>% unique()
clu567.bed.biv.d7 <- bt.intersect(a=clu567.bed,b=adi_7d.all[adi_7d.all$H3K27ac == 1 & adi_7d.all$H3K27me3 == 1,],wa=T) %>% unique()
clu567.bed.biv.d15 <- bt.intersect(a=clu567.bed,b=adi_15d.all[adi_15d.all$H3K27ac == 1 & adi_15d.all$H3K27me3 == 1,],wa=T) %>% unique()


df <- data.frame(
    category=c("control","clu567","control","clu567","control","clu567"),
    period=c("d0","d0","d7","d7","d14","d14"),
    ratio=c(nrow(clu567.control.biv.d0)/nrow(clu567.control),nrow(clu567.bed.biv.d0)/nrow(clu567.bed),nrow(clu567.control.biv.d7)/nrow(clu567.control),nrow(clu567.bed.biv.d7)/nrow(clu567.bed),nrow(clu567.control.biv.d15)/nrow(clu567.control),nrow(clu567.bed.biv.d15)/nrow(clu567.bed))
)
df$category <- factor(df$category,c("control","clu567"))
df$period <- factor(df$period,c("d0","d7","d14"))

p <- ggplot(df, aes(x = period, y = ratio, fill = category)) +
  # position_dodge places the bars side by side
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  # Set the colours: grey for Control, vivid orange for the ROI (clu567, representing Bivalent)
  scale_fill_manual(values = c("control" = "#BDC3C7", "clu567" = "#E67E22")) +
  # Display as percentages
  scale_y_continuous(labels = scales::percent_format(), expand = expansion(mult = c(0, 0.1))) +
  labs(
    y = "Proportion of H3K27me3-ac Bivalent State",
    fill = "Region Type"
  ) +
  theme_classic(base_size = 16) +
  theme(
    legend.position = "right",
    axis.text = element_text(color = "black"),
    axis.title = element_text(face = "bold")
  )
ggsave("outplots/fig4/fig4_d.control.vs.clu567.bivalent.ratio.pdf",p,width=8,height=5,dpi=300)


## Boxplot of the proportion of bivalent bins among all bins

clu567.d0.bivalent <- data.frame(matrix(nrow=nrow(clu567.bed),ncol=3))
colnames(clu567.d0.bivalent) <- c("peak","cluster","biv")
clu567.d7.bivalent <- data.frame(matrix(nrow=nrow(clu567.bed),ncol=3))
colnames(clu567.d7.bivalent) <- c("peak","cluster","biv")
clu567.d15.bivalent <- data.frame(matrix(nrow=nrow(clu567.bed),ncol=3))
colnames(clu567.d15.bivalent) <- c("peak","cluster","biv")


for(i in 1:nrow(clu567.bed)){
    input <- roi_list[clu567.bed$peak[i]][[1]]
    out1 <- length(which(input[,3,1] == 1 & input[,2,1] == 1))/nrow(input)
    clu567.d0.bivalent[i,] <- c(clu567.bed$peak[i],clu567.bed$cluster[i],out1)
    out2 <- length(which(input[,3,2] == 1 & input[,2,2] == 1))/nrow(input)
    clu567.d7.bivalent[i,] <- c(clu567.bed$peak[i],clu567.bed$cluster[i],out2)
    out3 <- length(which(input[,3,3] == 1 & input[,2,3] == 1))/nrow(input)
    clu567.d15.bivalent[i,] <- c(clu567.bed$peak[i],clu567.bed$cluster[i],out3)
}
clu567.d0.bivalent$biv <- as.numeric(clu567.d0.bivalent$biv)
clu567.d7.bivalent$biv <- as.numeric(clu567.d7.bivalent$biv)
clu567.d15.bivalent$biv <- as.numeric(clu567.d15.bivalent$biv)

boxplot(clu567.d0.bivalent$biv[clu567.d0.bivalent$biv!=0],clu567.d7.bivalent$biv[clu567.d7.bivalent$biv!=0],clu567.d15.bivalent$biv[clu567.d15.bivalent$biv!=0])


clu567.coltrol.bivalent <- data.frame(matrix(nrow=nrow(clu567.control),ncol=3))
l1 <- bt.intersect(a=msc.all[msc.all$H3K27ac == 1 & msc.all$H3K27me3 == 1,],b=clu567.control,wo=T)
l2 <- bt.intersect(a=adi_7d.all[adi_7d.all$H3K27ac == 1 & adi_7d.all$H3K27me3 == 1,],b=clu567.control,wo=T)
l3 <- bt.intersect(a=adi_15d.all[adi_15d.all$H3K27ac == 1 & adi_15d.all$H3K27me3 == 1,],b=clu567.control,wo=T)

for (i in 1:nrow(clu567.control)){
    if(any(clu567.control[i,4] %in% l1$V15,clu567.control[i,4] %in% l2$V15,clu567.control[i,4] %in% l3$V15) == F){
        clu567.coltrol.bivalent[i,] <- c(0,0,0)
    }
    else{
        out1 <- l1[l1$V15 == clu567.control[i,4],]
        out2 <- l2[l2$V15 == clu567.control[i,4],]
        out3 <- l3[l3$V15 == clu567.control[i,4],]
        clu567.coltrol.bivalent[i,] <- c(nrow(out1),nrow(out2),nrow(out3))/((clu567.control[i,3]-clu567.control[i,2])/200)
    }
}
colnames(clu567.coltrol.bivalent) <- c("d0","d7","d15")

boxplot(clu567.coltrol.bivalent$d0[clu567.coltrol.bivalent$d0 !=0],clu567.d0.bivalent$biv[clu567.d0.bivalent$biv!=0],clu567.coltrol.bivalent$d7[clu567.coltrol.bivalent$d7 !=0],clu567.d7.bivalent$biv[clu567.d7.bivalent$biv!=0],clu567.coltrol.bivalent$d15[clu567.coltrol.bivalent$d15 !=0],clu567.d15.bivalent$biv[clu567.d15.bivalent$biv!=0])

library(ggplot2)

# 1. Extract the data and build a long-format data frame
# Note: the code writes d15 while the description says d14; d15 is used here consistently with the variable names in the code, and the labels can be changed as needed
df_list <- list(
  data.frame(Value = clu567.coltrol.bivalent$d0[clu567.coltrol.bivalent$d0 != 0], Period = "d0", Category = "control"),
  data.frame(Value = clu567.d0.bivalent$biv[clu567.d0.bivalent$biv != 0],          Period = "d0", Category = "clu567"),
  
  data.frame(Value = clu567.coltrol.bivalent$d7[clu567.coltrol.bivalent$d7 != 0], Period = "d7", Category = "control"),
  data.frame(Value = clu567.d7.bivalent$biv[clu567.d7.bivalent$biv != 0],          Period = "d7", Category = "clu567"),
  
  data.frame(Value = clu567.coltrol.bivalent$d15[clu567.coltrol.bivalent$d15 != 0], Period = "d15", Category = "control"),
  data.frame(Value = clu567.d15.bivalent$biv[clu567.d15.bivalent$biv != 0],          Period = "d15", Category = "clu567")
)

# Merge all data
plot_data <- do.call(rbind, df_list)

# 2. Set the factor order (otherwise the x axis is sorted alphabetically as d0, d15, d7)
plot_data$Period <- factor(plot_data$Period, levels = c("d0", "d7", "d15"))
plot_data$Category <- factor(plot_data$Category, levels = c("control", "clu567"))

p <- ggplot(plot_data, aes(x = Period, y = Value, fill = Category)) +
  geom_boxplot(outlier.size = 0.8, alpha = 0.8) + # draw the boxplot
  scale_fill_manual(values = c("control" = "#BDC3C7", "clu567" = "#E67E22")) + # set the colours
  geom_boxplot(outlier.shape = NA) + # hide the outliers to avoid plotting them twice
  labs(
    x = "Time Period", 
    y = "Bivalent Level"
  ) +
  theme_classic(base_size = 16) + # use the clean classic theme
  theme(
    axis.text = element_text(color = "black"),
    legend.position = "right"
  )
ggsave("outplots/fig4/fig4_d.control.vs.clu567.bivalent.binratio.pdf",p,width=8,height=5,dpi=300)
```




# =================================================
## fig4_e Quadrant plot comparing the ac signal (d0-d7) with the me3 signal (d7-d15) in ROIs that contain bivalent structure
```{r}
clu567.biv.d7.ac <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig4.bivalent/clu567.d7.biv.bin.H3K27ac.CPM.tab")
clu567.biv.d7.me3 <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig4.bivalent/clu567.d7.biv.bin.H3K27me3.CPM.tab")


clu567.biv.d7.delta <- data.frame(
    ac_d0d7 = log2(clu567.biv.d7.ac[,5]/clu567.biv.d7.ac[,4]),
    ac_d7d15 = log2(clu567.biv.d7.ac[,6]/clu567.biv.d7.ac[,5]),
    me3_d0d7 = log2(clu567.biv.d7.me3[,5]/clu567.biv.d7.me3[,4]),
    me3_d7d15 = log2(clu567.biv.d7.me3[,6]/clu567.biv.d7.me3[,5])
)

# clu567.biv.d7.delta <- data.frame(
#     ac_d0d7 = clu567.biv.d7.ac[,5]-clu567.biv.d7.ac[,4],
#     ac_d7d15 = clu567.biv.d7.ac[,6]-clu567.biv.d7.ac[,5],
#     me3_d0d7 = clu567.biv.d7.me3[,5]-clu567.biv.d7.me3[,4],
#     me3_d7d15 = clu567.biv.d7.me3[,6]-clu567.biv.d7.me3[,5]
# )

quadrant_data <- clu567.biv.d7.delta %>%
  mutate(quadrant = case_when(
    ac_d0d7 > 0 & me3_d7d15 < 0 ~ "ac Up / me3 Down",
    ac_d0d7 > 0 & me3_d7d15 >= 0 ~ "ac Up / me3 Up",
    ac_d0d7 <= 0 & me3_d7d15 < 0 ~ "ac Down / me3 Down",
    ac_d0d7 <= 0 & me3_d7d15 >= 0 ~ "ac Down / me3 Up"
  )) %>%
  group_by(quadrant) %>%
  summarise(count = n()) %>%
  mutate(perc = paste0(round(count/sum(count)*100, 1), "%"))

# Plot: show the distribution density and the quadrant proportions
p <- ggplot(clu567.biv.d7.delta, aes(x=ac_d0d7, y=me3_d7d15)) +
  geom_bin2d(bins = 50) + # show the point density in heatmap form
  scale_fill_gradient(low = "lightblue", high = "red") +
  geom_vline(xintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0, linetype="dashed") +
  # Annotate the percentages in the four corners
  annotate("text", x = 4, y = -2.5, label = paste0("ac Up / me3 Down  ",quadrant_data[3,3]), size=6, color="darkred", fontface="bold") +
  labs(title="",
       x="H3K27ac (d0-d7)", y="H3K27me3 (d7-d15)") +
  theme_bw(base_size = 16)+
  theme(
    axis.text = element_text(color = "black")
  )
ggsave("outplots/fig4/fig4_d.Quadrant.Proportions.pdf",p,width=5.5,height=5,dpi=300)




```


# =================================================
## fig4_e Cumulative distribution function (CDF) curves to test whether the d15-d7 signal differs significantly between bivalent ROIs and me3 ROIs
```{r}

clu567.biv.d7.bed <- bt.intersect(a=clu567.bed,b=adi_7d.all[adi_7d.all$H3K27ac == 1 & adi_7d.all$H3K27me3 == 1,],wa=T) %>% unique()
colnames(clu567.biv.d7.bed) <- colnames(clu567.bed)
clu567.me3.d7.bed <- clu567.bed[!clu567.bed$peak %in% clu567.biv.d7.bed$peak,]

write.table(clu567.biv.d7.bed,file = "output/fig4.bivalent/clu567.biv.bin.bed",row.names = F,col.names = F,sep = "\t",quote = F)
write.table(clu567.me3.d7.bed,file = "output/fig4.bivalent/clu567.me3.bin.bed",row.names = F,col.names = F,sep = "\t",quote = F)

clu567.biv.d7.me3 <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig4.bivalent/clu567.biv.bin.H3K27me3.CPM.tab")

clu567.me3.d7.me3 <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig4.bivalent/clu567.me3.bin.H3K27me3.CPM.tab")


l1 <- clu567.biv.d7.me3[,6]-clu567.biv.d7.me3[,5]
l2 <- clu567.me3.d7.me3[,6]-clu567.me3.d7.me3[,5]



# 1. Build the plotting data frame
df_plot <- data.frame(
  delta = c(l1, l2),
  group = c(rep("Bivalent", length(l1)), 
            rep("H3K27me3 only", length(l2)))
)

# 2. Statistical test (K-S test)
# This test is designed to detect a significant shift between two distributions
ks_result <- ks.test(l1, l2)
p_value <- ks_result$p.value

# Print the P value
cat("K-S test P-value:", p_value, "\n")

library(ggplot2)

# Plot
p <- ggplot(df_plot, aes(x = delta, color = group)) +
  stat_ecdf(geom = "step", size = 1.5) + 
  # Set the colours: blue for the control group (l2), orange/red for the test group (l1)
  scale_color_manual(values = c("Bivalent" = "#E64B35FF", 
                                "H3K27me3 only" = "#4DBBD5FF")) +
  theme_bw(base_size = 16) +
  # Guide line: x = 0 means no change in signal
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
  # Guide line: y = 0.5 makes the median shift easy to read
  geom_hline(yintercept = 0.5, linetype = "dotted", color = "grey70") +
  labs(
    x = "Signal Change (d15 - d7 log2CPM)",
    y = "Cumulative Proportion",
    color = "Group at d7"
  ) +
  # Fine-tune the axis range according to the summary data
  coord_cartesian(xlim = c(-0.5, 0.3)) +
  theme(
    legend.position = "top",
    panel.grid.minor = element_blank(),
    plot.title = element_text(hjust = 0.5, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5)
  )

ggsave("outplots/fig4/fig4_e.biv.vs.me3.KS.test.CDF.curve.pdf",p,width=5,height=5,dpi=300)


```


# =====================================================
## fig4_g Spatial asymmetry within spatiotemporal asymmetry, i.e. the relationship between H3K27ac loss and H3K27me3 gain
## Expected to be approached from two angles: 1. co-enriched CRRs do indeed appear


```{r}

clu567.biv.d7.delta.me3 <- read.table(file = "~/Project/chromHMM/bamsubtract/d7.biv.H3K27me3.tab",skip = 3)

clu567.biv.d7.delta.ac <- read.table(file = "~/Project/chromHMM/bamsubtract/d7.biv.H3K27ac.tab",skip = 3)

library(tidyverse)

get_profile_df <- function(mat, mark_name) {
  tibble(
    bin = 1:ncol(mat),
    mean = colMeans(mat, na.rm = TRUE),
    sd = apply(mat, 2, sd, na.rm = TRUE),
    n = nrow(mat),
    sem = sd / sqrt(n),
    mark = mark_name
  )
}

ac_profile <- get_profile_df(clu567.biv.d7.delta.ac, "H3K27ac")
me3_profile <- get_profile_df(clu567.biv.d7.delta.me3, "H3K27me3")

df <- bind_rows(me3_profile, ac_profile)

start_bin <- 51
end_bin   <- 150
n_bin     <- 200

line_cols <- c(
  "H3K27ac"  = "#D55E00",
  "H3K27me3" = "#2C5AA0"
)

fill_cols <- c(
  "H3K27ac"  = "#F6C29A",
  "H3K27me3" = "#AFC6E9"
)

p <- ggplot(df, aes(x = bin, y = mean, color = mark, fill = mark)) +
  annotate("rect",
           xmin = start_bin, xmax = end_bin,
           ymin = -Inf, ymax = Inf,
           fill = "grey92", alpha = 0.8) +
  geom_hline(yintercept = 0, color = "red3", linewidth = 0.8) +
  geom_vline(xintercept = c(start_bin, end_bin),
             linetype = "dashed", color = "grey40", linewidth = 0.5) +
  geom_ribbon(aes(ymin = mean - sem, ymax = mean + sem),
              alpha = 0.35, color = NA) +
  geom_line(linewidth = 1.1) +
  facet_wrap(~mark, nrow = 1, scales = "fixed") +
  scale_color_manual(values = line_cols) +
  scale_fill_manual(values = fill_cols) +
  scale_x_continuous(
    breaks = c(1, start_bin, end_bin, n_bin),
    labels = c("-1.0 kb", "Start", "End", "1.0 kb"),
    expand = c(0.01, 0.01)
  ) +
  coord_cartesian(ylim = c(-0.3, 0.6)) +
  labs(x = NULL, y = NULL) +
  theme_classic(base_size = 14) +
  theme(
    strip.background = element_blank(),
    strip.text = element_text(face = "bold", size = 13),
    axis.text = element_text(color = "black", size = 11),
    axis.title = element_text(color = "black"),
    legend.position = "none",
    panel.spacing = unit(1.2, "lines")
  )

print(p)

ggsave("outplots/fig4/fig4_g.biv.d7.delta.profile.pdf", p, width = 8, height = 4, dpi = 300)

```





# =====================================================
## fig4_h TF enrichment with HOMER
```{r}
library(patchwork)

isbiv.known.motif <- fread(file = "~/Project/chromHMM/DiffCluster/output/homer/isbiv.vs.nobiv/knownResults.txt",header = T) %>% as.data.frame()
isbiv.known.motif$`P-value` <- exp(1)^isbiv.known.motif$`Log P-value`
isbiv.known.motif$FDR <- p.adjust(isbiv.known.motif$`P-value`)

isbiv.known.motif.list <- isbiv.known.motif[c(1,3,5,6,7),"Motif Name"]

nobiv.known.motif <- fread(file = "output/homer/nobiv.vs.isbiv/knownResults.txt",header = T) %>% as.data.frame()
nobiv.known.motif$`P-value` <- exp(1)^nobiv.known.motif$`Log P-value`
nobiv.known.motif$FDR <- p.adjust(nobiv.known.motif$`P-value`)
nobiv.known.motif.list <- nobiv.known.motif[c(1,2,4),"Motif Name"]



isbiv.homer.motif <- read.table(file = "output/homer/isbiv.vs.nobiv/homerResults.txt",header = T,sep = "\t") %>% as.data.frame()
isbiv.homer.motif$pvalue <- exp(1)^isbiv.homer.motif$log.P.pvalue
isbiv.homer.motif$FDR <- p.adjust(isbiv.homer.motif$pvalue)

isbiv.homer.motif.list <- isbiv.homer.motif[c(1,2,3,4,5,9,12,13,22,29),1]


nobiv.homer.motif <- read.table(file = "output/homer/nobiv.vs.isbiv/homerResults.txt",header = T,sep = "\t") %>% as.data.frame()
nobiv.homer.motif$pvalue <- exp(1)^nobiv.homer.motif$log.P.pvalue
nobiv.homer.motif$FDR <- p.adjust(nobiv.homer.motif$pvalue)

nobiv.homer.motif[c(2,4,7,9,10,15),]


df <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/homer/homer.selected.results.txt")
colnames(df) <- c("TF","category","condition","pvalue")

df <- df %>%
  mutate(
    negLogP = -log10(pvalue),
    # Unify the source spelling and convert it to a factor to control the display order (known first)
    source = factor(ifelse(category %in% c("know", "known"), "Known Motifs", "De Novo (Homer)"), 
                    levels = c("Known Motifs", "De Novo (Homer)")),
    # Set the condition order (isbiv first)
    condition = factor(condition, levels = rev(c("isbiv", "nobiv")))
  )


df1 <- df[df$source == "Known Motifs",]
df1$TF <- factor(df1$TF,levels=unique(df1$TF))

# Plot 1: Known Motifs (red scale)
p1 <- ggplot(df1, aes(y = condition, x = TF)) +
  geom_tile(aes(fill = negLogP), color = "white") +
  scale_fill_gradient2(
    low = "white",          # colour corresponding to 0
    mid = "#fff0e8",        # light red corresponding to 1.301 (can be replaced with any light red of your choice)
    high = "#de2d26",       # dark red corresponding to the maximum
    midpoint = 1.301,       # set the value of the midpoint
    limits = c(0, NA),      # make sure the scale starts at 0
    name = "Known\n-log10P"
  ) +
  theme_bw(base_size = 16) + labs(x = NULL, y = NULL, title = "Known Motifs") 

df2 <- df[df$source == "De Novo (Homer)",]
df2$TF <- factor(df2$TF,levels=unique(df2$TF))

# Plot 2: Homer Motifs (yellow/orange scale)
p2 <- ggplot(df2, aes(y = condition, x = TF)) +
  geom_tile(aes(fill = negLogP), color = "white") +
  scale_fill_gradient(low = "#fff7bc", high = "#d95f0e", name = "Homer\n-log10P") +
  theme_bw(base_size = 16) + labs(x = NULL, y = NULL, title = "De Novo Motifs")

# Combine
p3 <- p1 | p2 
ggsave("outplots/fig4/fig4_h.homer.heatmap.pdf", p3, width = 16, height = 5, dpi = 300)

```


## ==========================================================================
## supplementary fig 4 a,b,c,d
## a. Number of co-enriched CRRs at the three time points
## b. Flow and inheritance relationships of the three-time-point co-enriched CRRs
## c. H3K27me3 and H3K27ac trends of the three-time-point co-enriched CRRs (ideally independent ones)
## === The above is used to establish the central intermediate role of the d7 co-enriched CRRs: numerous, highly dynamic, in the middle of differentiation
## d. Quadrant plots of the d7 co-enriched CRRs for each combination of delta data, selecting the appropriate one, i.e. fig4_e
## ==========================================================================

```{r}
# sup4_a Number of co-enriched CRRs at the three time points

clu567.biv.crrs.counts <- data.frame(
    period = factor(c("d0","d7","d14"),levels = c("d0","d7","d14")),
    counts = c(
        length(which(clu567.d0.bivalent$biv !=0)),
        length(which(clu567.d7.bivalent$biv !=0)),
        length(which(clu567.d15.bivalent$biv !=0)))
)

my_colors <- c("#2b5c8f","#feb24c","#31a354")

# 3. Draw the bar plot
p <- ggplot(clu567.biv.crrs.counts, aes(x = period, y = counts, fill = period)) +
  geom_col(width = 0.5) + # geom_col draws bars of the specified height, width controls the bar width
  scale_fill_manual(values = my_colors) + # customise the colours of the three bars
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) + # set the bottom to 0 (flush with the x axis) and leave 10% headroom at the top
  labs(x = "", y = "Co-enriched CRRs counts") +
  theme_classic(base_size = 16) + # classic theme (includes x and y axis lines)
  theme(legend.position = "none") # hide the legend (the x axis is already labelled)

# Display the plot
print(p)
ggsave("outplots/fig4/sup4_a.clu567.period.crrs.counts.barplot.pdf", p, width = 5, height = 5, dpi = 300)

# sup4_b Inheritance relationships of the three-time-point co-enriched CRRs

l1 <- clu567.d0.bivalent[clu567.d0.bivalent$biv !=0,"peak"]
l2 <- clu567.d7.bivalent[clu567.d7.bivalent$biv !=0,"peak"]
l3 <- clu567.d15.bivalent[clu567.d15.bivalent$biv !=0,"peak"]

library(ggplot2)
library(dplyr)


# ==========================================
# 1. Set operations (count inherited, lost and newly emerged)
# ==========================================

# --- Stage 1 (l1) ---
s1_to_s2  <- intersect(l1, l2)        # retained from l1 into l2
s1_lost   <- setdiff(l1, l2)          # lost from l1

# --- Stage 2 (l2) ---
s2_from_s1     <- intersect(l2, l1)    # derived from l1
s2_from_s1_s3  <- intersect(s2_from_s1, l3) # derived from l1 and carried on into l3
s2_from_s1_lost<- setdiff(s2_from_s1, l3)   # derived from l1 but lost at l2

s2_new         <- setdiff(l2, l1)     # newly emerged at l2
s2_new_s3      <- intersect(s2_new, l3)     # newly emerged at l2 and carried into l3
s2_new_lost    <- setdiff(s2_new, l3)       # newly emerged at l2 and lost at l2

# --- Stage 3 (l3) ---
s3_from_s2_old <- intersect(l3, s2_from_s1) # derived from l1->l2
s3_from_s2_new <- intersect(l3, s2_new)     # derived from l2 newly emerged
s3_new         <- setdiff(l3, l2)           # newly emerged at l3

# ==========================================
# 2. Coordinate and layout calculation (y = 0 is the top alignment line)
# ==========================================
max_count <- max(length(l1), length(l2), length(l3))
GAP <- max_count * 0.08   # gap between the newly emerged block and the inherited block
SUBGAP <- max_count * 0.02 # small gap within the same block (retained vs lost)
bw <- 0.2                  # half-width of the bars

# --- L1 coordinates ---
# Inherited block (Top = 0)
y_l1_inh_top <- 0
y_l1_inh_bot <- -length(s1_to_s2)
# Lost block
y_l1_lost_top <- y_l1_inh_bot - SUBGAP
y_l1_lost_bot <- y_l1_lost_top - length(s1_lost)

# --- L2 coordinates ---
# Block inherited from L1 (aligned at Top = 0)
y_l2_inh_s3_top <- 0
y_l2_inh_s3_bot <- -length(s2_from_s1_s3)
y_l2_inh_lost_top <- y_l2_inh_s3_bot - SUBGAP
y_l2_inh_lost_bot <- y_l2_inh_lost_top - length(s2_from_s1_lost)

# Newly emerged block (with a GAP in between)
y_l2_new_s3_top <- y_l2_inh_lost_bot - GAP
y_l2_new_s3_bot <- y_l2_new_s3_top - length(s2_new_s3)
y_l2_new_lost_top <- y_l2_new_s3_bot - SUBGAP
y_l2_new_lost_bot <- y_l2_new_lost_top - length(s2_new_lost)

# --- L3 coordinates ---
# Inherited from L1->L2 (aligned at Top = 0)
y_l3_inh1_top <- 0
y_l3_inh1_bot <- -length(s3_from_s2_old)

# Block inherited from the L2 newly emerged pool (with a GAP)
y_l3_inh2_top <- y_l3_inh1_bot - GAP
y_l3_inh2_bot <- y_l3_inh2_top - length(s3_from_s2_new)

# L3 newly emerged block (with another GAP)
y_l3_new_top <- y_l3_inh2_bot - GAP
y_l3_new_bot <- y_l3_new_top - length(s3_new)


# ==========================================
# 3. Build the rectangle data set
# ==========================================
rects <- data.frame(
  x1 = c(
    # Stage 1
    1-bw, 1-bw,
    # Stage 2
    2-bw, 2-bw, 2-bw, 2-bw,
    # Stage 3
    3-bw, 3-bw, 3-bw
  ),
  x2 = c(
    1+bw, 1+bw,
    2+bw, 2+bw, 2+bw, 2+bw,
    3+bw, 3+bw, 3+bw
  ),
  ymin = c(
    y_l1_inh_bot, y_l1_lost_bot,
    y_l2_inh_s3_bot, y_l2_inh_lost_bot, y_l2_new_s3_bot, y_l2_new_lost_bot,
    y_l3_inh1_bot, y_l3_inh2_bot, y_l3_new_bot
  ),
  ymax = c(
    y_l1_inh_top, y_l1_lost_top,
    y_l2_inh_s3_top, y_l2_inh_lost_top, y_l2_new_s3_top, y_l2_new_lost_top,
    y_l3_inh1_top, y_l3_inh2_top, y_l3_new_top
  ),
  type = c(
    "Inherited Flow 1", "Lost/Terminated",
    "Inherited Flow 1", "Lost/Terminated", "Inherited Flow 2", "Lost/Terminated",
    "Inherited Flow 1", "Inherited Flow 2", "Newly Emerged"
  ),
  count = c(
    length(s1_to_s2), length(s1_lost),
    length(s2_from_s1_s3), length(s2_from_s1_lost), length(s2_new_s3), length(s2_new_lost),
    length(s3_from_s2_old), length(s3_from_s2_new), length(s3_new)
  )
)

# ==========================================
# 4. Helper function for smooth Sankey ribbons
# ==========================================
make_ribbon <- function(x1, x2, y1_top, y1_bot, y2_top, y2_bot, group_id, type) {
  t <- seq(0, 1, length.out = 100)
  s <- 3 * t^2 - 2 * t^3 # smooth S-shaped curve
  x <- x1 + (x2 - x1) * t
  ytop <- y1_top + (y2_top - y1_top) * s
  ybot <- y1_bot + (y2_bot - y1_bot) * s
  
  data.frame(
    x = c(x, rev(x)),
    y = c(ytop, rev(ybot)),
    group = group_id,
    type = type
  )
}

# Generate the 3 main inheritance ribbons
ribbon1 <- make_ribbon(1+bw, 2-bw, y_l1_inh_top, y_l1_inh_bot, 0, y_l2_inh_lost_bot, 1, "Flow: L1 -> L2")
ribbon2 <- make_ribbon(2+bw, 3-bw, y_l2_inh_s3_top, y_l2_inh_s3_bot, y_l3_inh1_top, y_l3_inh1_bot, 2, "Flow: L1 -> L2 -> L3")
ribbon3 <- make_ribbon(2+bw, 3-bw, y_l2_new_s3_top, y_l2_new_s3_bot, y_l3_inh2_top, y_l3_inh2_bot, 3, "Flow: L2 -> L3")

ribbons <- rbind(ribbon1, ribbon2, ribbon3)

# ==========================================
# 5. Plot with ggplot2
# ==========================================

# Colour scheme
colors <- c(
  "Inherited Flow 1" = "#2b5c8f", # colour of the lineage inherited all the way from L1 (dark blue)
  "Inherited Flow 2" = "#feb24c", # colour of the L2 newly emerged states inherited into L3 (yellow/orange)
  "Newly Emerged"    = "#31a354", # colour of the newest states (green)
  "Lost/Terminated"  = "#d9d9d9"  # colour of the states that were not inherited (light grey)
)

p2 <- ggplot() +
  # 1. Draw the connecting ribbons (semi-transparent flows)
  geom_polygon(data = ribbons, aes(x = x, y = y, group = group, fill = type), alpha = 0.35) +
  # 2. Draw the stratum bars
  geom_rect(data = rects, aes(xmin = x1, xmax = x2, ymin = ymin, ymax = ymax, fill = type), color = "white", size = 0.3) +
  # 3. Label the sample counts on the bars (optional)
  geom_text(data = dplyr::filter(rects, count > 0), 
            aes(x = (x1 + x2)/2, y = (ymin + ymax)/2, label = count), 
            color = "black", size = 6, fontface = "bold") +
  # 4. Adjust the Y axis display (the alignment uses negative values, so absolute values are shown)
  scale_y_continuous(labels = abs, name = "Sample Counts") +
  scale_x_continuous(breaks = c(1, 2, 3), labels = c("L1", "L2", "L3"), limits = c(0.5, 3.5)) +
  scale_fill_manual(values = c(
    colors,
    "Flow: L1 -> L2"      = "#2b5c8f",
    "Flow: L1 -> L2 -> L3"= "#2b5c8f",
    "Flow: L2 -> L3"      = "#feb24c"
  )) +
  theme_classic(base_size = 16) +
  theme(
    axis.line.y = element_line(color = "black"),
    axis.line.x = element_blank(),
    axis.ticks.x = element_blank(),
    legend.title = element_blank(),
    legend.position = "none"
  ) +
  labs(x = "Time Stages", title = "")

ggsave("outplots/fig4/sup4_b.clu567.period.inherit.barplot.pdf", p2, width = 5, height = 5, dpi = 300)


## sup4_c H3K27ac and H3K27me3 signals actually recorded at the three time points
df_me3 <- clu567.mean.df %>% dplyr::filter(category == "H3K27me3")

df_ac  <- clu567.mean.df %>% dplyr::filter(category == "H3K27ac")


## Then only the independent co-enriched CRRs are counted
# --- Stage 1 (l1) ---
s1_lost

# --- Stage 2 (l2) ---
s2_new_lost

# --- Stage 3 (l3) ---
s3_new

df1 <- rbind(df_me3[df_me3$peak %in% s1_lost,],df_ac[df_ac$peak %in% s1_lost,])
df2 <- rbind(df_me3[df_me3$peak %in% s2_new_lost,],df_ac[df_ac$peak %in% s2_new_lost,])
df3 <- rbind(df_me3[df_me3$peak %in% s3_new,],df_ac[df_ac$peak %in% s3_new,])

library(ggplot2)
library(dplyr)
library(patchwork) # for figure assembly; run install.packages("patchwork") first if it is missing

library(ggplot2)
library(dplyr)
library(patchwork)

plot_epigenomic_signals <- function(df, 
                                    error_type = "se", 
                                    y_min_ac = 0.01, 
                                    y_min_me3 = 0.2) {
  
  # 1. Make sure time is ordered (d0 -> d7 -> d15)
  df$time <- factor(df$time, levels = c("d0", "d7", "d15"))
  
  # 2. Compute the mean, SD and SE for each combination of time and category
  df_summary <- df %>%
    group_by(category, time) %>%
    summarise(
      mean_signal = mean(signal, na.rm = TRUE),
      sd_signal   = sd(signal, na.rm = TRUE),
      n           = n(),
      se_signal   = sd_signal / sqrt(n),
      .groups     = "drop"
    )
  
  # Choose whether to use SE or SD
  if (tolower(error_type) == "sd") {
    df_summary$err <- df_summary$sd_signal
  } else {
    df_summary$err <- df_summary$se_signal
  }
  
  # 3. Define the colour palettes
  colors_ac  <- c("#fc9272", "#ef3b2c", "#a50f15") # H3K27ac red gradient
  colors_me3 <- c("#9ecae1", "#4292c6", "#08519c") # H3K27me3 blue gradient

  # ----------------------------------------------------
  # 4. Draw the H3K27ac panel (Y axis starts at y_min_ac, default 0.2)
  # ----------------------------------------------------
  df_ac <- df_summary %>% dplyr::filter(category == "H3K27ac")
  # Compute the y axis upper limit automatically so that the bar tops and error bars stay inside the plot
  max_ac <- max(df_ac$mean_signal + df_ac$err, na.rm = TRUE)
  y_max_ac <- max_ac + (max_ac - y_min_ac) * 0.15 

  p_ac <- ggplot(df_ac, aes(x = time, y = mean_signal, fill = time)) +
    geom_col(width = 0.55, color = "black", linewidth = 0.3) +
    geom_errorbar(
      aes(ymin = mean_signal - err, ymax = mean_signal + err),
      width = 0.15, linewidth = 0.6
    ) +
    scale_fill_manual(values = colors_ac) +
    scale_y_continuous(expand = c(0, 0)) + # remove the default gap
    coord_cartesian(ylim = c(y_min_ac, y_max_ac)) + # set the absolute Y axis range (starting at y_min_ac)
    labs(title = "H3K27ac", x = "Time", y = "Signal") +
    theme_classic(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      legend.position = "none"
    )

  # ----------------------------------------------------
  # 5. Draw the H3K27me3 panel (Y axis starts at y_min_me3, default 0.01)
  # ----------------------------------------------------
  df_me3 <- df_summary %>% dplyr::filter(category == "H3K27me3")
  max_me3 <- max(df_me3$mean_signal + df_me3$err, na.rm = TRUE)
  y_max_me3 <- max_me3 + (max_me3 - y_min_me3) * 0.15

  p_me3 <- ggplot(df_me3, aes(x = time, y = mean_signal, fill = time)) +
    geom_col(width = 0.55, color = "black", linewidth = 0.3) +
    geom_errorbar(
      aes(ymin = mean_signal - err, ymax = mean_signal + err),
      width = 0.15, linewidth = 0.6
    ) +
    scale_fill_manual(values = colors_me3) +
    scale_y_continuous(expand = c(0, 0)) + # remove the default gap
    coord_cartesian(ylim = c(y_min_me3, y_max_me3)) + # set the absolute Y axis range (starting at y_min_me3)
    labs(title = "H3K27me3", x = "Time", y = "Signal") +
    theme_classic(base_size = 13) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
      legend.position = "none"
    )

  # 6. Assemble the figure
  combined_plot <- p_me3 + p_ac 
  
  return(combined_plot)
}

# SE (standard error) is used for the error bars by default
p1 <- plot_epigenomic_signals(df1)
p2 <- plot_epigenomic_signals(df2)
p3 <- plot_epigenomic_signals(df3)

p <- p1 / p2 / p3

ggsave("outplots/fig4/sup4_c.clu567.inherit.barplot.pdf", p, width = 6, height = 10, dpi = 300)


# To use SD (standard deviation) for the error bars instead:
p_sd <- plot_epigenomic_signals(df1, error_type = "sd")
print(p_sd)

```


## sup4_d Four quadrant plots, picking the one that best matches expectations
```{r}
clu567.biv.d7.ac <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig4.bivalent/clu567.d7.biv.bin.H3K27ac.CPM.tab")
clu567.biv.d7.me3 <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/fig4.bivalent/clu567.d7.biv.bin.H3K27me3.CPM.tab")


clu567.biv.d7.delta <- data.frame(
    ac_d0d7 = log2(clu567.biv.d7.ac[,5]/clu567.biv.d7.ac[,4]),
    ac_d7d15 = log2(clu567.biv.d7.ac[,6]/clu567.biv.d7.ac[,5]),
    me3_d0d7 = log2(clu567.biv.d7.me3[,5]/clu567.biv.d7.me3[,4]),
    me3_d7d15 = log2(clu567.biv.d7.me3[,6]/clu567.biv.d7.me3[,5])
)

quadrant_data1 <- clu567.biv.d7.delta %>%
  mutate(quadrant = case_when(
    ac_d0d7 > 0 & me3_d0d7 < 0 ~ "ac Up / me3 Down",
    ac_d0d7 > 0 & me3_d0d7 >= 0 ~ "ac Up / me3 Up",
    ac_d0d7 <= 0 & me3_d0d7 < 0 ~ "ac Down / me3 Down",
    ac_d0d7 <= 0 & me3_d0d7 >= 0 ~ "ac Down / me3 Up"
  )) %>%
  group_by(quadrant) %>%
  summarise(count = n()) %>%
  mutate(perc = paste0(round(count/sum(count)*100, 1), "%"))

# Plot: show the distribution density and the quadrant proportions
p1 <- ggplot(clu567.biv.d7.delta, aes(x=ac_d0d7, y=me3_d0d7)) +
  geom_bin2d(bins = 50) + # show the point density in heatmap form
  scale_fill_gradient(low = "lightblue", high = "red") +
  geom_vline(xintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0, linetype="dashed") +
  # Annotate the percentages in the four corners
  annotate("text", x = -2, y = -2, label = paste0(quadrant_data1[1,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = -2, y = 2, label = paste0(quadrant_data1[2,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = -2, label = paste0(quadrant_data1[3,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = 2, label = paste0(quadrant_data1[4,3]), size=6, color="black", fontface="bold") +
  labs(title="",
       x="H3K27ac (d0-d7)", y="H3K27me3 (d7-d15)") +
  theme_bw(base_size = 16)+
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "none"
  )
ggsave("outplots/fig4/sup4_d_1.Quadrant.Proportions.pdf",p1,width=5,height=5,dpi=300)


quadrant_data2 <- clu567.biv.d7.delta %>%
  mutate(quadrant = case_when(
    ac_d0d7 > 0 & me3_d7d15 < 0 ~ "ac Up / me3 Down",
    ac_d0d7 > 0 & me3_d7d15 >= 0 ~ "ac Up / me3 Up",
    ac_d0d7 <= 0 & me3_d7d15 < 0 ~ "ac Down / me3 Down",
    ac_d0d7 <= 0 & me3_d7d15 >= 0 ~ "ac Down / me3 Up"
  )) %>%
  group_by(quadrant) %>%
  summarise(count = n()) %>%
  mutate(perc = paste0(round(count/sum(count)*100, 1), "%"))

# Plot: show the distribution density and the quadrant proportions
p2 <- ggplot(clu567.biv.d7.delta, aes(x=ac_d0d7, y=me3_d7d15)) +
  geom_bin2d(bins = 50) + # show the point density in heatmap form
  scale_fill_gradient(low = "lightblue", high = "red") +
  geom_vline(xintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0, linetype="dashed") +
  # Annotate the percentages in the four corners
  annotate("text", x = -2, y = -2, label = paste0(quadrant_data2[1,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = -2, y = 2, label = paste0(quadrant_data2[2,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = -2, label = paste0(quadrant_data2[3,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = 2, label = paste0(quadrant_data2[4,3]), size=6, color="black", fontface="bold") +
  labs(title="",
       x="H3K27ac (d0-d7)", y="H3K27me3 (d7-d15)") +
  theme_bw(base_size = 16)+
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "none"
  )
ggsave("outplots/fig4/sup4_d_2.Quadrant.Proportions.pdf",p2,width=5,height=5,dpi=300)


quadrant_data3 <- clu567.biv.d7.delta %>%
  mutate(quadrant = case_when(
    ac_d7d15 > 0 & me3_d0d7 < 0 ~ "ac Up / me3 Down",
    ac_d7d15 > 0 & me3_d0d7 >= 0 ~ "ac Up / me3 Up",
    ac_d7d15 <= 0 & me3_d0d7 < 0 ~ "ac Down / me3 Down",
    ac_d7d15 <= 0 & me3_d0d7 >= 0 ~ "ac Down / me3 Up"
  )) %>%
  group_by(quadrant) %>%
  summarise(count = n()) %>%
  mutate(perc = paste0(round(count/sum(count)*100, 1), "%"))

# Plot: show the distribution density and the quadrant proportions
p3 <- ggplot(clu567.biv.d7.delta, aes(x=ac_d7d15, y=me3_d0d7)) +
  geom_bin2d(bins = 50) + # show the point density in heatmap form
  scale_fill_gradient(low = "lightblue", high = "red") +
  geom_vline(xintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0, linetype="dashed") +
  # Annotate the percentages in the four corners
  annotate("text", x = -2, y = -2, label = paste0(quadrant_data3[1,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = -2, y = 2, label = paste0(quadrant_data3[2,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = -2, label = paste0(quadrant_data3[3,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = 2, label = paste0(quadrant_data3[4,3]), size=6, color="black", fontface="bold") +
  labs(title="",
       x="H3K27ac (d0-d7)", y="H3K27me3 (d7-d15)") +
  theme_bw(base_size = 16)+
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "none"
  )
ggsave("outplots/fig4/sup4_d_3.Quadrant.Proportions.pdf",p3,width=5,height=5,dpi=300)


quadrant_data4 <- clu567.biv.d7.delta %>%
  mutate(quadrant = case_when(
    ac_d7d15 > 0 & me3_d7d15 < 0 ~ "ac Up / me3 Down",
    ac_d7d15 > 0 & me3_d7d15 >= 0 ~ "ac Up / me3 Up",
    ac_d7d15 <= 0 & me3_d7d15 < 0 ~ "ac Down / me3 Down",
    ac_d7d15 <= 0 & me3_d7d15 >= 0 ~ "ac Down / me3 Up"
  )) %>%
  group_by(quadrant) %>%
  summarise(count = n()) %>%
  mutate(perc = paste0(round(count/sum(count)*100, 1), "%"))

# Plot: show the distribution density and the quadrant proportions
p4 <- ggplot(clu567.biv.d7.delta, aes(x=ac_d7d15, y=me3_d7d15)) +
  geom_bin2d(bins = 50) + # show the point density in heatmap form
  scale_fill_gradient(low = "lightblue", high = "red") +
  geom_vline(xintercept = 0, linetype="dashed") +
  geom_hline(yintercept = 0, linetype="dashed") +
  # Annotate the percentages in the four corners
  annotate("text", x = -2, y = -2, label = paste0(quadrant_data4[1,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = -2, y = 2, label = paste0(quadrant_data4[2,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = -2, label = paste0(quadrant_data4[3,3]), size=6, color="black", fontface="bold") +
  annotate("text", x = 2, y = 2, label = paste0(quadrant_data4[4,3]), size=6, color="black", fontface="bold") +
  labs(title="",
       x="H3K27ac (d0-d7)", y="H3K27me3 (d7-d15)") +
  theme_bw(base_size = 16)+
  theme(
    axis.text = element_text(color = "black",size = 16),
    legend.position = "none"
  )
ggsave("outplots/fig4/sup4_d_4.Quadrant.Proportions.pdf",p4,width=5,height=5,dpi=300)
```

# =========================================================
## sup4.e Expression and GO enrichment analysis of bivalently modified genes and ordinary genes
```{r}

roi.proximal.tg <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/chromHMM/DiffCluster/output/roi_proximal.tg.tsv",header = T,sep = "\t")
clu567.d7.bivalent$name <- sapply(strsplit(clu567.d7.bivalent$peak,"_",fixed = T),"[",2)
l1 <- roi.proximal.tg[roi.proximal.tg$name %in% clu567.d7.bivalent[clu567.d7.bivalent$biv !=0,"name"],"target_gene_id"] %>% unique()
l1 <- intersect(l1,tg.keeped)
l2 <- roi.proximal.tg[roi.proximal.tg$name %in% clu567.d7.bivalent[clu567.d7.bivalent$biv ==0,"name"],"target_gene_id"] %>% unique()
l2 <- intersect(l2,tg.keeped)



l1.tpm <- log2(all.tpm[l1[!l1 %in% intersect(l1,l2)],]+1)
l2.tpm <- log2(all.tpm[l2[!l2 %in% intersect(l1,l2)],]+1)

l1.tpm <- log2(all.tpm[l1,]+1)
l2.tpm <- log2(all.tpm[l2,]+1)

boxplot(l1.tpm)
boxplot(l2.tpm)

t.test(l1.tpm$MSC,l1.tpm$ADI_7d,paired = T)
t.test(l1.tpm$MSC,l1.tpm$ADI_15d,paired = T)

t.test(l2.tpm$MSC,l2.tpm$ADI_7d,paired = T)
t.test(l2.tpm$MSC,l2.tpm$ADI_15d,paired = T)


isbiv.tpm.df <- data.frame(
    time=factor(c(rep("d0",nrow(l1.tpm)),rep("d7",nrow(l1.tpm)),rep("d15",nrow(l1.tpm))),levels=c("d0","d7","d15")),
    tpm=c(l1.tpm$MSC,l1.tpm$ADI_7d,l1.tpm$ADI_15d)
)

p <- ggplot(isbiv.tpm.df, aes(x = time, y = tpm, fill = time,color=time)) +
stat_boxplot(geom = "errorbar",
               width = 0.2,
               #size = 1.0,
               position = position_dodge(0.8)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1, alpha = 1) +
  scale_fill_manual(values = c("#dba880","#dba880","#dba880")) +
  scale_color_manual(values = c("black","black","black")) +
  labs(title = "", x = "", y = "log2(TPM+1)") +
  theme_classic(base_size = 16) +
  theme(legend.position = "up", # show the legend at the bottom
        legend.direction = "horizontal",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text = element_text(color = "black"),
        panel.grid.major.x = element_blank())

ggsave("outplots/fig4/fig4_g.isbiv.boxplot.pdf",p,width=7,height=7,dpi=300)

nobiv.tpm.df <- data.frame(
    time=factor(c(rep("d0",nrow(l2.tpm)),rep("d7",nrow(l2.tpm)),rep("d15",nrow(l2.tpm))),levels=c("d0","d7","d15")),
    tpm=c(l2.tpm$MSC,l2.tpm$ADI_7d,l2.tpm$ADI_15d)
)

p <- ggplot(nobiv.tpm.df, aes(x = time, y = tpm, fill = time,color=time)) +
stat_boxplot(geom = "errorbar",
               width = 0.2,
               #size = 1.0,
               position = position_dodge(0.8)) +
  geom_boxplot(outlier.shape = 21, outlier.size = 1, alpha = 1) +
  scale_fill_manual(values = c("#94d2ef","#94d2ef","#94d2ef")) +
  scale_color_manual(values = c("black","black","black")) +
  labs(title = "", x = "", y = "log2(TPM+1)") +
  theme_classic(base_size = 16) +
  theme(legend.position = "up", # show the legend at the bottom
        legend.direction = "horizontal",
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        axis.text = element_text(color = "black"),
        panel.grid.major.x = element_blank())

ggsave("outplots/fig4/fig4_g.nobiv.boxplot.pdf",p,width=7,height=7,dpi=300)



clu5.isbiv.go <- clusterProfiler::enrichGO(
      gene = ref[ref$gene_id %in% intersect(setdiff(l1,l2),tg.keeped),"gene_name"],
      universe = ref[ref$gene_id %in% tg.keeped,"gene_name"],
      OrgDb = org.Hs.eg.db::org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "BP",
      pAdjustMethod = "BH",
      pvalueCutoff = 0.1,
      qvalueCutoff = 0.1,
      #minGSSize = 10,
      #maxGSSize = 500,
      readable = TRUE
    ) %>% as.data.frame()

clu5.nobiv.go <- clusterProfiler::enrichGO(
      gene = ref[ref$gene_id %in% intersect(setdiff(l2,l1),tg.keeped),"gene_name"],
      universe = ref[ref$gene_id %in% tg.keeped,"gene_name"],
      OrgDb = org.Hs.eg.db::org.Hs.eg.db,
      keyType = "SYMBOL",
      ont = "BP",
      pAdjustMethod = "BH",
      pvalueCutoff = 0.1,
      qvalueCutoff = 0.1,
      #minGSSize = 15,
      #maxGSSize = 300,
      readable = TRUE
) %>% as.data.frame()



## The corresponding enrichment plots are drawn next
is.biv.go <- read.table(file = "output/fig4.bivalent/isbiv.vs.nobiv.GO.output.tsv",header = T,sep = "\t")


# 2. Process the data
# Compute -log10(FDR) as the height metric (standard practice for enrichment plots: the smaller the FDR, the longer the bar)
# To plot the raw FDR values instead, change -log10(FDR) in next_line to -FDR
df <- is.biv.go %>%
  mutate(
    FDR = -log10(FDR), # significance score
    category = factor(category, levels = c("isbiv", "nobiv")) # ensure the order
  ) %>%
  # Sort by FDR from small to large within each group (i.e. score from large to small) so that the most significant terms appear at the top
  arrange(category, FDR) %>%
  mutate(GO = str_wrap(GO, width = 30))  %>%
  mutate(GO = factor(GO, levels = rev(unique(GO))))


p <- ggplot(df, aes(y = FDR, x = GO, fill = category)) +
  geom_bar(stat="identity")+coord_flip() +
  scale_fill_manual(values = c("#dba880","#94d2ef")) +
  scale_color_manual(values = c("black","black","black")) +
  labs(title = "", x = "", y = "") +
  theme_classic(base_size = 16) +
  theme(legend.position = "up", # show the legend at the bottom
        legend.direction = "horizontal",
        plot.title = element_text(hjust = 0.5, face = "bold"),
        axis.text.y = element_text(color = "black",size = 28),

        panel.grid.major.x = element_blank())

ggsave("outplots/fig4/fig4_g.go.pdf",p,width=10,height=12,dpi=300)
```


