
<!--
Reproduces GWAS panels of main Figure 5: SNP enrichment/frequency analyses over the CRR classes. Inputs (examples): GWAS workspace Rdata (input/5.8.Rdata), roi_info.split.bed, gwas_expanded_set_with_pos.tsv and control/peak BED files. Outputs: outplots/fig5*.pdf.
--> 
# fig5-GWAS


```{r}
load("input/5.8.Rdata")


```



fig5.a Overall SNPs-per-1kb levels across the clusters
```{r}
roi_info.split <- read.table(file = "~/Project/GWAS/input/roi_info.split.bed",header = T,sep = "\t")
catelog_ref_snps <- read.table(file = "~/Project/GWAS/plink/gwas_expanded_set_with_pos.tsv",header = T)
catelog_ref_snps <- data.frame(
    chr=paste0('chr',ifelse(catelog_ref_snps$chr == 23,"X",catelog_ref_snps$chr)),
    start=catelog_ref_snps$position,
    end=catelog_ref_snps$position+1,
    snpid=catelog_ref_snps$snpid
)
l1 <- bt.intersect(a=catelog_ref_snps,b=roi_info.split[roi_info.split$cluster == 3,],wa=T)
l2 <- bt.intersect(a=catelog_ref_snps,b=roi_info.split[roi_info.split$cluster == 5,],wa=T)

snps.per1kb.results <- data.frame(matrix(nrow=9,ncol=2))

for (i in 1:7){
    inter <- bt.intersect(a=catelog_ref_snps,b=roi_info.split[roi_info.split$cluster == i,],wa=T) %>% unique
    len <- sum(roi_info.split[roi_info.split$cluster == i,"width"])/1000
    snps.per1kb.results[i,2] <- nrow(inter)/len
    snps.per1kb.results[i,1] <- paste0("cluster",i)
}

all.control <- read.table("~/Project/chromHMM/DiffCluster/output/all.roi.bg.bed", sep = "\t", stringsAsFactors = FALSE)
all.control <- bt.sort(i=unique(all.control),faidx="~/reference/hg19.chrom.sizes")
all.control <- all.control[all.control$V1 %in% chrom,]

l1 <- bt.intersect(a=catelog_ref_snps,b=all.control,wa=T) %>% unique
snps.per1kb.results[8,2] <- nrow(l1)/(sum(all.control$V3-all.control$V2)/1000)
snps.per1kb.results[8,1] <- "Control"

all.ac <- read.table(file = "~/Project/ChIP/H3K27ac/narrow/all.H3K27ac.peaks.merged.bed")
l2 <- bt.intersect(a=catelog_ref_snps,b=all.ac,wa=T) %>% unique

snps.per1kb.results[9,1] <- "Enhancer"
snps.per1kb.results[9,2] <- nrow(l2)/(sum(all.ac$V3-all.ac$V2)/1000)
colnames(snps.per1kb.results) <- c("category","ratio")



# 3. Set the order: control and ac first, then cluster1-7
target_order <- c("Control", "Enhancer", paste0("cluster", 1:7))
snps.per1kb.results$category <- factor(snps.per1kb.results$category, levels = target_order)

# 4. Define the color mapping
# grey for control and a bright green/cyan for ac (distinct from the blue-red palette)
my_colors <- c(
  "Control" = "#999999",     # grey
  "Enhancer" = "#e31a1c",     # spring green (bright and distinct)
  "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)

# 5. Plot
p <- ggplot(snps.per1kb.results, aes(x = category, y = ratio, fill = category)) +
  # Set the bar borders to black or none; a little transparency here highlights AC
  geom_bar(stat = "identity", width = 0.7, color = "white", alpha = 0.9) +
  # Further adjust the transparency of the clusters (optional)
  aes(alpha = ifelse(category %in% c("all.control", "all.H3K27ac"), 1.0, 0.7)) +
  scale_alpha_identity() +
  scale_fill_manual(values = my_colors) +
  # Key settings:
  # expand = c(0,0) keeps the bottoms of the bars flush with the X axis
  # breaks sets the tick marks
  scale_y_continuous(
    expand = c(0, 0), 
    breaks = seq(0.2, 0.32, by = 0.02)
  ) +
  # coord_cartesian sets the visible range, starting at 0.2
  coord_cartesian(ylim = c(0.2, 0.32)) +
  theme_classic(base_size = 16) + # the classic theme makes the X axis line clearly visible
  labs(
    title = "",
    x = NULL,
    y = "GWAS SNPs per 1kb"
  ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y = element_text(color = "black"),
    axis.line.x = element_line(color = "black"), # ensure the X axis line is visible
    legend.position = "none"
  )

print(p)
ggsave(file = "outplots/fig5_a.snps.per.1kb.barplot.pdf",p,width=8,height=4,dpi=300)


```



fig5.b SNP enrichment relative to background controls
```{r}

gwas.malist <- dir("~/Project/GWAS/ma")
all.enrich.result <- data.frame()

    all.control <- read.table("~/Project/chromHMM/DiffCluster/output/all.roi.bg.bed", sep = "\t", stringsAsFactors = FALSE)
    all.control <- bt.sort(i=unique(all.control),faidx="~/reference/hg19.chrom.sizes")
    all.control <- all.control[all.control$V1 %in% chrom,]
    all.ac <- read.table(file = "~/Project/ChIP/H3K27ac/narrow/all.H3K27ac.peaks.merged.bed",header=  F)
    all.ac.control <- read.table(file = "~/Project/ChIP/H3K27ac/narrow/all.ac.bg.bed")
    all.ac.control <- all.ac.control[all.ac.control$V1 %in% chrom,]
    all.ac.control <- bt.sort(i=all.ac.control,faidx="~/reference/hg19.chrom.sizes")


for (i in 1:length(gwas.malist)){
    ma.name <- gwas.malist[i]
    samplename <- unlist(strsplit(ma.name,".ma",fixed = T))[1]
    gwas <- fread(file = paste0('~/Project/GWAS/ma/',ma.name))
    #gwas_p <- gwas[gwas$P <= 5e-8,]
    gwas_clump <- read.table(file = paste0('~/Project/GWAS/clump/',samplename,'_clump_lead_snps.txt'))[,1]
    gwas_p <- gwas[gwas$SNP %in% gwas_clump,]
    gwas_p <- data.frame(chr=paste0('chr',gwas_p$CHR),start=gwas_p$BP,end=gwas_p$BP+1,rsid=gwas_p$SNP)
    enrich.result <- data.frame(matrix(nrow=9,ncol=5))
    colnames(enrich.result) <- c("GWAS","cluster","OR","pval","effect_ratio")
    enrich.result$GWAS <- unlist(strsplit(ma.name,".",fixed = T))[1]
    enrich.result$cluster <- c(paste0('cluster',1:7),"all.input","all.ac")
    for (j in 1:7){
        input <- roi_info.split[roi_info.split$cluster == j,]
        input_control <- read.table(file = paste0('~/Project/chromHMM/DiffCluster/output/clusters/controls/cluster',j,'.bg.bed'))
        input_control <- input_control[input_control$V1 %in% chrom,]
        inter1 <- bt.intersect(a=input,b=gwas_p,wa=T) %>% unique()
        inter2 <- bt.intersect(a=input_control,b=gwas_p,wa=T) %>% unique()
        print(matrix(c(nrow(inter1),nrow(input)-nrow(inter1),nrow(inter2),nrow(input_control)-nrow(inter2)),nrow=2,ncol=2))
        fisher <- fisher.test(matrix(c(nrow(inter1),nrow(input)-nrow(inter1),nrow(inter2),nrow(input_control)-nrow(inter2)),nrow=2,ncol=2))
        or <- fisher$estimate
        pval <- fisher$p.value
        enrich.result[j,3:5] <- c(or,pval,nrow(inter1)/nrow(input))
    }
    all.input <- roi_info.split
    ## all.input.H3K27me3.ROIs
    inter1 <- bt.intersect(a=all.input,b=gwas_p,wa=T) %>% unique()
    inter2 <- bt.intersect(a=all.control,b=gwas_p,wa=T) %>% unique()
    fisher <- fisher.test(matrix(c(nrow(inter1),nrow(all.input)-nrow(inter1),nrow(inter2),nrow(all.control)-nrow(inter2)),nrow=2,ncol=2))
    or <- fisher$estimate
    pval <- fisher$p.value
    enrich.result[8,3:5] <- c(or,pval,nrow(inter1)/nrow(all.input))
    ## all.H3K27ac peaks enhancers
    inter1 <- bt.intersect(a=all.ac,b=gwas_p,wa=T) %>% unique()
    inter2 <- bt.intersect(a=all.ac.control,b=gwas_p,wa=T) %>% unique()
    fisher <- fisher.test(matrix(c(nrow(inter1),nrow(all.ac)-nrow(inter1),nrow(inter2),nrow(all.ac.control)-nrow(inter2)),nrow=2,ncol=2))
    or <- fisher$estimate
    pval <- fisher$p.value
    enrich.result[9,3:5] <- c(or,pval,nrow(inter1)/nrow(all.ac))
    enrich.result$FDR <- p.adjust(enrich.result$pval)    
    
    all.enrich.result <- rbind(all.enrich.result,enrich.result)
}


all.enrich.result$logp <- -log10(all.enrich.result$pval)
all.enrich.result$logp <- ifelse(all.enrich.result$logp > -log10(5e-8),-log10(5e-8),all.enrich.result$logp)



# Load the required libraries
library(ggplot2)
library(dplyr)

## Assume your data frame is named all.enrich.result and contains at least the following columns:
## GWAS, cluster, OR, logp
## The example data below are for reference (skip them if you already have all.enrich.result)
# all.enrich.result <- data.frame(
#   GWAS   = c("ASAT", "ASAT", "ASAT", "ASAT", "ASAT", "ASAT"),
#   cluster = paste0("cluster", 1:6),
#   OR     = c(0, 5.650652, 1.47458, 0, 0, 0),
#   logp   = c(0.4550025, 1.3569231, 0.1890116, 0, 0, 0)
# )

## Preprocessing: set the significance threshold ---------------------------------------
logp_cutoff <- 1.301   # approximately -log10(0.05)

plot.data <- all.enrich.result %>%
  mutate(
    ## Whether the result is significant (used for the coloring logic)
    sig = logp >= logp_cutoff,
    
    ## Value used to compute the color intensity:
    ## For significant terms the intensity follows logp; non-significant terms are set to NA (shown in grey later)
    logp_for_color = ifelse(sig, logp, NA),
    
    ## Direction: OR > 1 is recorded as +1 (red), OR < 1 as -1 (blue), and OR == 1 or 0 as 0
    direction = case_when(
      OR > 1  ~  1,
      OR < 1  ~ -1,
      TRUE    ~  0
    ),
    
    ## Variable used for the continuous color scale:
    ## Rationale: significant & OR > 1 gives a positive value, significant & OR < 1 a negative value, the rest NA (grey)
    color_val = ifelse(sig & direction != 0, direction * logp_for_color, NA_real_)
  )

## Read the list of sample names used for plotting
final.samples <- fread(file = "final.intpu.samplename.3.24.txt",header = F,sep = "\t") %>% as.data.frame

plot.data <- plot.data[plot.data$GWAS %in% final.samples[,1],]
plot.data$GWAS <- final.samples[match(plot.data$GWAS,final.samples[,1]),3]
plot.data$GWAS <- factor(plot.data$GWAS,levels=rev(final.samples[,3]))
plot.data$cluster <- factor(plot.data$cluster,c("all.ac","all.input",paste0('cluster',1:7)))

plot.data <- plot.data %>%
  mutate(
    # Column grouping: the first two columns are GroupA and the rest GroupB
    # Adjust levels(cluster)[1:2] here to match your actual column names
    col_group = ifelse(cluster %in% levels(factor(cluster))[1:2], "Col_Part1", "Col_Part2"),

  )

## Build the color gradient ---------------------------------------------------
## A continuous blue-white-red color scale is used here:
##   negative (blue) <- 0 (white) -> positive (red)
## but non-significant cells are set to grey

# Compute the range of color_val (ignoring NA)
color_range <- range(plot.data$color_val, na.rm = TRUE)
# This also prevents errors when no enrichment/depletion is significant in your data
if (!is.finite(color_range[1])) {
  color_range <- c(-1, 1)
}

## Plotting ----------------------------------------------------------
p <- ggplot(plot.data, aes(x = cluster, y = GWAS)) +
  geom_tile(aes(fill = color_val),
            color = "grey80", size = 0.3) +
      
  # --- key change: adding facets ---
  facet_grid(
    . ~ col_group, 
    scales = "free", 
    space = "free"
  ) +
  
  ## Set the colors manually:
  ## Non-significant cells (color_val = NA) are set through na.value in scale_fill_gradient2
  scale_fill_gradient2(
    name = expression(paste("P value")),
    low = "#4575B4",      # blue
    mid = "white",        # neutral
    high = "#D73027",     # red
    midpoint = 0,
    limits = color_range, # keeps the blue-red scale symmetric
    na.value = "white"   # non-significant: grey/white
  ) +
  
  ## Theme and axis styling
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  
  labs(
    x = "",
    y = "",
    title = ""
  ) +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1,color="black"),
    axis.text.y = element_text(color="black",size=16),
    axis.title.x = element_text(margin = margin(t = 8)),
    axis.title.y = element_text(margin = margin(r = 8)),
    legend.position = "right",
    legend.title = element_text(face = "bold")
  )
print(p)
ggsave(file = "outplots/fig5_b.gwas.clumped.fisher.pheat.pdf",width=8,height=8,dpi=300)
```


fig5.c Overall and per-cluster S-LDSC heritability analysis
```{r}

total_input_dir <- "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/GWAS/ldsc/H3K27-S-LDSC"
clu_input_dir <- "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/GWAS/ldsc/S-LDSC"
samplelist <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/GWAS/input.GWAS.samplename",header = F)[,1]

# Two parts: (1) LDSC heritability enrichment for all ROIs and enhancers, (2) heritability by cluster

total.ldsc.result <- data.frame()
for (i in 1:length(samplelist)){
    input <- tail(read.table(file = paste0(total_input_dir,'/',samplelist[i],'.results'),header = T,sep = "\t"),2)
    input$Enrichment_FDR <- p.adjust(input$Enrichment_p)
    input$Enrichment_FDR <- input$Enrichment_p
    output <- data.frame(cluster=c("All ROIs","Enhancer"),GWAS=samplelist[i],pval=input$Enrichment_FDR,estimate=input$Enrichment)
    total.ldsc.result <- rbind(total.ldsc.result,output)
}


clu.ldsc.result <- data.frame()
for (i in 1:length(samplelist)){
    input <- tail(read.table(file = paste0(clu_input_dir,'/',samplelist[i],'.results'),header = T,sep = "\t"),7)
    input$Enrichment_FDR <- p.adjust(input$Enrichment_p)
    input$Enrichment_FDR <- input$Enrichment_p
    output <- data.frame(cluster=c(paste0('cluster',1:7)),GWAS=samplelist[i],pval=input$Enrichment_FDR,estimate =input$Enrichment)
    clu.ldsc.result <- rbind(clu.ldsc.result,output)
}

all.ldsc.result <- rbind(total.ldsc.result,clu.ldsc.result)


all.ldsc.result$logp <- -log10(all.ldsc.result$pval)

# Set a sensible upper bound for logp so that extremely small P values do not stretch the color scale and hide ordinary signals
# 10 or 20 is a reasonable choice; 5e-8 (7.3) also works
max_logp <- -log10(5e-8) 
all.ldsc.result$logp_capped <- ifelse(all.ldsc.result$logp > max_logp, max_logp, all.ldsc.result$logp)

logp_cutoff <- 1.301   # P < 0.05

plot.data <- all.ldsc.result %>%
  mutate(
    # Determine the direction: Enrichment > 1 is positive, < 1 is negative
    direction = ifelse(estimate > 1, 1, -1),
    
    # Continuous variable used for plotting: signed -log10(P)
    # so positive enrichment is red and negative enrichment (depletion) is blue
    color_val = direction * logp_capped,
    
    # Decide whether to add an asterisk
    sig_label = ifelse(logp >= logp_cutoff, "*", "")
  )

# 2. Filter and order the GWAS traits
final.samples <- fread(file = "final.intpu.samplename.3.24.txt", header = F, sep = "\t") %>% as.data.frame()
plot.data <- plot.data[plot.data$GWAS %in% final.samples[,1],]
plot.data$GWAS <- final.samples[match(plot.data$GWAS,final.samples[,1]),3]
plot.data$GWAS <- factor(plot.data$GWAS,levels=rev(final.samples[,3]))

plot.data$cluster <- factor(plot.data$cluster,c("Enhancer","All ROIs",paste0('cluster',1:7)))


# 3. Determine the symmetric color scale range
# Find the largest absolute value so that 0 sits exactly in the middle of the palette (white)
limit_val <- max(abs(plot.data$color_val), na.rm = TRUE)

## 4. Plotting ----------------------------------------------------------
plot.data <- plot.data %>%
  mutate(
    # Column grouping: the first two columns are GroupA and the rest GroupB
    # Adjust levels(cluster)[1:2] here to match your actual column names
    col_group = ifelse(cluster %in% levels(factor(cluster))[1:2], "Col_Part1", "Col_Part2"),

  )

# 2. Plotting
p <- ggplot(plot.data, aes(x = cluster, y = GWAS)) +
  geom_tile(aes(fill = color_val), color = "grey90", size = 0.1) +
  geom_text(aes(label = sig_label), color = "black", size = 5, vjust = 0.7) +
  
  # --- key change: adding facets ---
  facet_grid(
    . ~ col_group, 
    scales = "free", 
    space = "free"
  ) +
  
  scale_fill_gradientn(
    name = "P value",
    colors = c("#4575B4", "#74ADD1", "#ABD9E9", "white", "#FDAE61", "#F46D43", "#D73027"),
    values = scales::rescale(c(log10(5e-8), -1.301, -1.300, 0, 1.300, 1.301, -log10(5e-8))),
    limits = c(log10(5e-8), -log10(5e-8)),
    breaks = c(log10(5e-8), -1.301, 0, 1.301, -log10(5e-8)),
    guide = guide_colorbar(frame.colour = "black", ticks.colour = "black", barheight = 15)
  ) +
  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0)) +
  
  labs(x = "", y = "", title = "") +
  theme_minimal(base_size = 16) +
  theme(
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, color = "black"),
    axis.text.y = element_text(color = "black"),
    legend.position = "right",
    
    # --- key styling: hide the facet labels and adjust the spacing ---
    strip.background = element_blank(), # hide the grey box behind the facets
    strip.text = element_blank(),       # hide the facet text labels (Col_Part1 etc.)
    panel.spacing.x = unit(0.5, "cm"),  # horizontal spacing (separates columns)
    panel.spacing.y = unit(0.5, "cm")   # vertical spacing (separates rows)
  )

print(p)
ggsave(file = "outplots/fig5_c.s-ldsc.pheat.pdf",width=8,height=8,dpi=300)


```


fig5_d,e S-LDSC heritability enrichment during differentiation
```{r}
library(dplyr)
library(stringr)

calculate_tau_star <- function(
    results_path,
    log_path,
    M = 5961159
) {
  
  # ------------------------------------------------------------------
  # 1. Check and read the results file
  # ------------------------------------------------------------------
  
  if (!file.exists(results_path)) {
    stop("Results file not found: ", results_path)
  }
  
  results <- read.table(
    results_path,
    header = TRUE,
    sep = "\t",
    stringsAsFactors = FALSE,
    check.names = TRUE
  )
  
  required_cols <- c(
    "Category",
    "Prop._SNPs",
    "Enrichment",
    "Enrichment_p",
    "Coefficient",
    "Coefficient_std_error"
  )
  
  missing_cols <- setdiff(required_cols, colnames(results))
  
  if (length(missing_cols) > 0) {
    stop(
      "The results file is missing the following columns: ",
      paste(missing_cols, collapse = ", ")
    )
  }
  
  # ------------------------------------------------------------------
  # 2. Extract the observed-scale h2 from the log file
  # ------------------------------------------------------------------
  
  if (!file.exists(log_path)) {
    stop("Log file not found: ", log_path)
  }
  
  log_content <- readLines(log_path, warn = FALSE)
  
  h2_line <- grep(
    "^Total Observed scale h2:",
    log_content,
    value = TRUE
  )
  
  if (length(h2_line) != 1) {
    stop(
      "Could not uniquely locate 'Total Observed scale h2:'.\n",
      "Found ", length(h2_line), " lines; please check the log file: ",
      log_path
    )
  }
  
  # Extract only the first number after h2: so that the 2 inside h2 is not picked up
  h2_total <- as.numeric(
    sub(
      "^Total Observed scale h2:[[:space:]]*([-+0-9.eE]+).*$",
      "\\1",
      h2_line
    )
  )
  
  if (
    length(h2_total) != 1 ||
    !is.finite(h2_total) ||
    h2_total <= 0
  ) {
    stop(
      "The extracted total heritability is invalid: ",
      paste(h2_total, collapse = ", "),
      "\nCorresponding log line: ",
      h2_line
    )
  }
  
  # ------------------------------------------------------------------
  # 3. Compute tau*
  # ------------------------------------------------------------------
  
  results_processed <- results %>%
    mutate(
      # Only valid for 0/1 binary annotations
      sd_annot = sqrt(Prop._SNPs * (1 - Prop._SNPs)),
      
      scaling_factor = M * sd_annot / h2_total,
      
      tau_star = scaling_factor * Coefficient,
      
      # SE should always be non-negative
      tau_star_se = abs(scaling_factor) * Coefficient_std_error,
      
      # The Z-score is unchanged by standardization
      tau_star_z = Coefficient / Coefficient_std_error,
      
      tau_star_p = 2 * pnorm(
        abs(tau_star_z),
        lower.tail = FALSE
      ),
      
      h2_total = h2_total,
      M_total = M
    ) %>%
    select(
      Category,
      Prop._SNPs,
      Enrichment,
      Enrichment_p,
      Coefficient,
      Coefficient_std_error,
      sd_annot,
      h2_total,
      M_total,
      tau_star,
      tau_star_se,
      tau_star_z,
      tau_star_p
    )
  
  return(results_processed)
}

# --- Usage example ---
# results_file <- "BMI.results"
# log_file <- "BMI.log"
# final_df <- calculate_tau_star(results_file, log_file)

# Inspect the results for the clusters you are interested in
# print(head(final_df %>% filter(grepl("Clu", Category))))

clu_tau_results <- data.frame()
samplelist <- read.table(file = "period.sldsc.inputsamplelist")[,1]

for (i in 1:length(samplelist)){
    all_path <- "~/Project/GWAS/ldsc/S-LDSC"
    result_path_d0 <- "~/Project/GWAS/ldsc/S-LDSC-period/d0"
    result_path_d7 <- "~/Project/GWAS/ldsc/S-LDSC-period/d7"
    result_path_d15 <- "~/Project/GWAS/ldsc/S-LDSC-period/d15"
    l0 <- tail(calculate_tau_star(paste0(all_path,'/',samplelist[i],'.results'), paste0(all_path,'/',samplelist[i],'.log')),7)
    l1 <- tail(calculate_tau_star(paste0(result_path_d0,'/',samplelist[i],'.results'), paste0(result_path_d0,'/',samplelist[i],'.log')),7)
    l2 <- tail(calculate_tau_star(paste0(result_path_d7,'/',samplelist[i],'.results'), paste0(result_path_d7,'/',samplelist[i],'.log')),7)
    l3 <- tail(calculate_tau_star(paste0(result_path_d15,'/',samplelist[i],'.results'), paste0(result_path_d15,'/',samplelist[i],'.log')),7)
    output <- data.frame(
        GWAS=samplelist[i],
        cluster=paste0('cluster',1:7),
        period=c(rep("ALL",7),rep("d0",7),rep("d7",7),rep("d15",7)),
        tau_star=c(l0$tau_star,l1$tau_star,l2$tau_star,l3$tau_star),
        tau_star_se=c(l0$tau_star_se,l1$tau_star_se,l2$tau_star_se,l3$tau_star_se),
        tau_star_p=c(l0$tau_star_p,l1$tau_star_p,l2$tau_star_p,l3$tau_star_p)
    )
    clu_tau_results <- rbind(clu_tau_results,output)
}
#write.table(clu_tau_results,file = "output/clu_tau_results.tsv",row.names = F,col.names = T,sep = "\t",quote = F)


# 2. Data preprocessing
# Define colors
cluster_colors <- c(
    "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)

clu_tau_results <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/GWAS/ldsc/H3K27me3_period_tau_results/clu_tau_results.tsv",header = T,sep = "\t")
clu_tau_results <- clu_tau_results[clu_tau_results$GWAS %in% read.table("period.sldsc.inputsamplelist")[,1],]

plot_data <- clu_tau_results %>%
  # Ensure the period order
  mutate(period = factor(period, levels = c("ALL", "d0", "d7", "d15"))) %>%
  # Logic 1: check whether any period is significant for each cluster within each GWAS
  group_by(GWAS, cluster) %>%
  mutate(any_sig = any(tau_star_p <= 0.05)) %>%
  ungroup() %>%
  # Logic 2: assign colors according to any_sig, falling back to grey when not significant
  mutate(display_color = ifelse(any_sig, cluster, "NotSig")) %>%
  # Logic 3: decide whether the point is filled (p <= 0.05 is filled)
  mutate(is_solid = ifelse(tau_star_p <= 0.05, "Solid", "Hollow"))

# 3. Plotting
pd <- position_dodge(width = 0.7)
p <- ggplot(plot_data, aes(x = period, y = tau_star, group = cluster, color = display_color)) +
  # Draw the error bars (1.96 * SE)
  geom_errorbar(aes(ymin = tau_star - 1.96 * tau_star_se, 
                    ymax = tau_star + 1.96 * tau_star_se), 
                width = 0.4, position = pd, alpha = 0.8) +
  # Draw the connecting lines: only d0, d7 and d15 are linked, with ALL filtered out
  geom_line(data = filter(plot_data, period != "ALL"), 
            position = pd, size = 0.8) +
  # Draw the points: filled or open depending on significance
  # shape 21 is a fillable circle and 16 is a solid circle
  geom_point(aes(shape = is_solid, fill = display_color), 
             position = pd, size = 2.5, stroke = 1) +
  # Faceting: 2 rows by 5 columns
  facet_wrap(~GWAS, nrow = 2, ncol = 5, scales = "free_y") +
  # Set the colors: the seven cluster colors plus grey
  scale_color_manual(values = c(cluster_colors, "NotSig" = "gray80")) +
  scale_fill_manual(values = c(cluster_colors, "NotSig" = "gray80")) +
  # Set the shapes: Solid maps to a filled point (21 with fill) and Hollow to an open point (21 unfilled/white fill)
  scale_shape_manual(values = c("Solid" = 21, "Hollow" = 1)) +
  # Reference line: y = 0
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.3) +
  # Theme tweaks
  theme_bw(base_size = 16) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Forest Plot of tau_star by GWAS and Cluster",
       x = "Period",
       y = "tau_star",
       color = "Cluster Status",
       shape = "Significance (p < 0.05)")

# 4. Print the plot
print(p)


# Save the plot if needed
ggsave("outplots/sup5_a_GWAS_tar_star_Forest_Plot.pdf", width = 15, height = 8)


target_gwas <- c("WHRadjBMI", "logTG", "IWB", "Liver_fat")

library(patchwork) 

# 2. Define the core plotting function
# This function filters the data, computes a symmetric Y axis and draws the forest plot
draw_single_gwas <- function(data, target_gwas_name) {
  
  # A. Data preprocessing
    cluster_colors <- c(
    "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)
  cluster_input_list <- list(
    WHRadjBMI = c("cluster2","cluster3","cluster5"),
    IWB = c("cluster1","cluster5"),
    Liver_fat = c("cluster1","cluster3"),
    FIadj = c("cluster5","cluster7")
  )
  gwas_data <- data %>%
    filter(GWAS == target_gwas_name) %>%
    mutate(period = factor(period, levels = c("ALL", "d0", "d7", "d15"))) %>%
    # Keep only the clusters with a significant signal in this GWAS
    group_by(cluster) %>%
    filter(cluster %in% cluster_input_list[[target_gwas_name]]) %>%
    ungroup() %>%
    # Fill logic: significant points are filled and non-significant ones are open
    mutate(point_fill = ifelse(tau_star_p <= 0.05, cluster, "Hollow"))
  
  # B. Compute the symmetric Y axis range
  # Find the largest absolute value reached by the error bars
  max_abs_y <- gwas_data %>%
    summarise(val = max(c(abs(tau_star + 1.96 * tau_star_se), 
                          abs(tau_star - 1.96 * tau_star_se)), na.rm = TRUE)) %>%
    pull(val)
  
  y_limit <- max_abs_y * 1.1 # leave a 10% margin
  
  # C. Plotting
  pd <- position_dodge(width = 0.5)
  
  p <- ggplot(gwas_data, aes(x = period, y = tau_star, group = cluster, color = cluster)) +
    # Forest-plot error bars
    geom_errorbar(aes(ymin = tau_star - 1.96 * tau_star_se, 
                      ymax = tau_star + 1.96 * tau_star_se), 
                  width = 0.2, position = pd) +
    # Connecting lines (ALL excluded)
    geom_line(data = filter(gwas_data, period != "ALL"), position = pd, size = 0.8) +
    # Points (shape 21 supports fill)
    geom_point(aes(fill = point_fill), shape = 21, size = 3, stroke = 1, position = pd) +
    # Symmetric Y axis
    scale_y_continuous(limits = c(-y_limit, y_limit)) +
    # Color mapping
    scale_color_manual(values = cluster_colors) +
    scale_fill_manual(values = c(cluster_colors, "Hollow" = "white")) +
    # Reference line and theme
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
    theme_bw(base_size = 16) +
    labs(title = target_gwas_name, x = NULL, y = "tau_star") +
    theme(
      legend.position = "right",
      plot.title = element_text(size=10,hjust = 0.5),
      panel.grid.minor = element_blank()
    )
  
  return(p)
}

# 3. Generate the individual plots
p1 <- draw_single_gwas(clu_tau_results, "WHRadjBMI")
p2 <- draw_single_gwas(clu_tau_results, "IWB")
p3 <- draw_single_gwas(clu_tau_results, "Liver_fat")
p4 <- draw_single_gwas(clu_tau_results, "FIadj") 

# 4. Combine the plots with patchwork
# / stacks the plots vertically and guides = 'collect' merges the legends
combined_plot <- (p1 / p2 / p3 / p4) + 
  plot_layout(guides = 'collect')

# 5. Display the result
print(combined_plot)
ggsave("outplots/fig5_d_clu.tau.4gwas.forest.plots.pdf", combined_plot, width = 5, height = 8)
```


fig5_d Main panels show tau_star changes for the three GWAS WHRadjBMI/IWB/liver_fat, restricted to significant clusters
```{r}
library(dplyr)
library(stringr)


clu_tau_results_ac <- data.frame()
samplelist <- read.table(file = "period.sldsc.inputsamplelist")[,1]

for (i in 1:length(samplelist)){
    all_path <- "~/Project/GWAS/ldsc/S-LDSC"
    result_path_d0 <- "~/Project/GWAS/ldsc/S-LDSC-AC/d0"
    result_path_d7 <- "~/Project/GWAS/ldsc/S-LDSC-AC/d7"
    result_path_d15 <- "~/Project/GWAS/ldsc/S-LDSC-AC/d15"
    l0 <- tail(calculate_tau_star(paste0(all_path,'/',samplelist[i],'.results'), paste0(all_path,'/',samplelist[i],'.log')),7)
    l1 <- tail(calculate_tau_star(paste0(result_path_d0,'/',samplelist[i],'.results'), paste0(result_path_d0,'/',samplelist[i],'.log')),7)
    l2 <- tail(calculate_tau_star(paste0(result_path_d7,'/',samplelist[i],'.results'), paste0(result_path_d7,'/',samplelist[i],'.log')),7)
    l3 <- tail(calculate_tau_star(paste0(result_path_d15,'/',samplelist[i],'.results'), paste0(result_path_d15,'/',samplelist[i],'.log')),7)
    output <- data.frame(
        GWAS=samplelist[i],
        cluster=paste0('cluster',1:7),
        period=c(rep("ALL",7),rep("d0",7),rep("d7",7),rep("d15",7)),
        tau_star=c(l0$tau_star,l1$tau_star,l2$tau_star,l3$tau_star),
        tau_star_se=c(l0$tau_star_se,l1$tau_star_se,l2$tau_star_se,l3$tau_star_se),
        tau_star_p=c(l0$tau_star_p,l1$tau_star_p,l2$tau_star_p,l3$tau_star_p)
    )
    clu_tau_results_ac <- rbind(clu_tau_results_ac,output)
}
write.table(clu_tau_results_ac,file = "output/clu_tau_results_ac.tsv",row.names = F,col.names = T,sep = "\t",quote = F)



# 2. Data preprocessing
# Define colors
cluster_colors <- c(
    "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)

plot_data <- clu_tau_results_ac %>%
  # Ensure the period order
  mutate(period = factor(period, levels = c("ALL", "d0", "d7", "d15"))) %>%
  # Logic 1: check whether any period is significant for each cluster within each GWAS
  group_by(GWAS, cluster) %>%
  mutate(any_sig = any(tau_star_p <= 0.05)) %>%
  ungroup() %>%
  # Logic 2: assign colors according to any_sig, falling back to grey when not significant
  mutate(display_color = ifelse(any_sig, cluster, "NotSig")) %>%
  # Logic 3: decide whether the point is filled (p <= 0.05 is filled)
  mutate(is_solid = ifelse(tau_star_p <= 0.05, "Solid", "Hollow"))

# 3. Plotting
pd <- position_dodge(width = 0.7)
p <- ggplot(plot_data, aes(x = period, y = tau_star, group = cluster, color = display_color)) +
  # Draw the error bars (1.96 * SE)
  geom_errorbar(aes(ymin = tau_star - 1.96 * tau_star_se, 
                    ymax = tau_star + 1.96 * tau_star_se), 
                width = 0.4, position = pd, alpha = 0.8) +
  # Draw the connecting lines: only d0, d7 and d15 are linked, with ALL filtered out
  geom_line(data = filter(plot_data, period != "ALL"), 
            position = pd, size = 0.8) +
  # Draw the points: filled or open depending on significance
  # shape 21 is a fillable circle and 16 is a solid circle
  geom_point(aes(shape = is_solid, fill = display_color), 
             position = pd, size = 2.5, stroke = 1) +
  # Faceting: 2 rows by 5 columns
  facet_wrap(~GWAS, nrow = 2, ncol = 5, scales = "free_y") +
  # Set the colors: the seven cluster colors plus grey
  scale_color_manual(values = c(cluster_colors, "NotSig" = "gray80")) +
  scale_fill_manual(values = c(cluster_colors, "NotSig" = "gray80")) +
  # Set the shapes: Solid maps to a filled point (21 with fill) and Hollow to an open point (21 unfilled/white fill)
  scale_shape_manual(values = c("Solid" = 21, "Hollow" = 1)) +
  # Reference line: y = 0
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.3) +
  # Theme tweaks
  theme_bw(base_size = 16) +
  theme(
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    strip.background = element_rect(fill = "white"),
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  labs(title = "Forest Plot of tau_star by GWAS and Cluster",
       x = "Period",
       y = "tau_star",
       color = "Cluster Status",
       shape = "Significance (p < 0.05)")

# 4. Print the plot
print(p)


# Save the plot if needed
ggsave("outplots/sup5_b_GWAS_tar_star_ac_Forest_Plot.pdf", width = 15, height = 8)


draw_single_gwas <- function(data, target_gwas_name) {
  
  # A. Data preprocessing
    cluster_colors <- c(
    "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)
  cluster_input_list <- list(
    WHRadjBMI = c("cluster3","cluster5","cluster6"),
    logTG = c("cluster1","cluster5"),
    Liver_fat = c("cluster1","cluster3"),
    FIadj = c("cluster5","cluster7")
  )
  gwas_data <- data %>%
    filter(GWAS == target_gwas_name) %>%
    mutate(period = factor(period, levels = c("ALL", "d0", "d7", "d15"))) %>%
    # Keep only the clusters with a significant signal in this GWAS
    group_by(cluster) %>%
    filter(cluster %in% cluster_input_list[[target_gwas_name]]) %>%
    ungroup() %>%
    # Fill logic: significant points are filled and non-significant ones are open
    mutate(point_fill = ifelse(tau_star_p <= 0.05, cluster, "Hollow"))
  
  # B. Compute the symmetric Y axis range
  # Find the largest absolute value reached by the error bars
  max_abs_y <- gwas_data %>%
    summarise(val = max(c(abs(tau_star + 1.96 * tau_star_se), 
                          abs(tau_star - 1.96 * tau_star_se)), na.rm = TRUE)) %>%
    pull(val)
  
  y_limit <- max_abs_y * 1.1 # leave a 10% margin
  
  # C. Plotting
  pd <- position_dodge(width = 0.5)
  
  p <- ggplot(gwas_data, aes(x = period, y = tau_star, group = cluster, color = cluster)) +
    # Forest-plot error bars
    geom_errorbar(aes(ymin = tau_star - 1.96 * tau_star_se, 
                      ymax = tau_star + 1.96 * tau_star_se), 
                  width = 0.2, position = pd) +
    # Connecting lines (ALL excluded)
    geom_line(data = filter(gwas_data, period != "ALL"), position = pd, size = 0.8) +
    # Points (shape 21 supports fill)
    geom_point(aes(fill = point_fill), shape = 21, size = 3, stroke = 1, position = pd) +
    # Symmetric Y axis
    scale_y_continuous(limits = c(-y_limit, y_limit)) +
    # Color mapping
    scale_color_manual(values = cluster_colors) +
    scale_fill_manual(values = c(cluster_colors, "Hollow" = "white")) +
    # Reference line and theme
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray60") +
    theme_bw(base_size = 16) +
    labs(title = target_gwas_name, x = NULL, y = "tau_star") +
    theme(
      legend.position = "right",
      plot.title = element_text(size=10,hjust = 0.5),
      panel.grid.minor = element_blank()
    )
  
  return(p)
}


p1 <- draw_single_gwas(clu_tau_results_ac, "WHRadjBMI")
p2 <- draw_single_gwas(clu_tau_results_ac, "logTG")
p3 <- draw_single_gwas(clu_tau_results_ac, "Liver_fat")
p4 <- draw_single_gwas(clu_tau_results_ac, "FIadj") 

# 4. Combine the plots with patchwork
# / stacks the plots vertically and guides = 'collect' merges the legends
combined_plot <- (p1 / p2 / p3 / p4) + 
  plot_layout(guides = 'collect')

# 5. Display the result
print(combined_plot)
ggsave("outplots/fig5_e_clu.tau.ac.4gwas.forest.plots.pdf", combined_plot, width = 5, height = 8)


## Two additional panels are needed: one showing how H3K27me3 signal changes during differentiation for cluster12357, and one showing the H3K27ac signal changes


all.clu.coverage <- data.frame()
for (i in c(1,2,3,5,7)){
    a1 <- feats_coverage[roi_info.split[roi_info.split$cluster == i,"peak"],"H3K27ac_d0_coverage"]
    a2 <- feats_coverage[roi_info.split[roi_info.split$cluster == i,"peak"],"H3K27ac_d7_coverage"]
    a3 <- feats_coverage[roi_info.split[roi_info.split$cluster == i,"peak"],"H3K27ac_d15_coverage"]

    b1 <- feats_coverage[roi_info.split[roi_info.split$cluster == i,"peak"],"H3K27me3_d0_coverage"]
    b2 <- feats_coverage[roi_info.split[roi_info.split$cluster == i,"peak"],"H3K27me3_d7_coverage"]
    b3 <- feats_coverage[roi_info.split[roi_info.split$cluster == i,"peak"],"H3K27me3_d15_coverage"]
    output <- data.frame(
        cluster = paste0('cluster',i),
        period = rep(c("d0","d7","d15"),2),
        chip=c(rep("H3K27me3",3),rep("H3K27ac",3)),
        mean=c(mean(b1),mean(b2),mean(b3),mean(a1),mean(a2),mean(a3)),
        lower=c(quantile(b1,0.25),quantile(b2,0.25),quantile(b3,0.25),quantile(a1,0.25),quantile(a2,0.25),quantile(a3,0.25)),
        upper=c(quantile(b1,0.75),quantile(b2,0.75),quantile(b3,0.75),quantile(a1,0.75),quantile(a2,0.75),quantile(a3,0.75))
    )
    all.clu.coverage <- rbind(all.clu.coverage,output)
}


cluster_colors <- c(
    "cluster1" = "#2ca02c", # classic academic green
  "cluster2" = "#1b7837", # dark green       
  "cluster3" = "#3182bd", # standard blue    
  "cluster4" = "#08519c", # dark blue      
  "cluster5" = "#b10026", # scientific red     
  "cluster6" = "#feb24c", # soft orange-yellow  
  "cluster7" = "#f768a1"  # rose pink   
)

df <- all.clu.coverage
df$period <- factor(df$period,c("d0","d7","d15"))

# 4. Draw the H3K27me3 panel (solid outline)
p1 <- ggplot(df %>% filter(chip == "H3K27me3"), aes(x = period, y = mean, fill = cluster)) +

  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.8),
    color = "black",      # solid outline
    linetype = "solid",
    width = 0.8,linewidth=1.1
  ) +
  scale_fill_manual(values = cluster_colors) +
   coord_cartesian(ylim = c(0.5, 0.9)) +
  labs(title = "H3K27me3", y = "Mean Coverage", x = NULL) +
  theme_classic(base_size = 16) +
  theme(legend.position = "none") # temporarily hide the legend of the left panel

# 5. Draw the H3K27ac panel (dashed outline)
p2 <- ggplot(df %>% filter(chip == "H3K27ac"), aes(x = period, y = mean, fill = cluster)) +

  geom_bar(
    stat = "identity",
    position = position_dodge(width = 0.8),
    color = "black",      # black outline
    linetype = "dashed",   # dashed outline
    width = 0.8,linewidth=1.1
  ) +
  scale_fill_manual(values = cluster_colors) +
  labs(title = "H3K27ac", y = NULL, x = NULL) +
  theme_classic(base_size = 16)

# 6. Combine the panels with patchwork
# collect_guides = "collect" merges the legends of the two plots into one
combined_plot <- p1 + p2 + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "right")

# 7. Export the figure
print(combined_plot)
ggsave("outplots/fig5_d.K27me3&ac.allclu.barplots.pdf",combined_plot,width=10,height=4)

```


fig5_e Using WHRadjBMI as an example, two panels show the counts and the delta enrichment levels
UpSet plot showing how many clumped SNPs correspond to ROIs with no marked change, ROIs with a marked me3 change and ROIs with a marked ac change
```{r}

clu5.roi <- roi_info.split[roi_info.split$cluster == 5,c(1,2,3,4)]
clu5.roi$delta_cov_me3 <- feats_coverage[roi_info.split[roi_info.split$cluster == 5,"peak"],"H3K27me3_d15_coverage"]-feats_coverage[roi_info.split[roi_info.split$cluster == 5,"peak"],"H3K27me3_d0_coverage"]
clu5.roi$delta_cov_ac <- feats_coverage[roi_info.split[roi_info.split$cluster == 5,"peak"],"H3K27ac_d15_coverage"]-feats_coverage[roi_info.split[roi_info.split$cluster == 5,"peak"],"H3K27ac_d0_coverage"]

clu5.roi$me3_status <- "keeped"
clu5.roi$me3_status[which(clu5.roi$delta_cov_me3 <= -0.2)] <- "reduced"
clu5.roi$me3_status[which(clu5.roi$delta_cov_me3 >= 0.2)] <- "rised"

clu5.roi$ac_status <- "keeped"
clu5.roi$ac_status[which(clu5.roi$delta_cov_ac <= -0.2)] <- "reduced"
clu5.roi$ac_status[which(clu5.roi$delta_cov_ac >= 0.2)] <- "rised"


whr.summary <- fread(file = "~/Project/GWAS/ma/WHRadjBMI.ma")
whr.clumped.snplist <- read.table(file = "~/Project/GWAS/clump/WHRadjBMI_clump_lead_snps.txt")[,1]
whr.clumped <- whr.summary[whr.summary$SNP %in% whr.clumped.snplist,]
whr.clumped.bed <- data.frame(chr=paste0("chr",whr.clumped$CHR),start=whr.clumped$BP,end=whr.clumped$BP+1,rsid=whr.clumped$SNP)
whr.clumped.bed <- bt.sort(i=whr.clumped.bed,faidx="~/reference/hg19.chrom.sizes")
inter <- bt.intersect(a=whr.clumped.bed,b=clu5.roi,wo = T)
inter$status <- paste0(inter$V11,"-",inter$V12)

inter_data <- data.frame(
  rsid = inter[,4],
  me3_status = inter[,11],
  ac_status = inter[,12]
)

# 2. Convert the categorical variables into a binary matrix
# A 0/1 column is created for each state of me3 and ac
upset_input <- data.frame(
  rsid = inter_data$rsid,
  # the three me3 states
  me3_reduced = as.integer(inter_data$me3_status == "reduced"),
  me3_keeped  = as.integer(inter_data$me3_status == "keeped"),
  me3_rised   = as.integer(inter_data$me3_status == "rised"),
  # the three ac states
  ac_reduced  = as.integer(inter_data$ac_status == "reduced"),
  ac_keeped   = as.integer(inter_data$ac_status == "keeped"),
  ac_rised    = as.integer(inter_data$ac_status == "rised")
)

sets_to_show <- c("me3_keeped","ac_keeped","me3_reduced", "ac_rised", "me3_rised",   "ac_reduced" )

pdf("outplots/fig5_e.upset_plot_final.pdf", width = 7, height = 6) # the units are inches
library(UpSetR)
# Draw the UpSet plot
upset(upset_input, 
      sets = sets_to_show,
      keep.order = TRUE,               # keep the set order defined above
      main.bar.color = "steelblue",    # color of the top bar chart
      sets.bar.color = "gray30",       # color of the left bar chart
      mb.ratio = c(0.7, 0.3), # adjust the ratio between the top bars and the bottom matrix
      order.by = "freq",               # sort by intersection frequency
      queries = list(
        list(query = intersects, params = list("me3_reduced", "ac_rised"), color = "#be0f2d", active = T),
        list(query = intersects, params = list("me3_reduced", "ac_keeped"), color = "#E6A333", active = T),
        list(query = intersects, params = list("me3_keeped", "ac_rised"), color = "#D6604D", active = T)
      ),
      # in order: bar chart axis title, axis tick labels, set bar axis title, set bar tick labels, set names, numbers on the bars
      text.scale = c(2, 1.8, 1.8, 2, 1.8, 1.8), # adjust the font size
      number.angles = 0,               # angle of the numbers on the bars
      point.size = 4,                # point size
      line.size = 1.2                    # line thickness
)
dev.off()


# Next, draw a WHRadjBMI bar chart that directly shows delta tau*


input <- clu_tau_results[clu_tau_results$GWAS == "WHRadjBMI",]
input_ac <- clu_tau_results_ac[clu_tau_results_ac$GWAS == "WHRadjBMI",]

delta1_me3 <- input[input$period == "d7","tau_star"]-input[input$period == "d0","tau_star"]
delta2_me3 <- input[input$period == "d15","tau_star"]-input[input$period == "d7","tau_star"]

delta1_ac <- input_ac[input_ac$period == "d7","tau_star"]-input_ac[input_ac$period == "d0","tau_star"]
delta2_ac <- input_ac[input_ac$period == "d15","tau_star"]-input_ac[input_ac$period == "d7","tau_star"]

plot_df <- data.frame(
    period=c(rep("d7-d0",3),rep("d15-d7",3),rep("d7-d0",3),rep("d15-d7",3)),
    cluster = c(rep(paste0('cluster',c(1,3,5)),4)),
    chip=c(rep("H3K27me3 Defined",6),rep("H3K27ac Masked",6)),
    delta = c(delta1_me3[c(1,3,5)],delta2_me3[c(1,3,5)],delta1_ac[c(1,3,5)],delta2_ac[c(1,3,5)])
)
plot_df$period <- factor(plot_df$period,c("d7-d0","d15-d7"))

# 3. Define the color vector
cluster_colors <- c(
  "cluster1" = "#084594", "cluster2" = "#2171b5", "cluster3" = "#6baed6", 
  "cluster4" = "#bdd7e7", "cluster5" = "#b10026", "cluster6" = "#e31a1c", 
  "cluster7" = "#feb24c"
)

# 4. Write the plotting function with a linetype argument
draw_plot <- function(data, title_name, lt) {
  ggplot(data, aes(x = period, y = delta, fill = cluster)) +
    # color = "black" sets the border color
    # linetype sets the line type (1 = solid, 2 = dashed)
    # size sets the border thickness
    geom_bar(stat = "identity", 
             position = "dodge", 
             color = "black", 
             linetype = lt, 
             linewidth=1.0) + 
    scale_fill_manual(values = cluster_colors) +
    labs(title = title_name, x = "Period", y = "Delta") +
    theme_classic(base_size = 16) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.position = "right"
    )
}

# 5. Generate the two plots
# First plot: H3K27me3 Defined with a solid line (linetype = 1)
p1 <- draw_plot(subset(plot_df, chip == "H3K27me3 Defined"), "H3K27me3 Defined", 1)

# Second plot: H3K27ac Masked with a dashed line (linetype = 2)
p2 <- draw_plot(subset(plot_df, chip == "H3K27ac Masked"), "H3K27ac Masked", 2)

# 6. Combine the plots side by side with patchwork
# Merge the legends with collect and keep the outlines inside the legends consistent too
combined_plot <- p1 + p2 + plot_layout(guides = 'collect')

# 7. Display the figure
print(combined_plot)

ggsave("outplots/fig5_f.WHRadjBMI.delta.taustar.barplots.pdf", combined_plot, width = 7, height = 4)
```



```{r}

eqtl_list <- list()

eqtl_namelist <- read.table(file = "/storage/main/projects/bigcgpu-prj/zhangtianpei/reference/GTEX/GTEX.v7.eqtl.list.txt")[,1]
for (i in 1:length(eqtl_namelist)){
    ll <- fread(file = paste0('~/reference/GTEX/GTEx_Analysis_v7_eQTL/',eqtl_namelist[i],'.v7.signif_variant_gene_pairs.txt.gz'))
    eqtl_list[[eqtl_namelist[i]]] <- data.frame(
        chr = paste0('chr',sapply(strsplit(ll$variant_id,"_",fixed = T),"[",1)),
        start = as.integer(sapply(strsplit(ll$variant_id,"_",fixed = T),"[",2)),
        end = as.integer(sapply(strsplit(ll$variant_id,"_",fixed = T),"[",2)) + 1,
        gene_id = ll$gene_id, maf = ll$maf, pval = ll$pval_beta, beta = ll$slope, se = ll$slope_se
    )
    print(i)
}




all.eqtl.enrichment.results <- data.frame()
for (i in 1:7){
    input <- fread(file = paste0('~/Project/GWAS/output/eqtl.bg/cluster',i,'.eqtl.control.tsv')) %>% na.omit()
    input$gene_name <- ref[match(input$target_gene_id,ref$gene_id),"gene_name"]
    input.roi <- data.frame(input$chr,input$start,input$end,input$name,input$target_gene_id)
    input.control <- data.frame(input$control_chr,input$control_start,input$control_end,input$target_gene_id)
    all.results <- data.frame()
    for (j in 1:length(eqtl_list)){
        input.inter.tg <- intersect(input$target_gene_id,eqtl_list[[j]]$gene_id)
        eqtl <- eqtl_list[[j]][eqtl_list[[j]]$gene_id %in% input.inter.tg,]
        input.roi.eqtl <- bt.intersect(a=input.roi,b=eqtl,wo=T)
        input.roi.eqtl <- input.roi.eqtl[input.roi.eqtl$V5 == input.roi.eqtl$V9,1:5] %>% unique()
        input.control.eqtl <- bt.intersect(a=input.control,b=eqtl,wo=T)
        input.control.eqtl <- input.control.eqtl[input.control.eqtl$V4 == input.control.eqtl$V8,1:5] %>% unique()
        output <- fisher.test(matrix(c(nrow(input.roi.eqtl),nrow(input) - nrow(input.roi.eqtl),nrow(input.control.eqtl),nrow(input) - nrow(input.control.eqtl)),nrow=2,ncol=2))
        result <- data.frame(eqtl = names(eqtl_list)[j],cluster = paste0('cluster',i),or=output$estimate,lower=output$conf.int[1],upper = output$conf.int[2],pval = output$p.value)
        all.results <- rbind(all.results,result)
    }

    all.eqtl.enrichment.results <- rbind(all.eqtl.enrichment.results,all.results)
    rownames(all.eqtl.enrichment.results) <- 1:nrow(all.eqtl.enrichment.results)

}

all.eqtl.enrichment.results[all.eqtl.enrichment.results$eqtl == "Testis",]


df.plots <- all.eqtl.enrichment.results[all.eqtl.enrichment.results$eqtl %in% c("Adipose_Visceral_Omentum","Adipose_Subcutaneous","Whole_Blood"),]


# 1. Prepare the data (assuming your data frame is named df.plots)
# Keep only the three eqtl you specified
target_eqtls <- c("Adipose_Visceral_Omentum", "Adipose_Subcutaneous", "Whole_Blood")
df_filtered <- df.plots %>% 
  filter(eqtl %in% target_eqtls)

# 2. Preprocessing: fix the factor order so that the colors and the x-axis order are correct
df_filtered <- df_filtered %>%
  mutate(
    # Force the eqtl order to match the color order below
    eqtl = factor(eqtl, levels = target_eqtls),
    # Force the cluster order
    cluster = factor(cluster, levels = paste0("cluster", 1:7)),
    # Significance call
    is_significant = ifelse(pval < 0.05, "Significant", "Non-significant")
  )


# 3. Define the color vector
# Order: Visceral (red), Subcutaneous (orange), Whole_Blood (blue)
my_colors <- c(
  "Adipose_Visceral_Omentum" = "#E64B35", 
  "Adipose_Subcutaneous"     = "#F39B7F", 
  "Whole_Blood"              = "#3C5488"
)

# 4. Plotting
p <- ggplot(df_filtered, aes(x = cluster, y = or, color = eqtl, group = eqtl)) +
  # Reference line
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray60") +
  # Error bars (required for a forest plot)
  geom_errorbar(aes(ymin = lower, ymax = upper), 
                width = 0.4, 
                position = position_dodge(0.6),
                alpha = 1) +
  # Trend line
#   geom_line(position = position_dodge(0.6), 
#             size = 0.8, 
#             alpha = 0.6) +
  # OR points: filled (16) when significant and open (21) when not
  # shape = 21 would allow the interior to be controlled via fill, but 16 and 1 are used here for simplicity
  geom_point(aes(shape = is_significant), 
             size = 3.5, 
             position = position_dodge(0.6),
             stroke = 1) + # stroke controls the line thickness of the open circles
  # Set the shape manually
  scale_shape_manual(values = c("Significant" = 16, "Non-significant" = 1)) +
  # Set the colors manually
  scale_color_manual(values = my_colors) +
  # Axis and theme styling
  theme_classic(base_size = 16) +
  labs(
    title = "",
    x = "",
    y = "Odds Ratio (OR)",
    color = "eqtl",
    shape = "Significance (p < 0.05)"
  ) +
  theme(
    #panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1,color = "black"),
    axis.text.y = element_text(color = "black"),
    legend.position = "right"
  )
# Save the figure if needed
ggsave("outplots/fig5_d.eqtl_forest_plot.pdf",p, width = 10, height = 5)




whr.coloc <- fread(file = "~/Project/GWAS/coloc/Susie_coloc.WHRadjBMI.results.coverage.0.9")
whr.coloc <- whr.coloc[whr.coloc$cell %in% c("Adipose_Subcutaneous","Adipose_Visceral_Omentum"),]
whr.coloc <- cbind(data.frame(
    chr = paste0("chr",sapply(strsplit(whr.coloc$gwas_info,"_",fixed = T),"[",1)),
    start = paste0(sapply(strsplit(whr.coloc$gwas_info,"_",fixed = T),"[",2)),
    end = paste0(sapply(strsplit(whr.coloc$gwas_info,"_",fixed = T),"[",3))
),whr.coloc)



whr.locus <- data.frame(chr=whr.coloc$chr,start = whr.coloc$start,end = whr.coloc$end,gwas_info = whr.coloc$gwas_info,eqtl = whr.coloc$cell,gene_name = whr.coloc$ensg) %>% unique()


for (i in 1:7){
    clu.tg <- fread(file = paste0("~/Project/chromHMM/DiffCluster/output/peaks/cluster_tg/cluster",i,".rois.targetGene.tsv"))
    clu.tg$gene_name <- ref[match(clu.tg$target_gene_id,ref$gene_id),"gene_name"]

    l1 <- bt.intersect(a=whr.locus[whr.locus$eqtl == "Adipose_Subcutaneous",],b=clu.tg,wa=T) %>% unique()
    l1 <- l1[l1$V6 %in% clu.tg$gene_name,1:4] %>% unique

    input1 <- whr.coloc[whr.coloc$gwas_info %in% l1$V4 & whr.coloc$cell == "Adipose_Subcutaneous",]
    input1 <- input1[input1$PP.H4.abf >= 0.8,]
    input1 <- input1[input1$ensg %in% clu.tg$gene_name,]
    out1 <- unique(input1$gwas_info) %>% length

    l2 <- bt.intersect(a=whr.locus[whr.locus$eqtl == "Adipose_Visceral_Omentum",],b=clu.tg,wa=T) %>% unique()
    l2 <- l2[l2$V6 %in% clu.tg$gene_name,1:4] %>% unique
    input2 <- whr.coloc[whr.coloc$gwas_info %in% l2$V4 & whr.coloc$cell == "Adipose_Visceral_Omentum",]
    input2 <- input2[input2$PP.H4.abf >= 0.8,]
    input2 <- input2[input2$ensg %in% clu.tg$gene_name,]
    out2 <- unique(input2$gwas_info) %>% length
    print(c(out1/nrow(l1),out2/nrow(l2)))

}


whr.coloc.loci.annot <- data.frame()

for (i in 1:7){
    clu.tg <- fread(file = paste0("~/Project/chromHMM/DiffCluster/output/peaks/cluster_tg/cluster",i,".rois.targetGene.tsv"))
    clu.tg$gene_name <- ref[match(clu.tg$target_gene_id,ref$gene_id),"gene_name"]
    l1 <- bt.intersect(a=as.data.frame(whr.coloc),b=as.data.frame(na.omit(clu.tg)),wo = T)

    l2 <- l1[l1$V8 == "Adipose_Subcutaneous",]
    l3 <- l2[l2$V7 == l2$V26,] %>% na.omit()
    l4 <- l3[l3$V16 >= 0.8,]
    out1 <- data.frame(eqtl = "Adipose_Subcutaneous",cluster=paste0('cluster',i),ratio = nrow(l4)/nrow())
    nrow(l4)
}






```








fig5_g locuszoom plot to visualize marker genes
```{r}
library(locuszoomr)
library(EnsDb.Hsapiens.v75) # gene annotation database

# 1. Prepare your data (example dataframe)
# Must contain: chromosome, position, p
my_data <- fread("/storage/main/projects/bigcgpu-prj/zhangtianpei/Project/GWAS/ma/WHRadjBMI.ma", header=TRUE)
my_data <- data.frame(
    SNP=my_data$SNP,
    chromosome=my_data$CHR,
    position=my_data$BP,
    p=my_data$P
)




# 2. Define the region you want to plot
# This can be defined around the most significant SNP (index SNP)
loc <- locus(data = my_data, 
             gene = 'MLXIPL',   # or specify the range with seqname and xint
             flank = 5,    # extend 500 kb on each side
             ens_db = EnsDb.Hsapiens.v75)

# 3. Add LD information (optional; requires internet access to retrieve from LDlink)
# loc <- link_LD(loc, token = "YOUR_LDlink_TOKEN") 


# 4. Plotting

pdf("outplots/fig5_g.locuszoom.pdf",width=10,height=10)
locus_plot(loc)
dev.off()
```

